
# mlr3verse  
Functions and objects from the follow packages are imported by this meta package:  

|**Name**|**Title**|**URL**|
|:--|:--|:--|
|mlr3| Machine Learning in R - Next Generation|https://mlr3.mlr-org.com|
|mlr3cluster|Unsupervised Clustering|https://mlr3cluster.mlr-org.com|
|mlr3data|Additional data sets and tasks|https://mlr3data.mlr-org.com|
|mlr3filters|Filter Based Feature Selection|https://mlr3filters.mlr-org.com|
|mlr3select|Wrapper Based Feature Selection|https://mlr3fselect.mlr-org.com|
|mlr3learners|Recommended Learners|	https://mlr3learners.mlr-org.com|
|mlr3pipelines|Preprocessing Operators and Pipelines|	https://mlr3pipelines.mlr-org.com|
|mlr3proba|Probalistic Regression and Survival Analysis|https://mlr3proba.mlr-org.com|
|mlr3tuning|Hyperparameter Tuning|	https://mlr3tuning.mlr-org.com|
|mlr3vis|Visualizations|https://mlr3viz.mlr-org.com|
|paradox|Parameter Spaces|	https://paradox.mlr-org.com|


By loading the mlr3verse package, you are all set to deal with most regression, classification, cluster and survival tasks:

```{r}
library(mlr3verse)
```


If you want to get more detailed information about the loaded packages, you can call `mlr3verse_info()`:

```{r}
mlr3verse_info()
```

# Classification Learners

|**ID**|**Learner**|**Package**|
|:--|:--|:--|
|`classif.cv_glmnet`|Penalized Logistic Regression|`glmnet`|
|`classif.glmnet`|Penalized Logistic Regression|`glmnet`|
|`classif.kknn`|k-Nearest Neighbors|`kknn`|
|`classif.lda`|LDA|`MASS`|
|`classif.log_reg`|Logistic Regression|`stats`|
|`classif.multinom`|Multinomial log-linear model|`nnet`|
|`classif.naive_bayes`|Naive Bayes|`e1071`|
|`classif.nnet`|Single Layer Neural Network|`nnet`|
|`class.qda`|QDA|`MASS`|
|`classif.ranger`|Random Forest|`ranger`|
|`classif.svn`|SVM|`e1071`|
|`classif.xgboost`|Gradient Boosting|


# Regression Learners

|**ID**|**Learner**|**Package**|
|:--|:--|:--|
|`regr.cv_glmnet`|Penalized Linear Regression|`glmnet`|
|`regr.glmnet`|Penalized Linear Regression|`glmnet`|
|`regr.kknn`|k-Nearest Neighbors|`kknn`|
|`regr.lm`|Linear Regression|`stats`|
|`regr.ranger`|Random Forest|`ranger`|
|`regr.svm`|SVM|`e1071`|
|`regr.xgboost`|Gradient Boosting|`xgboost`|


# Survival Learners (via mlr3proba)

|**ID**|**Learner**|**Package**|
|:--|:--|:--|
|`surv.cv_glmnet`|Penalized Cox Regression|`glmnet`|
|`surv.glmnet`|Penalized Cox Regression|`glmnet`|
|`surv.ranger`|Random Forest|`ranger`|
|`surv.xgboost`|Gradient Boosting|xgboost|



github taken
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












