---
title: "tmap"
author: "YinCY"
date: "6/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Get started  
With the `tmap` package, thematic maps can be generated with great flexibility. The syntax for creating plots is similar to that of ggplot2, but tailored to maps.  

For more context on R’s geographic capabilities we recommend the online version of the book [Geocomputation with R](https://geocompr.robinlovelace.net/). The [Making maps with R](https://geocompr.robinlovelace.net/adv-map.html) chapter of the book provides many more context and abundant code examples of map making with `tmap` and other packages.  

A good place to start is to create a map of the world. After installing `tmap`, the following lines of code should create the map shown below:  

```{r}
library(tmap)
data("World")

tm_shape(World) +
  tm_polygons("HPI")
```

The object World is a spatial object of class sf from the `sf` package; it is a `data.frame` with a special column that contains a geometry for each row, in this case polygons. In order to plot it in `tmap`, you first need to specify it with `tm_shape`. Layers can be added with the `+` operator, in this case `tm_polygons`. There are many layer functions in tmap, which can easily be found in the documentation by their `tm_` prefix. See also `?'tmap-element'`.  

## Interactive maps
Each map can be plotted as a static image or viewed interactively using `"plot"` and `"view"` modes, respectively. The mode can be set with the function `tmap_mode`, and toggling between the modes can be done with the ‘switch’ `ttm()` (which stands for toggle thematic map.  

```{r}
tmap_mode(mode = "view")

tm_shape(World) +
    tm_polygons(col = "HPI")
```

## Multiple shapes and layers
A shape is a spatial object (with a class from `sf`, `sp`, `stars`, or `raster`). Multiple shapes and also multiple layers per shape can be plotted:  

```{r}
data(World, metro, rivers, land)

tmap_mode("plot")
tm_shape(shp = land) +
    tm_raster(col = "elevation", palette = terrain.colors(10)) +
tm_shape(shp = World) +
    tm_borders(col = "white", lwd = 0.5) +
    tm_text(text = "iso_a3", size = "area") +
tm_shape(shp = metro) +
    tm_symbols(col = "red", size = "pop2020", scale = 0.5) +
tm_legend(show = F)
```

## Facets
Facets can be created in three ways:  

1. By assigning multiple variable names to one aesthetic (in this example the first argument of `tm_polygons`):  

```{r}
tmap_mode("view")

tm_shape(World) +
    tm_polygons(col = c("HPI", "economy")) +
    tm_facets(sync = TRUE, ncol = 2)
```

2. By splitting the spatial data with the `by` argument of `tm_facets`:  

```{r}
tmap_mode("plot")

data("NLD_muni")
NLD_muni$perc_men <- NLD_muni$pop_men / NLD_muni$population * 100

tm_shape(NLD_muni) +
    tm_polygons(col = "perc_men", palette = "RdYlBu") +
    tm_facets(by = "province")
```

3. By using the `tmap_arrange` function:  

```{r}
tmap_mode(mode = "plot")

data("NLD_muni")
tm1 <- tm_shape(NLD_muni) + tm_polygons(col = "population", convert2density  =TRUE)
tm2 <- tm_shape(NLD_muni) + tm_bubbles(size = "population")

tmap_arrange(tm1, tm2)
```

## Basemaps and overlay tile maps
Tiled basemaps can be added with the layer function `tm_basemap`. Semi-transparent overlay maps (for example annotation labels) can be added with `tm_tiles`.  

```{r}
tmap_mode("view")

tm_basemap(server = "Stamen.Watercolor") +
   tm_shape(metro) + tm_bubbles(size = "pop2020", col = "red") +
    tm_tiles(server = "Stamen.TonerLabels")
```

## Options and styles
The functions `tm_layout` and `tm_view` are used to specify the map layout and the interactive aspects respectively. These functions can be used in the same way as the layer functions, e.g.  

```{r}
tmap_mode("plot")

tm_shape(World) +
    tm_polygons(col = "HPI") +
tm_layout(bg.color = "skyblue", inner.margins = c(0, 0.02, 0.02, 0.02))
```

These options, as well as a couple of others, can also be set within with `tmap_options`, which works in the same way as the base R function `options`. The main advantage is that these options are set globally, so they do not have to be specified in each map, for the duration of the session.  

```{r}
tmap_options(bg.color = "black", legend.text.color = "white")

tm_shape(World) +
    tm_polygons(col = "HPI", legend.title = "Happy Planet Index")
```

A style is a certain configuration of the tmap options.  

```{r}
tmap_style("classic")

tm_shape(World) +
    tm_polygons(col = "HPI", legend.title = "Happy Planet Index")
```

see what options have been changed  

```{r}
tmap_options_diff()
```

## Exporting maps
```{r}
tm <- tm_shape(World) +
    tm_polygons(col = "HPI", legend.title = "Happy Planet Index")

# save an image ("plot" mode)
tmap_save(tm, filename = "world_map.png")

# save as stand-alone HTML file ("view" mode)
tmap_save(tm, filename = "world_map.html")
```



