---
title: "TCGA F1000 Research"
author: "yincy"
date: "7/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Downloading data from TCGA data portal  
```{r}
library("TCGAbiolinks")
```




## Downloading data from Broad TCGA GDAC  
```{r, message=FALSE, warning=FALSE}
library(RTCGAToolbox)
```

The Bioconductor package `RTCGAToolbox` provides access to Firehose level 3 and 4 data through the function `getFirehoseData`.  

- `dataset` - tumor to download. A complete list of possibilities can be viewed with `getFirehoseDatasets` function.  
```{r}
getFirehoseDatasets()
```

- `runDate` - Stddata run dates. Dates can be viewed with `getFirehoseRunningDates()` function.  
```{r}
getFirehoseRunningDates()
```

- `gistic2_Date` - Analyze run dates. Dates can be viewed with `getFirehoseAnalyzeDates()` function.  
By default, `RTCGAToolbox` allows users to download up to 500 MB worth of data. To increase the size of the download, users are encouraged to use `fileSizeLimit` argument.   

get the last run dates  
```{r}
lastRunDate <- getFirehoseRunningDates()[1]

lastAnalyseDate <- getFirehoseAnalyzeDates(1)
```

get DNA methylation data, RNAseq2 and clinical data for LGG  
```{r}
lgg_data <- getFirehoseData(dataset = "GBM", 
                            runDate = lastRunDate, 
                            gistic2Date = getFirehoseAnalyzeDates(1), 
                            Methylation = TRUE, 
                            RNAseq2Norm = TRUE, 
                            clinical = TRUE, 
                            Mutation = TRUE, 
                            fileSizeLimit = 10000)
```


get DNA methylation data, RNAseq2 and clinical data for GBM  
```{r}
gbm_data <- getFirehoseData(dataset = "GBM")
```







