---
title: "visNetwork"
author: "yincy"
date: "12/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Minimal example
**visNetwork** needs at least two informations:  
- a nodes data.frame, with *id* column  
- a edges data.frame, with *from* and *to* columns  

# Minimal example  
```{r}
require(visNetwork, quietly = TRUE)
nodes <- data.frame(id = 1:3)
edges <- data.frame(from = c(1, 2), to = c(1, 3))
visNetwork(nodes = nodes, edges = edges, width = "100%")
```


## Simple individual nodes and edges customization  
- Adding more variables on nodes data.frame.  
```{r}
nodes <- data.frame(id = 1:10,
                    label = paste("Node", 1:10, sep = ""),
                    group = c("GrA", "GrB"),
                    value = 1:10,
                    shape = c("square", "triangle", "box", "circle", "dot", "star",
                    "ellipse", "database", "text", "diamond"),
                    title = paste0("<p><b>", 1:10, "</b><br>Node !<p>"),
                    color = c("darkred", "grey", "orange", "darkblue", "purple"),
                    shadow = c(F, T, F, T, T))

head(nodes)
```


- Adding more variables on edges data.frame.  
```{r}
edges <- data.frame(from = sample(1:10, 8), to = sample(1:10, 8),
                    label = paste("Edge", 1:8, sep = ""),
                    length = c(100, 500),
                    arrows = c("to", "from", "middle", "middle;to"),
                    dashes = c(T, F),
                    title = paste("Edge", 1:8),
                    smooth = c(F, T),
                    shadow = c(F, T, F, T))
head(edges)
```


```{r}
visNetwork(nodes = nodes,
           edges = edges,
           width = "100%")
```



## Global nodes/edges configuration  
- Set global optiona for nodes and edges using **visNodes** and **visEdges**, and use options per group using **visGroups**.  
```{r}
nodes <- data.frame(id = 1:5,
                    group = c(rep("A", 2), rep("B", 3)))

edges <- data.frame(from = c(2,5,3,3), 
                    to = c(1,2,4,2))

visNetwork(nodes, edges, width = "100%") %>% 
    visNodes(shape = "square") %>% 
    visEdges(arrows = "to") %>% 
    visGroups(groupname = "A", color = "darkblue") %>% 
    visGroups(groupname = "B", color = "red")
```


# Network configuration  
> Configuration options are available in *visOptions*, *visInteraction*, *visLayout*, *visHierarchicalLayout*, *visPhysics*:  

## Example Data  
```{r}
nb <- 10

nodes <- data.frame(id = 1:nb,
                    label = paste("Label", 1:nb),
                    group = sample(LETTERS[1:3], nb, replace = T), 
                    value = 1:nb,
                    title = paste0("<p>", 1:nb, "<br>Tooltip !</p>"),
                    stringsAsFactors = F)

edges <- data.frame(from = trunc(runif(nb)*(nb - 1)) + 1,
                    to = trunc(runif(nb) * (nb - 1)) + 1,
                    value = rnorm(nb, 10), label = paste("Edge", 1:nb),
                    title = paste0("<p>", 1:nb, "<br>Edge Tooltip !</p>"))
```


## Add legend  
New feature in **visNetwork**: you can now add more custom legend on nodes/edges!  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visLegend(width = 0.03, 
              position = "right")
```

Or passing data  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visLegend(useGroups = T, 
              addNodes = data.frame(label = "Nodes", shape = "circle"),
              addEdges = data.frame(label = "link", color = "black"), 
              position = "right")
```


## Highlight nearest  
You can highlight nearest nodes and edges clicking on a node with **highlightNearest** option:  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visOptions(highlightNearest = T)
```

Control the highlight degree  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visOptions(highlightNearest = list(enabled = T, degree = 1))
```


## Select by node id  
You can also seelct nodes by id/label with a list with **nodeIdSelection**:  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visOptions(highlightNearest = T, nodesIdSelection = T)
```


## Select by a column  
New feature in **visNetwork**: select some nodes by the values of a column using **selectedBy** option:  
```{r}
nodes$sel <- sample(c("sel1", "sel2"), nrow(nodes), replace = T)
visNetwork(nodes, edges, width = "100%") %>% 
    visOptions(selectedBy = "sel")
```


## Directed Network  
```{r}
visNetwork(nodes,edges, width =  "100%") %>% 
    visEdges(arrows = "to")
```


## Custom navigation   
*vis.js* propose some navigation tools:  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visInteraction(navigationButtons = T)
```


# Data Manipulation  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visOptions(manipulation = T)
```


## Hierarchical Layout  
```{r}
nodes <- data.frame(id = 1:7)

edges <- data.frame(from = c(1, 2, 2, 2, 3, 3),
                    to = c(2, 3, 4, 5, 6, 7))

visNetwork(nodes, edges, width = "100%", main = "Hierarchical layout") %>% 
    visEdges(arrows = "to") %>% 
    visHierarchicalLayout()
```

```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visEdges(arrows  ="from") %>% 
    visHierarchicalLayout(direction = "RL")
```


## Freeze network  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visInteraction(dragNodes = F, dragView = F, zoomView = F)
```

# Additional features  
New *visTree* function allows to cisulize and customize a rpart classification and regression tree. Have a look to *visTreeEditor* to edit and get back network, or to *visTreeModuleServer* to use custome tree module in R.  
```{r}
library(rpart)
data("solder")
res <- rpart(formula = Opening~., data = solder, control = rpart.control(cp = 0.00005))
visTree(object = res, height = "800px", nodesPopSize = T, minNodeSize = 10, maxNodeSize = 30)
```



## Use DOT language data  
```{r}
visNetwork(dot = 'dinetwork{1 -> 1 ->2; 2 -> 3; 2 -- 4, 2 -> 1}', width = "100%")
```



## Export 
```{r}
network <- visNetwork(nodes, edges, width = "100%")
visSave(graph = network, file = "network.html")
```


## Nodes  
### Individual configuration  
```{r}
nodes <- data.frame(
    id = 1:10,
    label = paste("Node", 1:10, sep = ""),
    group = c("GrA", "GrB"),
    value = 1:10,
    shape = c("square", "triangle", "box", "circle", "dot", "star", "ellipse", "database", "text", "diamond"),
    title = paste0("<p><b>", 1:10, "</b><br>Node !</b>"),
    color = c("darkred", "grey", "orange", "darkblue", "purple"),
    shadow = c(F, T, F, T, T)
)

head(nodes)
```


```{r}
edges <- data.frame(
    from = c(1, 2, 5, 7, 8, 10),
    to = c(9, 3, 1, 6, 4, 7)
)

visNetwork(nodes, edges, width = "100%", height = "300px")
```


### Global configuration  
If you want to set some configuration for all nodes, it's possible in a simple way, using the **visNodes** function directly.  
```{r}
nodes <- data.frame(id = 1:4)
edges <- data.frame(from = c(2, 4, 3, 3), to = c(1, 2, 4, 2))

visNetwork(nodes, edges, width = "100%") %>% 
    visNodes(shape = "square", color = list(background = "lightblue", 
                                            border = "darkblue",
                                            highlight = "yellow"),
             shadow = list(enable = T, size = 10)) %>% 
    visLayout(randomSeed = 12)
```

Combine individual and global options  
```{r}
nodes <- data.frame(
    id = 1:4, 
    shape = c("circle", "square"),
    label = LETTERS[1:4]
)


edges <- data.frame(from = c(2, 4, 3, 3), to = c(1, 2, 4, 2))

visNetwork(nodes, edges, width = "100%") %>% 
    visNodes(color = list(background = "lightblue",
                          border = "darkblue",
                          highlight = "yellow"),
             shadow = list(enable = T, size = 10)) %>% 
    visLayout(randomSeed = 12)
```


### Use complex configuration individually  
When options are available using a list, you can use it in a data.frame with specific notation like this (a dot per level)  
```{r}
nodes <- data.frame(id = 1:3,
                    color.background = c("red", "blue", "green"),
                    color.highlight.background = c("red", NA, "red"),
                    shadow.size = c(5, 10, 15))

edges <- data.frame(from = c(1, 2), 
                    to = c(1, 3),
                    label = LETTERS[1:2],
                    font.size = c(10, 20),
                    font.color = c("red", "blue"))

visNetwork(nodes, edges)
```


## Edges 
### Individual configuration  
*edge* must be a data.frame , with at least from and to columns, which make the link with id of nodes. You can add properties simply by adding variables on data.frame.   
```{r}
edges <- data.frame(
    from = sample(1:10, 8),
    to = sample(1:10, 8),
    label = paste("Edge", 1:8, sep = ""),
    length = c(100, 500),
    width = c(4, 1),
    arrows = c("to", "from", "middle", "middle;to"),
    dashes = c(T, F),
    title = paste("Edge", 1:8),
    smooth = c(F, T),
    shadow = c(F, T, F, T)
)

nodes <- data.frame(id = 1:10, group = c("A", "B"))

head(edges)
```


```{r}
visNetwork(nodes, edges, height = "400px", width = "100%")
```

### Global configuration  
If you want to set some configuration for all edges, it's possible in simper way, using **visEdges** function directly.  
```{r}
nodes <- data.frame(id = 1:4)
edges <- data.frame(from = c(2, 4, 3, 2), to = c(1,2,4,3))

visNetwork(nodes, edges, width = "100%") %>% 
    visEdges(shadow = T,
             arrows = list(to = list(enabled = T, scaleFactor = 2)),
             color = list(color = "lightblue", highlight = "red")) %>% 
    visLayout(randomSeed = 12)
```

### Combine individual and global options  
```{r}
nodes <- data.frame(id = 1:4, label = 1:4)
edges <- data.frame(
    from = c(2, 4, 3, 2),
    to = c(1, 2, 4, 3),
    dashes = c(T, F)
)

visNetwork(nodes, edges, width = "100%") %>% 
    visEdges(shadow = T,
             arrows = list(to = list(enabled = T, scaleFactor = 2)),
             color = list(color = "lightblue", highlight = "red")) %>% 
    visLayout(randomSeed = 12) %>% 
    visNodes(color = list(highlight = "dark"))
```

### Use complex configuration individually  
```{r}
nodes <- data.frame(
    id = 1:3,
    color = list(background = c("red", "blue", "green"), hightlight = c("red", NA, "red")),
    shadow = list(size = 5, 10, 15)
)

edges <- data.frame(
    from = c(1, 2),
    to = c(1, 3),
    label = LETTERS[1:2],
    font = list(color = c("red", "blue"),
                size = c(10, 20))
)

visNetwork(nodes, edges)
```


## Group  
You can affect group to a node using `group` column in the node data.frame, and then add a configuration per group using the **visGroups** function.  
```{r}
nodes <- data.frame(id = 1:5, group = c(rep("A", 2), rep("B", 3)))
edges <- data.frame(from = c(2, 5, 3, 3), to = c(1, 2, 4, 2))

visNetwork(nodes, edges, width = "100%") %>% 
    visGroups(groupname = "A", color = 'darkblue', shape = "square", shadow = list(enabled = T)) %>% 
    visGroups(groupname = "B", color = "red", shape = "triangle")
```

## Legend and Title  
### Title  
```{r}
nodes <- data.frame(id = 1:3, group = c("B", "A", "B"))
edges <- data.frame(from = c(1, 2), to = c(2, 3))

visNetwork(nodes, edges, main = "A really simple example", width = "100%")
```

### Legend based on groups  
We can simply add a legend, based on the definition of groups, using **visLegend**  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visGroups(groupname = "A", color = "red") %>% 
    visGroups(groupname = "B", color = "lightblue") %>% 
    visLegend()
```

### Placement and title  
It's also possible to adjust `position` (left or right), and `width`, and add a title (`main`)  
```{r}
visNetwork(nodes, edges, width = "100%") %>% 
    visGroups(groupname = "A", color = "red") %>% 
    visGroups(groupname = "B", color = "lightblue") %>% 
    visLegend(width = 0.1, position = "right", main = "Group")
```


### Custom nodes/edges  
you can add custom nodes(`addNodes`) and/or edges(`addEdges`) on the legend  
```{r}
lnodes <- data.frame(
    label = c("Group A", "Group B"),
    shape = c("ellipse")，
    color = c("red", "lightblue"),
    title = "Informations",
    id = 1:2
)

ledges <- data.frame(color = c("lightblue", "red"),
                     label = c("reverse", "depends"),
                     arrows = c("to", "from"))

visNetwork(nodes, edges, width = "100%") %>% 
    visGroups(groupname = "A", 
              color = "red") %>% 
    visGroups(groupname = "B", 
              color = "lightblue") %>% 
    visLegend(addEdges = ledges, 
              addNodes = lnodes, 
              useGroups = F,
              width = 0.2)
```


It's possible to use the definition of the groups(`useGroups = TRUE`) with custom nodes (`addNodes`) and/or edges(`addEdges`)  
```{r}
ledges <- data.frame(color = c("lightblue", "red"),
                     label = c("reverse", "depends"),
                     arrows = c("to", "from"))

visNetwork(nodes, 
           edges, 
           width = "100%") %>% 
    visGroups(groupname = "A", 
              color = "red") %>% 
    visGroups(groupname = "B", 
              color = "lightblue") %>% 
    visLegend(addEdges = ledges, 
              useGroups = T)
```

For more complex elements, if you prefer, you can use a **list** instead of **data.frame**  
```{r}
nodes <- data.frame(id = 1:3, group = c("B", "A", "B"))
edges <- data.frame(from = c(1, 2), to = c(2, 3))

visNetwork(nodes, edges) %>% 
    visGroups(groupname = "A", shape = "icon", icon = list(code = "f0c0", size = 75)) %>% 
    visGroups(groupname = "B", shape = "icon", icon = list(code = "f007", color = "red")) %>% 
    addFontAwesome() %>% 
    visLegend(addNodes = list(list(label = "Group", shape = "icon", icon = list(code = "f0c0", size = 25)),
                              list(label = "User", shape = "icon", icon = list(code = "f007", size = 50, color = "red"))),
              useGroups = F)
```


## Use image and icon  
### Images  
- to use images in nodes, you first have to set `shape` to **image** or **circularImage**.  
- Then, set the relative or absolute path on `image` value  
- An option is available in case of wrong path: `brokenImage`  
- And you can use borders with `shapeProperties` and `useBorderWithImage`  

```{r}
path_to_images <- "https://raw.githubusercontent.com/datastorm-open/datastorm-open.github.io/master/visNetwork/data/img/indonesia/"

nodes <- data.frame(
    id = 1:4, 
    shape = c("image", "circularImage"),
    image = paste0(path_to_images, 1:4, ".png"),
    label = "I'm an image"
)

edges <- data.frame(from = c(2, 4, 3, 3), to = c(1, 2, 4, 2))

visNetwork(nodes, edges, width = "100%") %>% 
    visNodes(shapeProperties = list(useBorderWithImage = T)) %>% 
    visLayout(randomSeed = 2)
```


## Options  
```{r}
nb <- 10
nodes <- data.frame(id = 1:nb, 
                    label = paste("Label", 1:nb), 
                    group = sample(LETTERS[1:3], nb, replace = T), 
                    value = 1:nb, 
                    title = paste0("<p>", 1:nb, "<br>Tooltip !</p>"), 
                    stringsAsFactors = F)

edges <- data.frame(from = c(8,2,7,6,1,8,9,4,6,2), 
                    to = c(3,7,2,7,9,1,5,3,2,9), 
                    value = rnorm(nb, 10), 
                    label = paste0("Edge", 1:nb), 
                    title = paste0("<p>", 1:nb, "<br>Edge Tooltip !</p>"))
```


### highlight nearest  
You can highlight nearest nodes and edges by clicking on a node with `highlightNearest`. Just click everywhere except on nodes to reset the network.  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(highlightNearest = T) %>% 
    visLayout(randomSeed = 123)
```


It's now possible to control the `degree` of depth, and to enable this option also hovering nodes (`hover`). Using `hover`, you can still use `click` to set a view.  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(highlightNearest =  list(enabled = T, degree = 2, hover = T)) %>% 
    visLayout(randomSeed = 123)
```

```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visGroups(groupname = "A", shape = "icon", icon = list(code = "foco", size = 75)) %>% 
    visGroups(groupname = "B", shape = "icon", icon = list(code = "f007", color = "red")) %>% 
    visGroups(groupname = "C", shape = "icon", icon = list(code = "f1b9", color = "black")) %>% 
    visOptions(highlightNearest = list(enabled = T, degree = 2, hover = T)) %>% 
    addFontAwesome() %>% 
    visLayout(randomSeed = 123)
```

### select by node id  
You can also seelct a node by id/label with a list using `nodesIdSelection`  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(highlightNearest = T, nodesIdSelection = T) %>% 
    visLayout(randomSeed = 123)
```


### Select by a column  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(selectedBy = "group") %>% 
    visLayout(randomSeed = 123)
```


```{r}
# or new column, with multiple groups  
nodes$sample <- paste(sample(LETTERS[1:3], nrow(nodes), replace = T), 
                      sample(LETTERS[1:3], nrow(nodes), replace = T), 
                      sep = ",")

nodes$label = nodes$sample # for see groups  
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(selectedBy = list(variable = "sample", multiple = T)) %>% 
    visLayout(randomSeed = 123)
```

### Customize options  
New features are available in `visNetwork` to customize `selectBy` and `nodesIdSelection`.  

- `values`: control the available values in the list  
- `selected`: initialize a selection  
- `style`: customize `css` of list  

```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visOptions(highlightNearest = T, 
               nodesIdSelection = list(enabled = T, 
                                       selected = "8", 
                                       values = c(5:10), 
                                       style = 'width: 200px; height: 26px; 
                                       background: #f8f8f8; 
                                       color: darkblue; 
                                       outline: none;
                                       border: none')) %>% 
    visLayout(randomSeed = 123)
```

### Collapse/Uncollapse Nodes  
```{r}
nodes <- data.frame(id = 1:15, 
                    label = paste("Label", 1:15), 
                    group = sample(LETTERS[1:3], 15, replace = T))

edges <- data.frame(from = trunc(runif(15) * (15-1)) + 1, 
                    to = trunc(runif(15)*(15-1)) + 1)

visNetwork(nodes = nodes, 
           edges = edges) %>% 
    visEdges(arrows = "to") %>% 
    visOptions(collapse = T)
```

### Data Manipulation  
Using `manipulation`, you can edit your network  

- Add/delete nodes/edges  
- Change links   

```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visLayout(randomSeed = 123) %>% 
    visOptions(manipulation = T)
```


### “Les miserables” example  
```{r}
nodes <- jsonlite::fromJSON("https://raw.githubusercontent.com/datastorm-open/datastorm-open.github.io/master/visNetwork/data/nodes_miserables.json")

edges <- edges <- jsonlite::fromJSON("https://raw.githubusercontent.com/datastorm-open/datastorm-open.github.io/master/visNetwork/data/edges_miserables.json")

visNetwork(nodes = nodes, 
           edges = edges, 
           height = "700px", 
           width = "100%") %>% 
    visOptions(selectedBy = "group", 
               highlightNearest = T, 
               nodesIdSelection = T, 
               collapse = T) %>% 
    visPhysics(stabilization = F)
```


## Layout  
The layout is configurable with `visLayout()` and `visHierarchicalLayout()` functions.  

### Set seed  
The nodes are randomly positioned initially. This means that the settled result is different every time. If you provide a random seed manually (`randomSeed`), then the layout will be the same every time.  


### Hierarchical Layout  
You can use and control hierarchical layout with some options  

- `levelSeparation`, the distance between the different levels  
- `direction`, the direction of the hierarchical layout  
- `sortMethod`, the algorithm used to ascertain the levels of the nodes based on the data  
```{r}
nodes <- data.frame(id = 1:7)
edges <- data.frame(from = c(1, 2, 2, 2, 3, 3), 
                    to = c(2, 3, 4, 5, 6, 7))
visNetwork(nodes = nodes, 
           edges = edges,
           width = "100%") %>% 
    visHierarchicalLayout() %>% 
    visEdges(arrows = "from")
```

```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           width = "100%") %>% 
    visEdges(arrows = "from") %>% 
    visHierarchicalLayout(direction = "LR", levelSeparation = 500)
```

It's also possible to define the `level` of each node  
```{r}
#with level 
nodes <- data.frame(id = 1:4, level = c(2, 1, 1, 1))
edges <- data.frame(from = c(1, 1, 1), 
                    to = c(2, 3, 4))

visNetwork(nodes = nodes, 
           edges = edges, 
           width = "100%") %>% 
    visEdges(arrows = "from") %>% 
    visHierarchicalLayout()
```

```{r}
# without level 
nodes$level <- NULL
visNetwork(nodes = nodes, 
           edges = edges, 
           width = "100%") %>% 
    visEdges(arrows = "from") %>% 
    visHierarchicalLayout()
```

## igraph  
It's possible to make the link between the features of package igraph and those of visNetwork  

- Use igraph layout to compute coordinates for fast rendering with `visIgraphLayout()`  
- Transform igraph network to visNetwork network using `visIgraph()` or `toVisNetworkData()`   

### Use igraph layout  
With `visIgraphLayout()`, you can use all available layouts in igraph and calculate coordinates before the sending to vis.js.  
```{r}
nnodes <- 100
nedges <- 100

nodes <- data.frame(id = 1:nnodes)
edges <- data.frame(from = sample(1:nnodes, nedges, replace = T), 
                    to = sample(1:nnodes, nedges, replace = T))

visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px") %>% 
    visIgraphLayout(layout = "layout_with_fr") %>% 
    visNodes(size = 10)
```

```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px") %>% 
    visIgraphLayout(layout = "layout_in_circle") %>% 
    visOptions(highlightNearest = list(enabled = T, hover = T), 
               nodesIdSelection = T)
```


### Use igraph network  
`visIgraph` directly transforms an `igraph` network into a **visNetwork**. It's a little bit experiemntal. You can also use `toVisNetworkData` to just get data into **visNetwork** format.  
```{r}
library(igraph)
igraph_network <- graph.famous(name = "Walther")
plot(igraph_network)
```


get data and plot  
```{r}
data <- toVisNetworkData(igraph_network)
visNetwork(nodes = data$nodes, edges = data$edges, height = "500px")
```

or plot directly  
```{r}
visIgraph(igraph_network)
```


## CART  
you can visualize Classification and Regression Trees from the output of the **rpart** package, simply using **visTree**.  

### Classification Tree  
```{r}
library(rpart)

# basic classification tree 
res <- rpart(Species ~., data = iris)
visTree(res, main = "Iris classification tree", width = "100%")
```

### Regression tree  
```{r}
res <- rpart(Petal.Length ~., data = iris)
visTree(object = res, edgesFontSize = 14, nodesFontSize = 16, width = "100%")
```


```{r}
data("solder")
res <- rpart(Opening ~., data = solder, control = rpart.control(cp = 0.00005))
visTree(res, height = "800px", width = "100%", nodesPopSize = T, minNodeSize = 10, maxNodeSize = 30)
```


## Interactions  
It's possible to control the interactions of the network with **visInteractions()**  

### Frozen network  

- `dragNodes`: enable or not the selection and movement of nodes (click on a node, and move your mouse)?  
- `dragView`: enable or not the movement of the full network (click everywhere except node, and move your mouse)?  
- `zoomView`: enable or not the zoom (use mouse scroll)?  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visInteraction(dragNodes = F, 
                   dragView = F, 
                   zoomView = F) %>% 
    visLayout(randomSeed = 123)
```


### Hide edges/nodes on drag  
Can increase the performance  

- `hideEdgesOnDrag`: hide edges when dragging the view  
- `hideNodesOnDrag`: hide nodes when dragging the view  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visInteraction(hideEdgesOnDrag = T) %>% 
    visLayout(randomSeed = 123)
```


### Navigation buttons  
We can show navigation buttons with `navigationButtons`  
```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visInteraction(navigationButtons = T, 
                   keyboard = T, 
                   tooltipDelay = 1)
```


## Physics  
The physics of the network is highly configurable  
Using **visPhysics()** function, you can play with the physics of the network:  

- choose the `solver` ('barnesHut', 'repulsion', 'hierarchicalRepulsion', 'forceAtlas2Based'), and set options.  

- control the `stabilization` processus.  

+ play with `maxVelocity` and `minVelocity`  
+ disable stabilization for rendering initial network and see stabilization processes  

```{r}
nodes <- data.frame(id = 1:10)

edges <- data.frame(from = round(runif(8)*10), 
                    to = round(runif(8)*10))

visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visPhysics(solver = "forceAtlas2Based", 
               forceAtlas2Based = list(gravitationalConstant = -500))
```


```{r}
visNetwork(nodes = nodes, 
           edges = edges, 
           height = "500px", 
           width = "100%") %>% 
    visPhysics(solver = "barnesHut")
```


## Configure tools  
Use **visConfigure()** to configure and view options directly on your network   
Don't show here. Test and open the network in a web browser. Can be usefull to configure your network as wanted.  
```{r}
nodes <- data.frame(id = 1:3, label = LETTERS[1:3])
edges <- data.frame(from = c(1, 2), to  = c(1, 3))
visNetwork(nodes, edges, width = "100%") %>% 
    visConfigure(enabled = T)
```



