---
title: "igraph"
author: "yincy"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load package, message=FALSE}
library(igraph)
library(magrittr)
library(tidyverse)
```

**all igraph's parameter is here!!**  
```{r}
?igraph.plotting
```

The discription of an `igraph` object starts with up to four letters  
```
- D or U for a directed or undirected graph  
- N for a named graph (where nodes have a name attribute)  
- W for a weighted graph (where edges have a weight attribute)  
- B for a bipartite (two-mode) graph (where nodes have a type attribute)  
```

```
(g/c) - graph-level character attribute  
(v/c) - vertex-level character attribute  
(e/n) - edge-level numeric attribute  
```


# Networks in igraph  
```{r}
g1 <- graph(edges = c(1,2, 2,3, 3, 1), n = 3, directed = F)
plot(g1)
```

```{r}
class(g1)
g1
```


```{r}
g2 <- graph( edges=c(1,2, 2,3, 3, 1), n=10 )
plot(g2)
```

when the edge list has vertex names, the number of nodes is not needed  
```{r}
g3 <- graph(edges = c("John", "Jim", "Jim", "Jill", "Jill", "John")) # named vertices  
plot(g3)
```

```{r}
g3
```

```{r}
g4 <- graph( c("John", "Jim", "Jim", "Jack", "Jim", "Jack", "John", "John"), 
             isolates=c("Jesse", "Janis", "Jennifer", "Justin") )

plot(g4, 
     edge.size = 0.5,
     vertex.color = "gold",
     vertex.size = 15,
     vertex.frame.color = "gray",
     vertex.label.color = "black",
     veretex.label.cex = 0.8,
     vertex.label.dist = 2,
     edge.curved = 0.2)
```


```{r}
plot(graph_from_literal(a---b, b---c)) # the number of dashes doesn't matter
```


```{r}
plot(graph_from_literal(a--+b, b+--c))
```

```{r}
plot(graph_from_literal(a+-+b, b+-+c))
```


```{r}
g1 <- graph_from_literal(a-b-c-d-e-f, a-g-h-b, h-e:f:i, j)
plot(g1)
```


## edge, vertex, and network attributes  
access vertices and edges  
```{r}
E(g4)
```

```{r}
V(g4)
```

```{r}
g4[] # adjacency matrix, the default direction is row to column
```


```{r}
g4[1, ]
```


Add attributes to the network, vertices, or edges  
```{r}
V(g4)$name
```


```{r}
V(g4)$gender <- c("male", "male", "male", "male", "female", "female", "male")
E(g4)$type <- "email"
E(g4)$weight <- 10
```



```{r}
edge_attr(g4)
```

```{r}
vertex_attr(g4)
```

```{r}

```


```{r}
graph_attr(g4, name = "description") <- "The first graphic attributs"
graph_attr(g4)
```


Another way to set attributes  
```{r}
g4 <- set_graph_attr(g4, name = "name", value = "Email Network")
g4 <- set_graph_attr(g4, name = "something", value = "a thing")

graph_attr(g4)
```


```{r}
graph_attr(g4, name = "name")
```


```{r}
g4 <- delete_graph_attr(graph = g4, name = "something")
```

```{r}
graph_attr(g4)
```


```{r}
plot(g4,
     edge.arrow.size = 0.5, 
     vertex.label.color = "black",
     vertex.label.dist =1.5,
     vertex.color = c("pink", "skyblue")[1 + (V(g4)$gender == "male")])
```


simplify graph to remove loops and multiple edges between the same nodes  
`edge.attr.comb` possible options: sum, mean, prod, min, max, first, last and ignore.  
`ignore`: the attributes be disregarded and dropped  
```{r}
g4s <- simplify(graph = g4, 
                remove.loops = TRUE,
                remove.multiple = TRUE,
                edge.attr.comb = c(type = "ignore", weight = "sum"))
```

```{r}
plot(g4s,
     vertex.label.dist = 1.5, 
     edge.arrow.size = 0.4)
```

```{r}
g4s
```


## Specific graphs and graph models  
empty graph  
```{r}
eg <- make_empty_graph(40)
plot(eg, vertex.size = 10,
     vertex.label = NA)
```


Full graph  
```{r}
fg <- make_full_graph(n = 40)

plot(fg, 
     vertex.size = 10,
     vertex.label = NA)
```

Simple star graph  
```{r}
st <- make_star(40)

plot(st, 
     vertex.size = 10,
     vertex.label = NA)
```

Tree graph  
```{r}
tr <- make_tree(n = 100,
                children = 3,
                mode = "undirected")

l <- layout_with_fr(tr) 
```

```{r}
par(mar = c(0, 0, 0, 0))
plot(tr, 
     vertex.size = 80,
     vertex.label = NA, 
     layout = l, 
     rescale = F, 
     xlim = c(-5, 5), 
     ylim = c(-20,22), 
     vertex.color = "orange", 
     vertex.frame.color = "grey")
```


Ring graph  
```{r}
rn <- make_ring(n = 40)

plot(rn, 
     vertex.size = 10,
     vertex.label = NA)
```


Erdos-Renyi random graph model  
```{r}
er <- sample_gnm(n = 100, m = 40)

plot(er,
     vertex.size = 6,
     vertex.label = NA)
```

Watts-Strogatz small-world model  
```{r}
sw <- sample_smallworld(dim = 2, 
                        size = 10,
                        nei = 1, 
                        p = 0.1)

plot(sw,
     vertex.size = 6,
     vertex.label = NA,
     layout = layout_in_circle)
```


Barabasi-Albert preferential attachment model for scale-free graphs  
```{r}
ba <- sample_pa(n = 100, 
                power = 1,
                m = 1,
                directed = F)

plot(ba, vertex.size = 6,
     vertex.label = NA)
```

```{r}
zach <- graph(edges = "Zachary")

plot(zach,
     vertex.size = 10,
     vertex.label = NA)
```

Rewiring a graph  
`each_edge()` is a rewiring method that changes the edge endpoints uniformaly randomly with a probability **prob**  
```{r}
rn.rewiring <- rewire(rn, each_edge(prob = 0.1))

plot(rn.rewiring, vertex.size = 10, vertex.label = NA)
```

Rewire to connect vertices to other vertices at a certain distance.  
```{r}
rn.neigh <- connect.neighborhood(graph = rn, order = 5)

plot(rn.neigh, vertex.size = 8, vertex.label = NA)
```


Combine graphs (disjoint union, assuming separate vertex sets): **%du%**  
```{r}
par(mfrow = c(1, 2))
plot(rn, vertex.size = 10, vertex.label = NA)
plot(tr, vertex.size = 10, vertex.label = NA)
```

```{r}
plot(rn %du% tr, vertex.size = 10, vertex.label = NA)
```


# Reading network data from files  
## Dataset 1 edgelist  
```{r}
nodes <- read.csv(file = "Dataset1-Media-Example-NODES.csv", header = TRUE, as.is = TRUE)
# the first column is the vertex id, other column trited as vertex attributes

links <- read.csv(file = "Dataset1-Media-Example-EDGES.csv", header = TRUE, as.is = TRUE)
# the first two columns are the vertex ids thant this edge between, and the other columns are the edge attributes
```

the number of nodes and the unique links between nodes (directed, from --> to)  

The number of unique links less than the number of total links(49 < 52), so duplicated links are exists.  
```{r}
nrow(nodes); unique(links[, c("from", "to")]) %>% nrow()
```

```{r}
links %>% distinct(from, to) %>% nrow()
```

 Summary Statistics of Data Subset  
```{r}
?aggregate
links <- aggregate(x = links[, 3], by = links[, -3], FUN = sum)
links <- links[order(links$from, links$to), ]
# reorder links according to from and to 
```


```{r}
colnames(links)[4] <- "weight"
rownames(links) <- NULL
```

## Dataset 2 matrix  
```{r}
nodes2 <- read.csv(file = "Dataset2-Media-User-Example-NODES.csv", header = TRUE, as.is = TRUE)
links2 <- read.csv(file = "Dataset2-Media-User-Example-EDGES.csv", header = TRUE, as.is = TRUE)
```


```{r}
links2 <- as.matrix(links2)
dim(links2); dim(nodes2)
```


## Turning networks into igraph objects  
```{r}
net <- graph_from_data_frame(d = links, vertices = nodes, directed = TRUE)
class(net)
```

```{r}
net
```

```{r}
E(net)
```

```{r}
V(net)
```


```{r}
E(net)$type
V(net)$media
```

```{r}
edge_attr(net)
print("-------------------------")
vertex_attr(net)
```


```{r}
par(mar = c(0, 0 , 0, 0))
plot(net,
     edge.arrow.size = 0.4,
     vertex.label = V(net)$media, 
     vertex.label.cex = 0.7, 
     vertex.label.dist = 1.5, 
     layout = layout_with_fr, 
     edge.curved = 0.1, 
     vertex.color = brewer.pal(n = 3, name = "Dark2")[V(net)$media.type], 
     edge.width = E(net)$weight * 0.2, 
     edge.arrow.size = 3, 
     edge.lty = c(1, 3)[as.factor(E(net)$type)])
```

```{r}
net <- simplify(graph = net, 
                remove.loops = TRUE, 
                remove.multiple = F,
                edge.attr.comb = list(weight = "sum", type = "first"))

par(mar = c(0, 0 , 0, 0))
plot(net,
     edge.arrow.size = 0.4,
     vertex.label = V(net)$media, 
     vertex.label.cex = 0.7, 
     vertex.label.dist = 1.5, 
     layout = layout_with_fr, 
     edge.curved = 0.1, 
     vertex.color = brewer.pal(n = 3, name = "Dark2")[V(net)$media.type], 
     edge.width = E(net)$weight * 0.2, 
     edge.arrow.size = 3, 
     edge.lty = c(1, 3)[as.factor(E(net)$type)])
```


```{r}
net2 <- graph_from_incidence_matrix(incidence = links2)
net2
```

To transform a one-mode network matrix into an igraph object, use instead `graph_from_adjacency_matrix`.   

```{r}
colors <- c("gray50", "tomato", "gold")
V(net)$color <- colors[V(net)$media.type]
edge.start <- ends(net, es = E(net), names = F)[,1]
edge.col <- V(net)$color[edge.start]

plot(net, edge.color = edge.col, edge.curved = 0.1, vertex.label = NA)
```

By default, the coordinates of the plots are rescaled to the [-1,1] interval for both x and y. You can change that with the parameter `rescale=FALSE` and rescale your plot manually by multiplying the coordinates by a scalar.  

You can use `norm_coords` to normalize the plot with the boundaries you want.  

```{r}
net.bg <- sample_pa(n = 80)
V(net.bg)$size <- 8
V(net.bg)$frame.color <- "white"
V(net.bg)$label <- NA
E(net.bg)$arrow.mode <- 0

l <- layout_with_fr(net.bg)
ll <- norm_coords(layout = l, ymin = -1, ymax = 1, xmin = -1, xmax = 1)

par(mfrow = c(2, 2), mar = c(0, 0, 0, 0))
plot(net.bg, rescale = F, layout = ll*0.4)
plot(net.bg, rescale = F, layout = ll*0.6)
plot(net.bg, rescale = F, layout = ll*0.8)
plot(net.bg, rescale = F, layout = ll*1.0)
```



```{r}
layouts <- grep("^layout_", ls("package:igraph"), value = T)[-1]

layouts <- layouts[!grepl("bipartite|merge|norm|sugiyama|tree", layouts)]
```

```{r}
par(mfrow = c(5, 3), mar = c(1, 1, 1, 1))

for(layout in layouts){
        l <- do.call(what = layout, args = list(graph = net))
        plot(net, 
             edge.arrow.mode = 0, 
             layout = l, 
             main = layout,
             vertex.label = NA)
}
```



```{r}
net.m <- net - E(net)[E(net)$type == "hyperlink"]
net.h <- net - E(net)[E(net)$type == "mention"]
```


```{r}
par(mfrow = c(1,2))
plot(net.h, vertex.color = "orange", main = "Tie:Hyperlink", vertex.label = NA, rescale = T)
plot(net.m, vertex.color = "orange", main = "Tie:mention", vertex.label = NA, rescale = T)
```



# Network and node descriptives  
### density  
```{r}
edge_density(net, loops = F)
```

```{r}
ecount(net) / (vcount(net)*(vcount(net) - 1))
```

### Reciprocity: the proportion of reciprocated ties (for a directed network)  
```{r}
reciprocity(net)

dyad_census(net) # mutal, asymmetric and null node pairs

2*dyad.census(net)$mut / ecount(net) # for a directed network 
```

```{r}
plot(net,edge.arrow.width = 0.5)
```


### Transitivity  
- global - ratio of triangles (directed disregarded) to connected triples  
- local - ratio of triangles to connected triples each vertex is part of.  
```{r}
transitivity(net, type = "global") # net is treated as an undirected network 
transitivity(as.undirected(net, mode = "collapse")) # same as above
transitivity(net, type = "local")
triad.census(net) # for directed networks  
```

### Diameter   
A network diameter is the longest geodesic distance (length of the shortest path between two nodes) in the network.     

In igraph, `diameter()` returns the distance, while `get_diameter()` returns the nodes along the first found path of that distance.  
Note that edge weights are used by defualt, unless set to NA  
```{r}
diameter(net, directed = F, weights = NA)
```

```{r}
diameter(net, directed = F)
```

```{r}
diam <- get_diameter(net, directed = F) # get the nodes 
```

Note though that when asked to behaved as a vector, a vertex sequence will produce the numeric indexes of the nodes in it. The same applies for edge sequences.  

```{r}
as.vector(diam)
```

```{r}
vcol <- rep("grey40", vcount(net))
vcol[diam] <- "gold"

ecol <- rep("grey80", ecount(net))
ecol[E(net, path = diam, directed = F)] <- "orange"

par(mar = c(0, 0, 0, 0))
plot(net, 
     vertex.color = vcol,
     edge.color = ecol,
     edge.arrow.mode = 0, 
     vertex.size = V(net)$audience.size, 
     layout = layout_with_lgl,
     rescale = T)
```


### Node degrees  
```{r}
deg <- degree(net, mode = "all")
plot(net, vertex.size = deg*3,
     vertex.label = NA)
```

```{r}
hist(deg, breaks = 1:vcount(net)-1, main = "Histogram of node degree")
```

### degree distribution  
```{r}
deg.dist <- degree_distribution(graph = net, cumulative = TRUE, mode = "all")

plot(x = 0:max(deg), y = 1-deg.dist, pch = 19, cex = 1.2, col = "orange", xlab = "Degree")
```


### Centrality and centralization  
Centrality functions (vertex level) and centralization functions (graph level).  

The centralization functions return `res` - vertex centrality, centralization, and theoretical_max - maximum centralization score for a graph of that size  

The centrality function can run on a subset of nodes (set with the vids parameter).  

degree: number of ties  
```{r}
degree(net, mode = "in")
```

```{r}
centr_degree(net, mode = "in", normalized = TRUE)
```


Closeness:(centrality based on distance to others in the graph)  
```{r}
closeness(net, mode = "all", weights = NA)
centr_clo(net, mode = "all", normalized = T)
```


Eigenvector: centrality proportional to the sum of connection centralities  
```{r}
eigen_centrality(net, directed = T, weights = NA)
centr_eigen(net, directed = T, normalized = T)
```



Betweenness: centrality based on a broker position connecting others  
```{r}
betweenness(net, directed = T, weights = NA)
edge_betweenness(net, directed = T, weights = NA)
centr_betw(net, directed = T, normalized = T)
```


### Hubs and authorities  
Hubs were expected to contain catalogs with a large number of outgoing links; while authorities would get many incoming links from hubs, presumably because of their high-quality relevant information.  

```{r}
hs <- hub_score(net, weights = NA)$vector
as <- authority_score(net, weights = NA)$vector
```

```{r}
par(mfrow = c(1, 2), mar = c(0, 0, 2, 0))
plot(net, 
     vertex.size = hs*20, 
     main = "Hubs",
     vertex.label = V(net)$name,
     edge.arrow.size = 0.5, 
     vertex.label.dist = 1.8)

plot(net, 
     vertex.size = as*20, 
     main = "Authorities",
     vertex.label = V(net)$name,
     edge.arrow.size = 0.5, 
     vertex.label.dist = 1.8)
```


### Distances and paths  
**Average path length**: the mean of the shortest distance between each pair of nodes in the network (in both directions for directed graphs).  
```{r}
mean_distance(net, directed = F)
```

```{r}
mean_distance(net, directed = T)
```

find all shortest paths in the graph  
```{r}
distances(net) # with edge weights
distances(net, weights = NA) # ignore weights
```

get the distance of every media from the New York Times  
```{r}
dist.from.NYT <- distances(graph = net, v = V(net)[media == "NY Times"], to = V(net), weights = NA)
```

```{r}
oranges <- colorRampPalette(colors = c("dark red", "gold"))
col <- oranges(max(dist.from.NYT)+1)
col <- col[dist.from.NYT + 1]
```

```{r}
plot(net,
     vertex.color = col,
     vertex.label = dist.from.NYT,
     edge.arrow.size = 0.6,
     vertex.label.color = "white")
```


find shortest path between specific nodes  
```{r}
news.path <- shortest_paths(graph = net,
                            from = V(net)[media == "MSNBC"],
                            to = V(net)[media == "New York Post"],
                            output = "both")


```

```{r}
ecol <- rep("gray80", ecount(net))
ecol[unlist(news.path$epath)] <- "orange"

ew <- rep(2, ecount(net))
ew[unlist(news.path$epath)] <- 4


vcol <- rep("gray70", vcount(net))
vcol[unlist(news.path$vpath)] <- "gold"

vlab <- V(net)$media
vlab[setdiff(x = 1:vcount(net), y = unlist(news.path$vpath))] <- NA
# not selected vertices do not display label

par(mar = c(0, 0, 0, 0))
plot(net,
     vertex.color = vcol,
     edge.color = ecol,
     edge.width = ew,
     edge.arrow.mode = 0,
     vertex.label = vlab,
     vertex.label.cex = 0.7,
     vertex.label.color = "blue")
```


Identify the edges going into or out of a vertex, for instance the WSJ. For a single node, use `incident()`, for multiple nodes use `incident_edges()`.  

```{r}
inc.edges <- incident(graph = net, 
                      v = V(net)[media == "Wall Street Journal"], 
                      mode = "out")

ecol <- rep("gray80", ecount(net))
ecol[unlist(inc.edges)] <- "orange"
```

```{r}
vcol <- rep("gray40", vcount(net))
vcol[V(net)$media == "Wall Street Journal"] <- "gold"

par(mar = c(0, 0, 0, 0))
plot(net, 
     vertex.color = vcol, 
     edge.color = ecol, 
     edge.arrow.size = 0.5,
     vertex.label = NA, 
     vertex.size = 10)
```

find immediate neighbors of a vertex  

`neighbors()` finds all nodes one step out from the focal actor.  
```{r}
neigh.nodes <- neighbors(graph = net, v = V(net)[media == "Wall Street Journal"], mode = "out")
vcol <- rep("gray40", vcount(net))
vcol[neigh.nodes] <- "tomato"
vcol[V(net)$media  == "Wall Street Journal"] <- "orange"
vlab <- V(net)$media
vlab[setdiff(x = 1:vcount(net), y = unlist(neigh.nodes))] <- NA

plot(net, 
     vertex.color = vcol,
     vertex.label = vlab,
     edge.arrow.size = 0.5,
     vertex.label.cex = 0.7)
```


To find the neighbors for multiple nodes, use `adjacent_vertices()` instead of `neighbors()`.  

To find node neighborhoods going more than one step out, use function `ego()` with parameter order set to the number of steps out to go from the focal node(s).  

Special operators for the indexing of edge sequences: %-%, %->%, %<-%  
```
E(network)[X %-% Y]: selects edges between vertex sets X and Y, ignoring direction  

E(network)[X %->% Y]: selects edges from vertex sets X to vertex set Y  

E(network)[X %<-% Y]: selects edges from vertex sets Y to vertex set X  
```

example: select edges from newspapers to online sources:  
```{r}
E(net)[V(net)[type.label == "Newspaper"] %->% V(net)[type.label == "Online"]]
```


### Subgroups and communities  
```{r}
net.sym <- as.undirected(graph = net,
                         mode = "collapse",
                         edge.attr.comb = list(weight = "sum", "ignore"))
```


```{r}
sub_graph <- cliques(net.sym) # list of cliques  
sub_graph %>% length()
```


```{r}
sapply(cliques(net.sym), length) #cliques size
```

```{r}
largest.cliques(net.sym) # cliques with max number of nodes
```


```{r}
vcol <- rep("gray80", vcount(net.sym))
vcol[unlist(largest_cliques(net.sym))] <- "gold"
plot(as.undirected(net.sym), vertex.label = V(net.sym)$name, vertex.color = vcol)
```


### community detection  
A number of algorithms aim to detect groups that consist of densely connectedd nodes with fewer connections across groups.  

**Community detection based on edge betweenness (Newman-Girvan)**  
High-betweenness edges are removed sequentially (recalculating at each step) and the best partitioning of the network is selected.  
```{r}
ceb <- cluster_edge_betweenness(net)
dendPlot(ceb, mode = "hclust")
```


```{r}
par(mar = c(0,0,0,0))
plot(x = ceb, y = net, 
     edge.arrow.size = 0.3, 
     edge.width = 2, 
     edge.color = "grey")
```



```{r}
class(ceb)
length(ceb) # number of communities
membership(ceb) # community membership for each node
modularity(ceb) # how modular the graph partitioning is
crossing(ceb, net) # boolean vector: TRUE for edges across communities
```

High modularity for a partitioning reflects dense connections within communities and sparse connections across communities.  


**Community detection based on propagating labels**  
Assigns node labels, randomizes, than replaces each vertex’s label with the label that appears most frequently among neighbors. Those steps are repeated until each vertex has the most common label of its neighbors.  
```{r}
clp <- cluster_label_prop(graph = net, weights = E(net)$weight)
```

```{r}
par(mar = c(0,0,0,0))
plot(x = clp, 
     y = net, 
     edge.arrow.size = 0.4, 
     edge.color = "grey", 
     edge.curved = 0.2, 
     edge.width = (E(net)$weight %>% scale(center = -2) %>% as.numeric())*5)
```


**Community detection based on greedy optimization of modularity**  
```{r}
# fast greedy only works for undirected net
cfg <- cluster_fast_greedy(as.undirected(net), weights = E(net)$weight)
```

```{r}
par(mar = c(0, 0, 0, 0))
plot(cfg, as.undirected(net, edge.attr.comb = c(weight = "sum", "ignore")), 
     edge.width = (E(net)$weight %>% scale(center = -2) %>% as.numeric())*5, 
     EDGE.ARROW.SIZE = 0.2, 
     edge.color = "grey", 
     edge.curved = 0.2)
```



```{r}
V(net)$community <- cfg$membership
colors <- adjustcolor(col = c("gray50", "tomato", "gold", "yellowgreen"), alpha = 0.6)
plot(net,
     vertex.color = colors[V(net)$community],
     edge.arrow.size = 0.3,
     vertex.label = NA)
```



