---
title: "igraph"
author: "yincy"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load package, message=FALSE}
library(igraph)
library(magrittr)
```

**all igraph's parameter is here!!**  
```{r}
?igraph.plotting
```

The discription of an `igraph` object starts with up to four letters  
```
- D or U for a directed or undirected graph  
- N for a named graph (where nodes have a name attribute)  
- W for a weighted graph (where edges have a weight attribute)  
- B for a bipartite (two-mode) graph (where nodes have a type attribute)  
```

```
(g/c) - graph-level character attribute  
(v/c) - vertex-level character attribute  
(e/n) - edge-level numeric attru=ibute 
```


# Networks in igraph  
```{r}
g1 <- graph(edges = c(1,2, 2,3, 3, 1), n = 3, directed = F)
plot(g1)
```

```{r}
class(g1)
g1
```


```{r}
g2 <- graph( edges=c(1,2, 2,3, 3, 1), n=10 )
plot(g2)
```

when the edge list has vertex names, the number of nodes is not needed  
```{r}
g3 <- graph(edges = c("John", "Jim", "Jim", "Jill", "Jill", "John")) # named vertices  
plot(g3)
```

```{r}
g3
```

```{r}
g4 <- graph( c("John", "Jim", "Jim", "Jack", "Jim", "Jack", "John", "John"), 
             isolates=c("Jesse", "Janis", "Jennifer", "Justin") )

plot(g4, 
     edge.size = 0.5,
     vertex.color = "gold",
     vertex.size = 15,
     vertex.frame.color = "gray",
     vertex.label.color = "black",
     veretex.label.cex = 0.8,
     vertex.label.dist = 2,
     edge.curved = 0.2)
```


```{r}
plot(graph_from_literal(a---b, b---c)) # the number of dashes doesn't matter
```


```{r}
plot(graph_from_literal(a--+b, b+--c))
```

```{r}
plot(graph_from_literal(a+-+b, b+-+c))
```


```{r}
g1 <- graph_from_literal(a-b-c-d-e-f, a-g-h-b, h-e:f:i, j)
plot(g1)
```


## edge, vertex, and network attributes  
access vertices and edges  
```{r}
E(g4)
```

```{r}
V(g4)
```

```{r}
g4[]
```

```{r}
g4[1, ]
```


Add attributes to the network, vertices, or edges  
```{r}
V(g4)$name
```


```{r}
V(g4)$gender <- c("male", "male", "male", "male", "female", "female", "male")
E(g4)$type <- "email"
E(g4)$weight <- 10
```


```{r}
edge_attr(g4)
```

```{r}
vertex_attr(g4)
```


```{r}
graph_attr(g4)
```


Another way to set attributes  
```{r}
g4 <- set_graph_attr(g4, name = "name", value = "Email Network")
g4 <- set_graph_attr(g4, name = "something", value = "a thing")

graph_attr(g4)
```


```{r}
graph_attr(g4, name = "name")
```


```{r}
g4 <- delete_graph_attr(graph = g4, name = "something")
```

```{r}
graph_attr(g4)
```


```{r}
plot(g4,
     edge.arrow.size = 0.5, 
     vertex.label.color = "black",
     vertex.label.dist =1.5,
     vertex.color = c("pink", "skyblue")[1 + (V(g4)$gender == "male")])
```


simplify graph to remove loops and multiple edges between the same nodes  
`edge.attr.comb` possible options: sum, mean, prod, min, max, first, last and ignore.  
`ignore`: the attributes be disregarded and dropped  
```{r}
g4s <- simplify(graph = g4, 
                remove.loops = TRUE,
                remove.multiple = TRUE,
                edge.attr.comb = c(type = "ignore", weight = "sum"))
```

```{r}
plot(g4s,
     vertex.label.dist = 1.5)
```

```{r}
g4s
```


## Specific graphs and graph models  
empty graph  
```{r}
eg <- make_empty_graph(40)
plot(eg, vertex.size = 10,
     vertex.label = NA)
```


Full graph  
```{r}
fg <- make_full_graph(n = 40)

plot(fg, 
     vertex.size = 10,
     vertex.label = NA)
```

Simple star graph  
```{r}
st <- make_star(40)

plot(st, 
     vertex.size = 10,
     vertex.label = NA)
```

Tree graph  
```{r}
tr <- make_tree(n = 40,
                children = 3,
                mode = "undirected")
plot(tr, 
     vertex.size = 10,
     vertex.label = NA)
```

Ring graph  
```{r}
rn <- make_ring(n = 40)

plot(rn, 
     vertex.size = 10,
     vertex.label = NA)
```


Erdos-Renyi random graph model  
```{r}
er <- sample_gnm(n = 100, m = 40)

plot(er,
     vertex.size = 6,
     vertex.label = NA)
```

Watts-Strogatz small-world model  
```{r}
sw <- sample_smallworld(dim = 2, 
                        size = 10,
                        nei = 1, 
                        p = 0.1)

plot(sw,
     vertex.size = 6,
     vertex.label = NA,
     layout = layout_in_circle)
```


Barabasi-Albert preferential attachment model for scale-free graphs  
```{r}
ba <- sample_pa(n = 100, 
                power = 1,
                m = 1,
                directed = F)

plot(ba, vertex.size = 6,
     vertex.label = NA)
```

```{r}
zach <- graph(edges = "Zachary")

plot(zach,
     vertex.size = 10,
     vertex.label = NA)
```

Rewiring a graph  
`each_edge()` is a rewiring method that changes the edge endpoints uniformaly randomly with a probability **prob**  
```{r}
rn.rewiring <- rewire(rn, each_edge(prob = 0.1))

plot(rn.rewiring, vertex.size = 10, vertex.label = NA)
```

Rewire to connect vertices to other vertices at a certain distance.  
```{r}
rn.neigh <- connect.neighborhood(graph = rn, order = 5)

plot(rn.neigh, vertex.size = 8, vertex.label = NA)
```


Combine graphs (disjoint union, assuming separate vertex sets): **%du%**  
```{r}
par(mfrow = c(1, 2))
plot(rn, vertex.size = 10, vertex.label = NA)
plot(tr, vertex.size = 10, vertex.label = NA)
```

```{r}
plot(rn %du% tr, vertex.size = 10, vertex.label = NA)
```


# Reading network data from files  
```{r}
nodes <- read.csv(file = "Dataset1-Media-Example-NODES.csv", header = TRUE, as.is = TRUE)
links <- read.csv(file = "Dataset1-Media-Example-EDGES.csv", header = TRUE, as.is = TRUE)
```

```{r}
head(nodes)
```

```{r}
head(links)
```


```{r}
nrow(nodes); unique(links[, c("from", "to")]) %>% nrow()
```

```{r}
?aggregate
links <- aggregate(x = links[, 3], by = links[, -3], FUN = sum)
links <- links[order(links$from, links$to), ]
```


```{r}
colnames(links)[4] <- "weight"
rownames(links) <- NULL
```














