---
title: "circlize"
author: "yincy"
date: "4/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(circlize)
```

# Chapter 1 Introduction  
Circular layout is very useful to represent complicated information. First, it elegantly represents information with long axes or a large amount of categories; second, it intuitively shows data with multiple tracks focusing on the same object; third, it easily demonstrates relations between elements. It provides an efficient way to arrange information on the circle and it is beautiful.  

Circos is a pioneer tool widely used for circular layout representations implemented in Perl. It greatly enhances the visualization of scientific results (especially in Genomics field). Thus, plots with circular layout are normally named as “circos plot”. Here the circlize package aims to implement Circos in R. One important advantage for the implementation in R is that R is an ideal environment which provides seamless connection between data analysis and data visualization. circlize is not a front-end wrapper to generate configuration files for Circos, while completely coded in R style by using R’s elegant statistical and graphic engine. We aim to keep the flexibility and configurability of Circos, but also make the package more straightforward to use and enhance it to support more types of graphics.  

In this book, chapters in Part I give detailed overviews of the general circlize functionalities. Part II introduces functions specifically designed for visualizing genomic datasets. Part III gives comprehensive guilds on visualizing relationships by Chord diagram.  

## Principle of design  
A circular layout is composed of sectors and tracks. For data in different categories, they are allocated into different sectors and for multiple measurements of the same category, they are represented as stacked tracks from outside of the circle to the inside. The intersection of a sector and a track is called a cell (or a grid, a panel), which is the basic unit in a circular layout. It is an imaginary plotting region for data points in a certain category.  

Since most of the figures are composed of simple graphics, such as points, lines, polygon, circlize implements low-level graphic functions for adding graphics in the circular plotting regions, so that more complicated graphics can be easily generated by different combinations of low-level graphic functions. This principle ensures the generality that types of high-level graphics are not restricted by the software itself and high-level packages focusing on specific interests can be built on it.  

Currently there are following low-level graphic functions that can be used for adding graphics. The usage is very similar to the functions without `circos`. prefix from the base graphic engine, except there are some enhancement specifically designed for circular visualization.  

+ `circos.points()`: adds points in a cell.  
+ `circos.lines()`: adds lines in a cell.  
+ `circos.segments()`: adds segments in a cell.  
+ `circos.rect()`: adds rectangles in a cell.  
+ `circos.polygon()`: adds polygons in a cell.  
+ `circos.text`: adds text in a cell.  
+ `circos.axis()` ands `circos.yaxis()`: add axis in a cell.  

Following functions arrange the circular layout  

+ `circos.initialize()`: allocates sectors on the circle.  
+ `circos.track()`: create plotting regions for cells in one single track.  
+ `circos.update()`: updates an existed cell.  
+ `circos.par()`: graphic parameters.  
+ `circos.clear()`: resets graphic parameters and internal variables.  

Thus, theoretically, you are able to draw most kinds of circular figures by the above functionalities. 

## A quick glance  





















