---
title: "01 geographic data models"
author: "YinCY"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# vector data model

The **vector data model** represents the world using points, lines and polygons. These have discrete, well-defined borders, meaning that vector datasets usually have a high level of precision.

Take care when using the word 'vector' as it can have two meanings in this book: geographic vector data and the `vector` class in R. The former is a data model, the latter is an R class just like `data.frame` and `matrix`. Still, there is a link between the two: the spatial coordinates which are at the heart of the geographic vector data model can be represented in R using `vector` objects.

The geographic vector data model is based on points located within a coordinate reference system (CRS). Points can represent self-standing features or they can be linked together to form more complex geometries such as lines and polygons. Most point geometries contain only two dimensions (3-dimensional CRSs contain an additional z value, typically representing height above sea level).

## Simple Features (sf)
Simple Features is a hierarchical data model that represents a wide ranges of geometry types. Of 17 geometry types supported by the specification, only 7 used in the vast majority of geographic research; these core geometry types are fully supported by the R package **sf**.

**sf** can represent all common vector geometry types (raster data classes are not supported by **sf**): points, lines, polygons and their respective 'multi' versions (which group together features of the same type into a single feature). **sf** also supports geometry collections, which can contain multiple geometry types in a single object. **sf** largely supersedes the **sp** ecosystem, which comprises **sp**, **rgdal** for data read/write and **rgeos** for spatial operations. 

```{r}
library(sf)
library(spData)
library(magrittr)

data("world")
world %>% class()
world %>% names()

plot(world)
```

```{r}
world[, "lifeExp"] %>% summary
```

```{r}
world[1:2, 1:3]
```

**Advantages of **sf** over **sp**:  

- **sf** objects can be treated as data frames in most operations.  
- **sf** functions can be combined using %>% operator and works well with the tidyverse collection of R packages.  
- **sf** function names are relatively consistent and intuitive, all begin with `st_`.  

```{r}
world_asia <- world[world$continent == "Asia", ]
asia <- st_union(world_asia)

plot(world["pop"], reset = F)
plot(asia, add = T, col = "red")
```

```{r}
plot(world["continent"], reset  =F)
cex <- sqrt(world$pop) / 10000
world_cents <- st_centroid(world, of_largest_polygon = T)
plot(st_geometry(world_cents), add = T, cex = cex)
```

```{r}
india <- world[world$name_long == "India", ]
plot(st_geometry(india), expandBB = c(0, 0.2, 0.6, 1), col = "gray", lwd = 3)
plot(world_asia[0], add = T)
```

The use of `world_asia[0]` to keep only the geometry column.  

### Geometry types
Geometries are the basic building blocks of simple features. Simple features in R can take on one of the 17 geometry types supported by the **sf** package. In this vignette we only focus on the seven most commonly used types: POINT, LINESTRING, POLYGON, MULTIPOINT, MULTILINESTRING, MULTIPOLYGON and GEOMETRYCOLLECTION. Find the whole list of possible feature types in the PostGIS manual.  

### Simple feature geometries (sfg)
The `sfg` class represents the different simpel feature geometry types in R: point, linestring, polygon (and their 'multi' equivalents, such as multipoints) or geometry collection.  


#### Functions to create simple feature geometry objects (sfg) from scratch
The names of these functions are simple and consistent, as they all start with the `st_` prefix and end with the name of the geometry type in lowercase letters.  

- A point: `st_point()`  
- A linestring: `st_linestring()`  
- A polygon: `st_polygon()`  
- A multipoint: `st_multipoint()`  
- A multilinestring: `st_multilinestring()`  
- A multipolygon: `st_multipolygon()`  
- A geometry collection: `st_geometrycollection()`  

`sfg` objects can be created from three base R data types:  

1. A numeric vector: a single point  
2. A matrix: a set of points, where each row represents a point, a multipoint or linestring  
3. A list: a collection of objects such as matrice, multilinestrings or geometry collections  

```{r}
# Point
st_point(c(5,2))
st_point(c(5, 2, 3))
st_point(c(5,2,1), dim = "XYM")
st_point(c(5, 2, 3, 1))

multipoint_matrix <- rbind(c(5, 2), c(1, 3), c(3, 4), c(3, 2))
st_multipoint(multipoint_matrix)
```

```{r}
# linestring
linestring_matrix <- rbind(c(1, 5), c(4, 4), c(4, 1), c(2, 2), c(3, 2))
st_linestring(linestring_matrix)

# polygon
polygon_list <- list(rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5)))
st_polygon(polygon_list)

# Polygon with a hole
polygon_border <- rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5))
polygon_hole <- rbind(c(2, 4), c(3, 4), c(3, 3), c(2, 3), c(2, 4))
polygon_with_hole_list <- list(polygon_border, polygon_hole)

st_polygon(polygon_with_hole_list) 

# Multilinestring
multilinestring_list <- list(rbind(c(1, 5), c(4, 4), c(4, 1), c(2, 2), c(3, 2)), 
                             rbind(c(1, 2), c(2, 4)))
st_multilinestring(multilinestring_list)

# multipolygon
multipolygon_list <- list(list(rbind(c(1, 5), c(2, 2), c(4, 1), c(4, 4), c(1, 5))), 
                          list(rbind(c(0, 2), c(1, 2), c(1, 3), c(0, 3), c(0, 2))))
st_multipolygon(multipolygon_list) 

# geometrycollection
geometrycollection_list <- list(st_multipoint(multipoint_matrix), 
                                st_linestring(linestring_matrix))

st_geometrycollection(geometrycollection_list) 
```

In most cases, an `sfc` object contains objects of the same geometry type. Therefore, when we convert `sfg` objects of type polygon into a simple feature geometry column, we would also end up with an `sfc` object of type polygon, which can be verified with `st_geometry_type()`.  

It is also possible to create an sfc object from sfg objects with different geometry types.  

```{r}
point1 = st_point(c(5, 2))

multilinestring_list1 = list(rbind(c(1, 5), c(4, 4), c(4, 1), c(2, 2), c(3, 2)), 
                             rbind(c(1, 2), c(2, 4)))
multistring1 <- st_multilinestring(multilinestring_list1)

st_sfc(point1, multistring1) %>% st_geometry_type()
```

```{r}
st_crs(point1)
```

All geometries in an `sfc` object must have the same CRS. 

```{r}
points_sfc_wgs <- st_sfc(point1, crs = 4326)
st_crs(points_sfc_wgs)
```

```{r}
st_sfc(point1, crs = "+proj=longlat +datum=WGS84 +no_defs")
```

### The sf class
```{r}
lnd_point <- st_point(c(0.1, 51.5))
lnd_geometry <- st_sfc(lnd_point, crs = 4326)
lnd_attrib <- data.frame(
  name = "London", 
  temperature = 25, 
  date = as.Date("2017-06-21")
)

lnd_sf <- st_sf(lnd_attrib, geometry = lnd_geometry)
lnd_sf %>% class()
```


# raster data model

The **raste data model** divides the surface up into cells of constant size. Raster datasets are the basis of background images used in web-mapping and have been a vital source of geographic data since the origins of aerial photography and satellite-based remote sensing devices. Rasters aggregate spatially specific features to a given resolution, meaning that they are consistent over space and scalable (many worldwide raster datasets are available).

```{r}
library(spDataLarge)
library(raster)
library(magrittr)

raster_filepath <- system.file("raster/srtm.tif", package = "spDataLarge")
new_raster <- raster(x = raster_filepath)
```

Dedicated functions for `RasterLayer` class

```{r}
new_raster %>% dim
new_raster %>% ncell()
new_raster %>% extent()
new_raster %>% crs()
new_raster %>% inMemory() # do not stored in memory
```

```{r}
new_raster %>% plot
```

### Raster classes
The `RasterLayer` class represents the simplest form of a raster object, and consists of only one layer.

```{r}
new_raster2 <- raster(nrows = 6, 
                      ncols = 6, 
                      res = 0.5, 
                      xmn = -1.5, 
                      xmx = 1.5, 
                      ymn = -1.5, 
                      ymx = 1.5, 
                      vals = 1:36)

# ?raster for more methods to create raster object
```


Aside from `RasterLayer`, there are two additional classes: `RasterBrick` and `RasterStack`. Both can handle multiple layers, but differ regarding the number of supported file formats, type of internal representation and processing speed.

A `RasterBrick` consists of multiple layers, which typically correspond to a single multispectral satellite file or a single multilayer object in memory. The `brick()` function creates a `RasterBrick` object. Usually, you provide it with a filename to a multilayer raster file but might also use another raster object and other spatial objects.  

```{r}
multi_raster_file <- system.file("raster/landsat.tif", 
                                 package = "spDataLarge")

r_brick <- brick(multi_raster_file)
```

```{r}
r_brick %>% nlayers()
```

A `RasterStack` is similar to a `RasterBrick` in the sense that it consists also of multiple layers. However, in contrast to `RasterBrick`, `RasterStack` allows you to connect several raster objects stored in different files or multiply objects in memory. More specifically, a `RasterStack` is a list of `RasterLayer` objects with the same extent and resolution. Hence, one way to create it is with the help of spatial objects already existing in R’s global environment.

```{r}
raster_on_disk <- raster(r_brick, layer = 1)
raster_in_memory <- raster(xmn = 301905, 
                           xmx = 335745, 
                           ymn = 4111245, 
                           ymx = 4154085, 
                           res = 30)
values(raster_in_memory) <- sample(seq_len(ncell(raster_in_memory)))
crs(raster_in_memory) <- crs(raster_on_disk)

r_stack <- stack(raster_in_memory, raster_on_disk)
r_stack
```

Another difference is that the processing time for `RasterBrick` objects is usually shorter than for `RasterStack` objects.

Decision on which `Raster*` class should be used depends mostly on a character of input data. Processing of a single mulitilayer file or object is the most effective with `RasterBrick`, while `RasterStack` allows calculations based on many files, many `Raster*` objects, or both.

Operations on `RasterBrick` and `RasterStack` objects will typically return a `RasterBrick`.

















