---
title: "rlist"
author: "yincy"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(rlist)
library(pipeR)
```

This package supports filtering, mapping, grouping, sorting, updating, searching and many other functions. It is pipe-friendly and strongly recommends functional programming style in list operations.  

# Quick overview  

```{r}
people <- list.load(file = "data/people.json")

people %>>% 
    list.filter(Expertise$R >= 1 & Expertise$Python >= 1) %>>% 
    list.class(Interests) %>>% 
    list.sort(-length(.)) %>>% 
    list.take(3) %>>% 
    list.map(. %>>% list.table(Age))
```

## Working in pipeline  
`PipeR` is recommended to work with rlist functions since they do not have any conflicting design with rlist.  


## File formats  
`rlist` is designed to deal with non-tabular data:  
 - `JSON`: `rlist` package imports `jsonlite` package to read/write JSON data.  
 - `YAML`: `rlist` import `yaml` package to read/wirte YAML data.  


# Features  
## mapping  
`list.load()` is designed for loading data from given data source. The data source can be either local or remote and the function by default uses the file extension to decide the way to read it.  

```{r}
people <- list.load(file = "data/sample.json")
str(people)
```

to extract the name of each people  
```{r}
lapply(people, function(x){
    x$Name
})
```

```{r}
list.map(people, Name)
```

List mapping is to evaluate an expression for each list member. It is the fundamental operation in `rlist` functionality. Almost all functions in this package that work with expressions are using mapping but in different ways.  

### list.map  
The simplest way of mapping is provided by `list.map()` as we have just demonstrated. Basically, it evaluates an expression for each list element.

The function makes it easier to query a list by putting all fields of the list member in mapping to the environment where the expression is evaluated. In other words, the expression is evaluated in the context of one list member each time.  
```{r}
list.map(people, Age)
```

Since the expression does not have to be a field name of the list member, we can evaluate whatever we want in the context of a list member.  

```{r}
list.map(people, sum(as.numeric(Expertise)))
```

```{r}
list.map(people, Expertise)
```

```{r}
list.map(people, list(age = Age, range = range(as.numeric(Expertise))))
```

In some cases we need to refer to the item itself, or its index in the list, or even its name. In the expression, `.` represents the item itself, `.i` represents its index, and `.name` represents its name.  

```{r}
nums <- c(a = 3, b = 2, c = 1)
list.map(nums, . + 1)
```

```{r}
list.map(nums, .i)
```

```{r}
list.map(nums, paste0("name: ", .name))
```

> `rlist` functions are general enough to work smoothly with vectors. list.map() works very much like lapply() so that the input will be finally transformed to list.  


### list.mapv  
if we want to get the mapping results as a vector rather than a list, we can use `list.mapv()`, which basically calls `unlist()` to the resulted from `list.map()`.  

```{r}
list.mapv(people, Age)
```

```{r}
list.mapv(people, sum(as.numeric(Expertise)))
```

### list.select  
In contrast to `list.map()`, `list.select()` provides an easier way to map each list member to a new list. This functions basically evaluates all given expressions and put the results into a list.   

If a field name a list member is selected, its name will automatically preserved. If a list item evaluated from other expression is selected, we may better give it a name, or otherwise it will only have an index.  

```{r}
list.select(people, Name, Age)
```

```{r}
list.select(people, Name, Age, nlang = length(Expertise))
```

### list.iter  
Sometimes we don't really need the result of a mapping but its side effects. For example, if we only need to print out something about each list member, we don't need to carry on the output of mapping.  

`list.iter()` performs iterations over a list and returns the input data invisibly for further data transformation.  

```{r}
list.iter(people, cat(Name, ":", Age, "\n"))
```


### list.maps  
All the previous functions work with a single list. However, there are scenarios where mapping multiple lists is needed. `list.maps()` evaluates an expression with multiple lists each of which is represented by a user-defined symbol at the function call.  

```{r}
l1 <- list(p1 = list(x = 1, y = 2), p2 = list(x = 3, y = 4), p3 = list(x = 1, y = 3))
l2 <- list(2, 3, 5)
list.maps(a$x*b+a$y, a = l1, b = l2)
```

`list.maps()` does not follow the conventions of many other functions like `list.map()` and `list.iter()` where the data comes first and expression comes the second. Since `list.maps()` supports multi-mapping with a group of lists, only implicit lambda expression is supported to avoid ambiguity. After that the function still allows users to define the symbol that represents each list being mapped in `...`.  

In the example above, `...` means `a = l1, b = l2`, so that a and b are meaningful in the first expression `a$x*b+a$y` where `a` and `b` mean the iterating element of each list, respectively.  

The function does not require named be supplied with the lists as arguments. In this case, we can use `..1`, `..2`, etc. to refer to the first, second or other lists.  

```{r}
list.maps(..1$x * ..2 + ..1$y, l1, l2)
```


## Filtering  
List filtering is to select list elements by given criteria. In `rlist` package, more than ten functions are related with list filtering. Basically, they all perform mapping first but then aggregate the results in different ways.  

### list.filter  
`list.filter()` filters a list by an expression that returns `TRUE` or `FALSE`. The results only contain the list elements for which the value of that expression turns out to be `TRUE`.  

Different from list mapping which evaluates an expression given each list element, list filtering evaluates an expression to decide whether to include the entire element in the results.  

```{r}
list.filter(people, Age >= 25) %>% str()
```

```{r}
library(pipeR)
#library(magrittr) # pipe from this package works as well

people %>>% 
    list.filter(Age >= 25) %>>% 
    list.mapv(Name)
```

```{r}
people %>% 
    list.filter(mean(as.numeric(Expertise)) >= 3) %>% 
    list.mapv(Name)
```

Meta-symbols like `.`, `.i` and `.name` can also be used. The following code will pick up the list element whose index is even.  

```{r}
people %>% 
    list.filter(.i %% 2 == 0) %>% 
    list.mapv(Name)
```

### list.find  
In some cases, we don't need to find all the instances given the criteria. Rather, we only need to find a few, sometimes only one. `list.find()` avoids searching across all list element but stops at a specific number of items found.  

```{r}
people %>% 
    list.find(Age >= 25, n = 1) %>% 
    list.mapv(Name)
```

### list.findi  
Similar with `list.find()`, `list.findi()` only returns the index of the elements found.  

```{r}
list.findi(people, Age >= 23, n = Inf)
```

### list.first, list.last 
`list.first()` and `list.last()` are used to find the first and last element that meets certain condition if specified, respectively.  

```{r}
list.first(people, Age >= 23) %>% str
```

```{r}
list.last(people, Age >= 23) %>% str
```

These two functions also works when the condition is missing. In this case, they simply take out the first/last element from the list or vector.  

```{r}
list.first(1:10)
```

```{r}
list.last(1:10)
```

### list.take  
`list.take()` takes at most a given number of elements from a list. **If the number is even larger than the length of the list, the function will by default return all elements in the list**.  

```{r}
list.take(1:10, n = 3)
```

```{r}
list.take(1:5, n = 8)
```

### list.skip  
As opposed to `list.take()`, `list.skip()` skips at most a given number of elements in the list and take all the rest as the results. If the number of elements to skip is equal or greater than the length of that list, an empty one will be returned.  

```{r}
list.skip(1:10, 3)
```

```{r}
list.skip(1:5, 8)
```

### list.takeWhile  
Similar to `list.take()`, `list.takeWhile()` is also designed to take out some elements from a list but subject to a condition. Basically, it keeps taking elements while a condition holds true.  

```{r}
people %>% 
    list.takeWhile(cond = Expertise$R >= 2) %>% 
    list.map(list(name = Name, R = Expertise$R))
```

### list.skipWhile  
`list.skipWhile()` keeps skipping elements while a condition holds true.  

```{r}
people %>% 
    list.skipWhile(cond = Expertise$R <= 2) %>% 
    list.map(list(name = Name, R = Expertise$R)) %>% 
    str
```

### list.is  
`list.is()` returns a logical vector that indicates whether a condition holds for each member of a list.  

```{r}
list.is(people, cond = "music" %in% Interests)
```

```{r}
list.is(people, "Java" %in% names(Expertise))
```

### list.which  
`list.which()` returns a integer vector of the indices of the elements of a list that meet a given condition.  

```{r}
list.which(people, "music" %in% Interests)
```

```{r}
list.which(people, "Java" %in% names(Expertise))
```

### list.all  
`list.all()` returns `TRUE` if **all** the elements of a list satisfy a given condition, or `FALSE` otherwise.  

```{r}
list.all(people, mean(as.numeric(Expertise)) > 3)
```

```{r}
list.all(people, "R" %in% names(Expertise))
```

### list.any  
`list.any()` returns `TRUE` if at least one of the elements of a list satisfies a given condition, or `FALSE` otherwise.  

```{r}
list.any(people, mean(as.numeric(Expertise)) > 3)
```

```{r}
list.any(people, "Python" %in% names(Expertise))
```

### list.count  
`list.count()` return a scalar integer that indicates the number of elements of a list that satisfy a given condition.  

```{r}
list.count(people, mean(as.numeric(Expertise)) >= 3)
```

```{r}
list.count(people, "R" %in% names(Expertise))
```

### list.match  
`list.match()` filters a list **by matching the names** of the list elements by a regular expression pattern.  

```{r}
data <- list(p1 = 1, p2 = 2, a1 = 3, a2 = 4)

list.match(data, pattern = "p[12]")
```

### list.remove  
`list.remove()` removes list elements by index or name.  

```{r}
list.remove(data, range = c("p1", "p2"))
```

```{r}
data %>% 
    list.which(cond = . < 3) %>% 
    list.remove(data, range = .)
```

### list.exclude  
`list.exclude()` removes list elements that satisfy given condition.  

```{r}
people %>% 
    list.exclude(cond = 'sports' %in% Interests) %>% 
    list.mapv(Name)
```

### list.clean  
`list.clean()` is used to clean a list by a function either recursively or not. The function can be built-in function like `is.null()` to remove all `NULL` values from the list, or can be user-defined function like `function(x) length(x) == 0` to remove all empty objects like `NULL`, `character(0L)`, etc.  

```{r}
x <- list(a = 1, b = NULL, c = list(x = 1, y = NULL, z = logical(0L), w = c(NA, 1)))
str(x)
```

to clean all `NULL` values in the list recursively  
```{r}
list.clean(x, fun = is.null, recursive = T) %>% str
```


to remove all empty values including `NULL` and zero-length vectors  

```{r}
list.clean(x, function(x) length(x) == 0L, recursive = TRUE) %>% 
    str
```

The function can also be related to missing values. For example, exclude all empty values and vectors with at least `NA`s.  

```{r}
list.clean(x, 
           fun = function(x) length(x) == 0L || anyNA(x), 
           recursive = T) %>% 
    str
```

### subset  
`subset()` is implemented for list object in a way that combines `list.filter()` and `list.map()`. This function basically filters a list while at the same time maps the qualified list elements by an expression.  

```{r}
people %>% 
    subset(Age >= 24, Name)
```

```{r}
people %>% 
    subset("reading" %in% Interests, sum(as.numeric(Expertise)))
```

## Updating  
`list.update()` partially modifies the given list by a number of lists resulted from expressions.  

```{r}
people %>% 
    list.select(Name, Age) %>% 
    list.stack()
```

`list.stack()` converts a list to a data frame with equivalent structure.  

Suppose we find that the age of each people is mistakenly recorded, say, 1 year less than their actual ages, respectively, we need to update the original data by refresh the age of each element.  

```{r}
people %>% 
    list.update(Age = Age + 1) %>% 
    list.select(Name, Age) %>% 
    list.stack()
```

`list.update()` can also be used to exclude certain fields of the elements. Once we update the fields we want to exclude to `NULL`, those fields are removed.  

```{r}
people %>% 
    list.update(Interests = NULL, Expertise = NULL, N = length(Expertise)) %>% 
    list.stack()
```

## Sorting  
### list.order  
`list.order()` evaluates the given lambda expressions and find out the order by default ascending. If the values for some members tie, the next values of the next expression, if any, will count.  

To get the order in descending, use `()` to enclose the expression or simply write a minus operator (`-`) before the expression if its value is numeric.  

```{r}
people %>% 
    list.order(Age)
```


```{r}
list.order(people, length(Interests))
```

```{r}
list.order(people, (Expertise$R))
```

```{r}
list.order(people, (length(Interests)), (Expertise$R))
```

### list.sort  
`list.sort()` produces a sorted list of the original list members. Its usage is exactly the same as `list.order()`.  

```{r}
people %>>% 
    list.sort(Age) %>>% 
    list.select(Name, Age) %>>% 
    str
```

```{r}
people %>% 
    list.sort(length(Interests)) %>% 
    list.select(Name, nint = length(Interests)) %>% 
    str
```

```{r}
people %>% 
    list.sort((Expertise$R)) %>% 
    list.select(Name, R = Expertise$R) %>% 
    str
```

```{r}
people %>% 
    list.sort((length(Interests)), (Expertise$R)) %>% 
    list.select(Name, nint = length(Interests), R = Expertise$R) %>% 
    str
```

```{r}
people %>% 
    list.sort(max(unlist(Expertise))) %>% 
    list.mapv(Name)
```


## Grouping  
rlist supports multiple types of grouping.  

### list.group  
`list.group()` is used to put list elements into subgroups by evaluating an expression. The expression often produces a scalar value such as a logical value, a character value, or a number. Each group denotes a unique value that expression takes for at least one list element, and all elements are put into one and only one group.  

Divide numbers from 1 to 10 into even and odd numbers.  

```{r}
list.group(1:10, . %% 2 == 0)
```

```{r}
list.group(people, Age) %>% 
    str
```

```{r}
people %>>% 
    list.group(Age) %>>% 
    list.map(. %>>% list.mapv(Name))
```

The mapping runs at the first-level, that is, for each group. The mapper expression `. %>>% list.mapv(Name)` means that each people in the group maps to the name.  


### list.ungroup  
`list.group()` produces a nested list in which the first level are groups and the second level are the original list elements put into different groups.

`list.ungroup()` reverts this process. In other words, the function eradicates the group level of a list.  

```{r}
ageGroup <- list.group(people, Age)
str(ageGroup)
```

```{r}
list.ungroup(ageGroup)
```

### list.cases  







