---
title: "dplyr"
author: "yincy"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to dplyr  
## Single table verbs  
dplyr aims to provide a function for each basic verb of data manipulation. These verbs can be organised into three categories based on the component of the dataset that they work with.  

Rows:  
- `filter()` choose rows based on column values.  
- `slice()` choose rows based on location.  
- `arrange()` changes the order of the rows.  

Columns:
- `select()` changes whether or not a column is included.  
- `rename()` changes the name of columns.  
- `mutate()` changes the values of columns and creates new columns.  
- `relocate()` changes the order of the columns. 

Groups of rows:  
- `summarise()` collapses a group into a single row.  


# From base R to dplyr  
## Overview  
1. The code dpltr verbs input and output data frames. This contrasts with base R functions which more frequenctly work with individual vectors.  
2. dplyr relies heavily on "non-standard evaluation" so that you don't need to use $ to refer to culumns in the "current" data frame. This behaviour is inspired by the base functions `subset()` and `transform()`.  
3. dplyr solutions tend to use a variety of single purpose verbs, while base R solutions typically tend to use [ in a variety of ways, depending on the task at hand.  
4. Multiple dplyr verbs are often strung together into a pipeline by %>% . In base R, you'll typically save intermediate results to a variable that you either discard, or repeatedly overwrite.  
5. All dplyr verbs handle "grouped" data frames so that the code to perform a computation per-group looks very similar to code that works on a whole data frame. In base R, per-group operations tend to have varied forms.  


## One table verbs  
The following table shows a condensed translation between dplyr verbs and their base R equivalents.  

**dplyr**| **base**  
---------|------------
`arrange(df,x)`| `df[order),,drop = F]` 
`distinct(df, x)`| `df[!duplicated(x), , drop = F]`, `unique()`
`filter(df, x)`| `df[which(x), , drop = F]`, `subset()`
`mutate(df, z = x + y)`| `df$z = df$x + df$y`, `transform()`
`pull(df, 1)`| `df[[1]]`
`pull(df, x)`| `df$x`
`rename(df, y = x)`| `names(df)[names(df) == "x] <- "y"`
`relocate(df, y)`| `df[union("y", names(df))]`
`select(df, x, y)`| `df[c("x", "y")], subset()`
`select(df, starts_with("x"))`| `df[grepl(names(df), "^x")]`
`summarise(df, mean(x))`| `mean(df$x)`, `tapply()`, `aggregate()`, `by()`
`slice(df, c(1, 2, 5))`| `df[c(1, 2, 5), , drop = F]`


```{r}
library(dplyr)

mtcars <- as_tibble(mtcars)
iris <- as_tibble(iris)
```



### `arrange()`: Arrange rows by variables  
`dplyr::arrange()` orders the rows of a data frame by the values of one or more columns.  

```{r}
mtcars %>% 
    arrange(cyl, disp)
```

The `desc()` helper allows you to order selected variables in descending order.  

```{r}
mtcars %>% 
    arrange(desc(cyl), desc(disp))
```

```{r}
mtcars %>% 
    arrange(-cyl, -disp)
```


base R  
```{r}
mtcars[order(mtcars$cyl, mtcars$disp, decreasing = T), , drop = F]
```


### `distinct`: Select distinct/unique rows  

```{r}
df <- tibble(
    x = sample(10, 100, replace = T), 
    y = sample(10, 100, replace = T)
)

df %>% distinct(x)
```


```{r}
df %>% distinct(x, .keep_all = T)
```


### `filter()`: Returns rows with matching conditions  
`dplyr::filter()` select rows where an expression is `TRUE`.  

```{r}
starwars %>% filter(species == "Human")
```


```{r}
starwars %>% filter(mass > 1000)
```

```{r}
starwars %>% filter(hair_color == "none" & eye_color == "black")
```


### `mutate()`: Create or transform variables  
`dplyr::mutate()` creates new variables from existing variables  

```{r}
df %>% mutate(z = x + y, z2 = z ^2)
```


```{r}
gf <- tibble(g = c(1, 1, 2, 2), x = c(0.5, 1.5, 2.5, 3.5))

gf %>% 
    group_by(g) %>% 
    mutate(x_mean = mean(x), x_rank = rank(x))
```


### `pull()`: Pull out a single variable  
`dplyr::pull()` extracts a variable either by name or position.  

```{r}
mtcars %>% pull(1)

mtcars %>% pull(cyl)
```


### `relocate()`: Change column order  
`dplyr::relocate()` makes it easy to remove a set of columns to a new position.  

```{r}
# to front 
mtcars %>% relocate(gear, carb)
```


```{r}
# to back
mtcars %>% relocate(mpg, cyl, .after = last_col())
```


### `rename()`: Rename variables by name  
`dplyr::rename()` allows you to rename variables by name or position.  

```{r}
iris %>% rename(sepal_length = Sepal.Length, sepal_width = 2)
```


### `rename_with()`: Rename variables with a function  
`dplyr::rename_with()` transform column names with a function.  

```{r}
iris %>% rename_with(toupper)
```


### `select()`: Select variables by name  
`dplyr::select()` subsets columns by position, name, function of name, or other property.  

```{r}
iris %>% select(1:3)
```

```{r}
iris %>% select(Species, Sepal.Length)
```

```{r}
iris %>% select(starts_with("Petal"))
```

```{r}
iris %>% select(where(is.factor))
```


### `summarise()`: Reduce multiple values down to a single value  
`dplyr::summarise()` compuates one or more summaries for each group.  

```{r}
mtcars %>% 
    group_by(cyl) %>% 
    summarise(
        mean = mean(disp), 
        n = n()
    )
```


### `slice()`: Choose rows by position  
`slice` select rows with their location.  

```{r}
slice(mtcars, 25:n())

mtcars %>% slice(25:n())
```


## Two-table verbs  

**dplyr**| **base**
---------|------------------
`inner_join(df1, df2)`| `merge(df1, df2)`
`left_join(df1, df2)`| `merge(df1, df2, all.x = T)`
`right_join(df1, df2)`| `merge(df1, df2, all.y = T)`
`full_join(df1, df2)`| `merge(df1, df2, all = T)`
`semi_join(df1, df2)`| `df1[df1$x %in% df2$x, , drop = F]`
`anti_join(df1. df2)`| `df1[!df$x %in% df2$x, , drop = F]`


### Mutating joins  
dplyr's `inner_join()`, `left_join()`, `right_join()` and `full_join()` add new columns from y to x, matching rows based on a set of "keys", and after only in how missing matches are handled. They are equivalent to calls to `merge()` with various settings of the `all`, `all.x` and `all.y` arguments. The main difference is the order of the rows.  

- `dplyr` preserves the order of the x data frame.  
- `merge` sorts the key columns. 


### Filtering joins  
dplyr's `semi_join()` and `anti_join()` affect only the rows, not the columns.  

```{r}
band_members %>% semi_join(band_instruments)
```

```{r}
band_members %>% anti_join(band_instruments)
```


# Row-wise operations  
dplyr, and R in general, are particularly well suited to performing operations over columns, and performing operations over rows is much harder. In this vignette, you will learn dplyr's approach centered around the row-wise data frame created by `rowwise()`.  

There are three common use cases that we discuss in this viggnette:  
- Row-wise aggregates (e.g. compute the mean of x, y, z).  
- Callign a function multiple times with varying arguments.  
- Working with list-columns.  

## Creating  
Row-wise operations require a special type of grouping where each group consist of a single row. You can create this with `rowwise()`.  

```{r}
df <- tibble(x =1:2, y = 3:4, z = 5:6)

df %>% rowwise()
```

Like `group_by()`, `rowwise()` doesn't really do anything itself, it just changes how the other verbs work.  

```{r}
df %>% mutate(m = mean(c(x, y, z)))
```

```{r}
df %>% rowwise() %>% mutate(m = mean(c(x, y, z)))
```

If you use `mutate()` with a data frame, it computes the mean of x, y and z across all rows. If you apply it to a row-wise data frame, it computes the mean for each row.  

You can optionally supply "identifier" variables in your call to `rowwise()`. These variables are preserved when you call `summarise()`, so they behave somewhat similarly to the grouping variable passed to `group_by()`.  

```{r}
df <- tibble(name = c("Mara", "Hadley"), x = 1:2, y = 3:4, z = 5:6)

df %>% 
    rowwise() %>% 
    summarise(m = mean(c(x, y, z)))
```

```{r}
df %>% 
    rowwise(name) %>% 
    summarise(m = mean(c(x, y, z)))
```

`rowwise()` is just a special form of grouping, so if you want to remove it from a data frame, just call `ungroup()`.  


## Per row summary statistics  
`dplyr::summarise()` makes it really easy to summarise values across rows within one column. When combined with `rowwise()` it also makes it easy to summarise values across columns within one row.  
```{r}
df <- tibble(
    id = 1:6, 
    w = 10:15, 
    x = 20:25, 
    y = 30:35, 
    z = 40:45
)

rf <- df %>% rowwise()

rf %>% mutate(total = sum(c(w, x, y, z)))
```

```{r}
rf %>% summarise(total = sum(c(w, x, y, z)))
```

If you have a lot of variables, it's going to be tedious to type every variable name. instead, you can use `c_across()` which uses tidy selection syntax so you can to succinctly select many variables.  

```{r}
rf %>% mutate(totle = sum(c_across(w:z)))
```

```{r}
rf %>% mutate(total = sum(c_across(where(is.numeric))))
```

You could combine this with column-wise operations to compute the proportion of the total for each column.  

```{r}
rf %>% 
    mutate(total = sum(c_across(w:z))) %>% 
    ungroup() %>% 
    mutate(prop = across(w:z, ~. / total))
```


## Row-wise summary functions  
The `rowwise()` approach will work for any asummary funciton. But if you need greater speed, it's worth looking for a built-in row-wise variant of your summary function. These are more efficient because they operate on the data frame as whole; they don't split it into rows, compute the summary, and then join the result back together agin.  

```{r}
df %>% 
    mutate(total = rowSums(across(where(is.numeric))))
```

```{r}
df %>% mutate(mean = rowMeans(across(where(is.numeric))))
```

## List-columns  
`rowwise()` operations are a natural pairing when you have list-columns. They allow you to avoid explicit loops and/or functions from the `apply()` or `purrr::map()` families.  

### Motivation  
Imagine you have this data frame, and you want to count the lengths of each element.  

```{r}
df <- tibble(
    x = list(1, 2:3, 4:6)
)
```

```{r}
df %>% mutate(l = length(x))

df %>% mutate(l = lengths(x))

df %>% mutate(l = sapply(x, length))

df %>% mutate(l = purrr::map_int(x, length))
```

```{r}
df %>% 
    rowwise() %>% 
    mutate(l = length(x))
```

### Subsetting  
Before we continue on, I wanted to brieflt mention the magic that makes this work. This isn't something you'll generally need to think about, but it's useful to know about when something goes wrong.  

There's an important difference between a grouped data frame where each group happens to have one row, and a row-wise data frame where every group always has one row. Take these two data frames.  

```{r}
df <- tibble(g = 1:2, y = list(1:3, "a"))
gf <- df %>% group_by(g)
rf <- df %>% rowwise(g)
```

```{r}
gf %>% mutate(type = typeof(y), length = length(y))
```

```{r}
rf %>% mutate(type = typeof(y), length = length(y))
```

They key difference is that when `muate()` slices up the columns to pass to `length(y)` the grouped mutate uses `[` and the row-wise muate use `[[`. The following code gives a flaour of the difference if you used a for loop.  

```{r}
# groupped
out1 <- integer(2)

for(i in 1:2){
    out1[[1]] <- length(df$y[i])
}

out1
```


```{r}
#row-wise
out2 <- integer(2)

for(i in 1:2){
    out2[[i]] <- length(df$y[[i]])
}

out2
```

### Modelling  
`rowwise()` data frames allows you to solve a variety of modeling problems in what I think is a particularly elegent way.  

```{r}
by_cyl <- mtcars %>% nest_by(cyl)
by_cyl
```

This is a little different to the usual `group_by()` output: we have visibly changed the structure of the data. Now we have three rows, and we have a list-col, `data`, that stores the data for that group. Also note that the output is `rowwise()`; this is important because it's going to make working with that list of data frames much easier.  

One we have one data frame per row, it's straightforward to make one model per row.  
```{r}
mods <- by_cyl %>% mutate(mod = list(lm(mpg ~ wt, data = data)))
mods
```

And supplement that with one set of predictions per row.  
```{r}
mods <- mods %>% mutate(pred = list(predict(mod, data)))
mods
```

You could then summarize the model in a variety of ways. 

```{r}
mods %>% summarise(rmse = sqrt(mean((pred - data$mpg)^2)))
```

```{r}
mods %>% summarise(rsq = summary(mod)$r.squared)
```

```{r}
mods %>% summarise(broom::glance(mod))
```

access the parameters of each model  
```{r}
mods %>% summarise(broom::tidy(mod))
```

## Repeated function calls  
`rowwise()` doesn't just work with functions that return a length-1 vector (aka summary functions); it can work with any function if the result is a list. This means that `rowwise()` and `mutate()` provide an elegant way to call a function many times with varying arguments, storing the outputs alongside the inputs.  

## Simulations  



































