---
title: "tidyselect"
author: "yincy"
date: "5/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyselect)
library(magrittr)
```

`tidyselect` implements a specialised sublanguage of R for selecting variables from data frames and other data structures.  

## Selections as dots or as named arguments  
There are two major ways of designing a function that takes selections.  

+ Passing **dots** as in `dplyr::select`.  
```{r}
mtcars %>% dplyr::select(mpg, cyl)
```

+ Interpolating **named arguments** as in `tidyr::pivot_longer()`. In this case, multiple inputs can be provided inside `c()` or by using boolean operators.  

```{r}
mtcars %>% tidyr::pivot_longer(c(mpg, cyl))
mtcars %>% tidyr::pivot_longer(mpg | cyl)
```

Our general recommendation is to take dots when the main purpose of the function is to create a new data structure based on a selection. When the selection is accessory to the main purpose of the function, take it as a named argument. In doubt, we recommend using named arguments because it is easier to change a named argument to dots than the other way around.  

## Do you need tidyselect?  
The tools described in this vignette are rather low level. Depending on your use case, it may be easier to wrap `dplyr::select()`. You’ll get a data frame containing the columns selected by your user, which you can then handle in various ways.  

The following examples illustrate how you could write a function that takes a selection of data and returns the corresponding data frame with capitalised names.  
```{r}
# pasing dots  
toupper_dots <- function(data, ...){
  sel <- dplyr::select(data, ...)
  rlang::set_names(sel, toupper)
}

# Interpolating a named argument with {{ }}
toupper_arg <- function(data, arg){
  sel <- dplyr::select(data, {{arg}})
  rlang::set_names(sel, toupper)
}


mtcars %>% toupper_dots(mpg:disp, vs)
mtcars %>% toupper_arg(c(mpg:disp, vs))
```

The main advantage of the lower level tidyselect tools is that they offer a bit more information and flexibility. Instead of returning the selected data, they return the locations of selected elements inside the input data. If you don’t need the selected locations and can afford the dependency, you may consider wrapping dplyr instead.  


## The selection evaluators  
`tidyselect` is implemented with non-standard evaluation (NSE). This unique feature of the R language refers to the ability of functions to **defuse** some or all of their arguments, and resume evaluation later on. Crucially, evaluation can be resumed in a different context or according to different rules, which is often how domain-specific languages are created in R.  

### Deufsing and resuming evaluation of R code  
When a function argument is defused, R halts the evaluation of the code and returns a defused expression instead. This expression contains the code that describes how compute the intended value.  

defuse your own R code with `expr()`  
```{r}
own <- rlang::expr(1 + 2)
own
```




























