---
title: "Lab-4"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

```{r, message=FALSE, echo=FALSE}
library(mosaic)
library(magrittr)
```


# Question 1
```{r}
pnorm(145, mean = 127.1, sd = 11.7, lower.tail = F)
```

# Question 2
```{r}
z <- (145 - 127.1) / 11.7
z
pnorm(z, lower.tail = F)
```

# Question 3
```{r}
pt(2.3, df = 35, lower.tail = F)
```

```{r}
college <- read.table(file = "CollegeMidwest.txt", header = T)
head(college)

set.seed(90095)

sample_index <- sample(2919, size = 40)
college_sample <- college[sample_index, ]

x_bar <- mean(college_sample$CumGpa)
x_bar

s <- sd(college_sample$CumGpa)
t_stat <- (x_bar - 3.5) / (s / sqrt(40))
t_stat

2 * pt(t_stat, df = 40 - 1)

t.test(college_sample$CumGpa, mu = 3.5)
```


# Question 4
```{r}
t.test(college_sample$CumGpa, 
       mu = 3.5,
       conf.level = 0.99)
```

# Question 5
With the p-value > 0.01, we don't have sufficient evidence to suggest that the mean cumulative GPA of the students at the College of the Midwest is different from 3.5.  


# Question 6
```{r}
ex_vec <- c(4,1,2,6,8,5,3,7)

#(a)
ex_vec > 4 & ex_vec < 6

#(b)
ex_vec < 4 | ex_vec > 6
```

# Question 7
```{r}
cdc <- read.csv("cdc.csv", header = T)
head(cdc)

#(a)
hlth_notgood <- cdc$genhlth == "poor" | cdc$genhlth == "fair"

tally(~ hlth_notgood, 
      format = "proportion")

#(b)
hlth_good <- cdc$exerany == "1" & 
    (cdc$genhlth == "good" | cdc$genhlth == "very good" | cdc$genhlth == "excellent")

tally(~ hlth_good, format = "proportion")

#(c)
hlth_noex <- cdc$exerany == "0" & 
    (cdc$genhlth == "good" | cdc$genhlth == "very good" | cdc$genhlth == "excellent")

tally(~ hlth_noex, format = "proportion")
```

# Question 8
The statistic is of the difference between health and exercise at least once a week and health but do not exercise.  

# Question 9
```{r}
diff_props <- numeric(1000)
set.seed(147)

for(i in 1:1000){
  exerany_shuffle_i <- sample(cdc$exerany)
  cond_props_i <- tally(hlth_notgood ~ exerany_shuffle_i, format = "proportion")
  diff_props[i] <- cond_props_i[2,2] - cond_props_i[2, 1]
}

obs_table <- tally(hlth_notgood ~ exerany, data = cdc, format = "proportion")
obs_diff <- obs_table[2, 2] - obs_table[2, 1]

t.test(diff_props, mu = obs_diff, alternative = 'two.side')
```

`mean(diff_props) > 0.01` compute the proportion of diff_props bigger then 0.01.  

# Question 10
As the p-value less than 0.05, we reject the null hypothesis.  

# Question 11
```{r}
#(a)
x <- mean(diff_props)
s <- sd(diff_props)
t <- (x - obs_diff) / (s / sqrt(1000))
t

#(b)
mu <- mean(diff_props)
de <- sqrt((obs_diff * (1 - obs_diff))/ 1000)
z <- (mu - obs_diff) / (de / sqrt(1000))
z
```

# Question 12
```{r}
2 * pnorm(z)
```

The theory-based p-value is smaller than the simulation-based p-value.  
