---
title: "OOP_Advanced_R"
author: "yincy"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Chapter 7 OOfield guide  
R has three object oriented systems and the base type.  

Central to any object-oriented system are the concepts of **class** and **method**.  

A **class** defines the behaviour of **objects** by describing their attributes and their relationship to other classes. The class is also used when selecting **methods**, functions that behave differently depending on the class of their input.  

Classes are usually organised in a hierarchy: if a method does not exist for a child, then the parent's method is used instead; the child **inherits** behavious from the parent.  

- **S3** implements a style of OO programming called generic-function OO.  
- **S4** works similarly to S3, but is more formal. There are two major differences to S3. S4 has formal class definitions, which describe the representation and inheritance for each class, and has special helper functions for defining generics for each class, and has special helper functions for defining generics and methods. S4 also has multiple dispatch, which means that generic functions can pick methods based on the class of any number of arguments, not just one.  

- **Reference classes**, called RC for short, are quite different from S3 and S4. RC implements message-passing OO, so methods belong to classes, not functions.  

- **base types**, the internal C-level types that underlie the other OO systems. Base types are mostly manipulated using C code, but they're important to known about because they provide the building blocks for the other OO systems.  


## Base types  
Underlying every R object is a C structure (or struct) that describes how that object is stored in memory. The struct includes the contents of the object, the information needed for memory management, and, most importantly a **type**. This is the **base type** of an R object.  

Base types are not really an object system because only the R core team can create new types.  

determine an object's base type with `typeof()`.  

Unfortunately the names of base types are not used consistently throughout R, and type and the corresponding 'is' function may use different names.  

```{r}
library(pryr)

f <- function(){}
typeof(f)
is.function(f)
```

```{r}
typeof(sum)
is.primitive(sum)
```

to see if an object is an pure base type. -- doesn't have S3, S4 or RC behaviour   
```{r}
is.object(f)
```

## S3  
```{r}
library(pryr)
```

```{r}
df <- data.frame(x = 1:10, y = factor(letters[1:10]))
otype(df) # data frame is an S3 class
otype(df$x) # vector isn't an S3 class
otype(df$y) # factor is an S3 class
```

In S3, methods belong to functions, called **generic functions**, or generic for short.  
S3 methods do not belong to objects or classes. This is different from most other programming languages, but is a legitimate OO style.  


`pryr` also provides **ftype()** which describes the object system associated with a function  
```{r}
mean
ftype(mean)
```

You can see all the methods that belong to a generic with `methods()`  
```{r}
methods("mean")
```


### Defining classes and creating objects  
S3 is a simple and ad hoc system; it has no formal definition of a class. To make an object an instance of a class, you just take an existing base object and set the class attribute.  

You can do that during creating with `structure()`, or after the fact with class() <- ""  
```{r}
foo <- structure(.Data = list(), class = "foo")
otype(foo)
```

or  
```{r}
foo <- list()
class(foo) <- "foo"
otype(foo)
```


**S3 objects are usually built on top of lists, or atomic vectors with attributtes**. 

```{r}
mod <- lm(log(mpg) ~ log(disp), data = mtcars)
class(mod)
print(mod)
```

```{r}
class(mod) <- "data.frame"
class(mod)
print(mod)
class(mod) <- "lm"
print(mod)
```

You can determine the class of any object using `class(x)`, and see if an object inherits from a specific class using `inherits(x, "classname")`  
```{r}
class(foo)
inherits(foo, "foo")
```

The class of an S3 object can be a vector, which describes behaviour from most to least specific.  

Apart from developer supplied constructor functions, S3 has no checks for correctness.  





## Creating new methods and generics  
To add a new generic, create a function that calls `UseMethod()`.  

`UseMethod()` takes two arguments: the name of the generic function, and the argument to use for method dispatch.  

If you omit the second argument it will dispatch the first argument to the function.  

There is no need to pass any of the arguments of the generic to `UseMethod()` and you shouldn't do so. `UseMethod()` uses black magic to find them out for itself.  


**A generic isn't useful without some methods**.  

```{r}
f <- function(x)UseMethod(generic = "f")
f.a <- function(x)"Class a" # class a's method
a <- structure(list(), class = "a")
class(a)
f(a)
```

Adding a method to an existing generic works in the same way  
```{r}
mean.a <- function(x)"class a method"
mean(a)
```


## Method dispatch  
`UseMethod()` creates a vector of function names, like paste0("generic", ".", c(class(x), "defualt"))  

The defualt class makes it possible to set up a fall back method for otherwise unknown classes.  

```{r}
f <- function(x) UseMethod(generic = "f")
f.a <- function(x) "class a method"
f.default <- function(x) "unknown class"
```

```{r}
f(structure(list(), class = "a"))
```

no method for b class, so use method for a class  
```{r}
f(structure(list(), class = c("b", "a")))
```

no method for c class, so falls back to default  
```{r}
f(structure(list(), class = "c"))
```


Group generic mehtods add a little more complexity.  

Group generics makes it possible to implement methods for multiple generics with one function.  

The four group generics and the functions they include are:  
1. **Math**: abs, sign, sqrt, floor, cos, sin, log, exp, ...  
2. **Ops**: +, -, *, /, ^, %%, %/%, &, |, !, ==, !=, <, <=, >, >=  
3. **Summary**: all, any, sum, prod, min, max, range  
4. **Complex**: Arg, Conj, Im, Mod, Re  

find more information `?groupGeneric`  

The most important thing to take away from this is to recognise that **Math**, **Ops**, **Summary**, and **Complex** aren't real functions, but instead represent groups of functions.  

Note that inside a group generic function a special variable `.Generic` provides the actual generic function called.  

methods are normal R functions, you can call them directly  
```{r}
c <- structure(list(), class = "c")

# call the correct method
f.default(c)


# force R to call the wrong method  
f.a(c)
```



## S4  
S4 works in a similar way to S3, but it adds formality and rigour. Methods still belong to functions, not classes, but:  
1. Classes have formal definitions which describe their fields and inheritance structures (parent classes).  
2. Method dispatch can be based on multiple arguments to a generic function, not just one.  
3. There is a special operator, @, for extracting slots (aka fields) from an S4 object.  

All S4 related code is stored in the `methods` package.  


## Recognising objects, generic functions, and methods  

## identify S4 object  
1. `str()` describe it as a 'formal' class  
2. `isS4()` returns TRUE  
3. `pryr::otype()` reurn 'S4'  

## identify S4 generics and methods  
1. `pryr::ftype()`  

```{r}
library(pryr)
library(stats4)
```

```{r}
y <- c(26, 17, 13, 12, 20, 5, 9, 8, 5, 4, 8)
nLL <- function(lambda) - sum(dpois(y, lambda, log = TRUE))
fit <- mle(nLL, start = list(lambda = 5), nobs = length(y))
```

```{r}
isS4(fit)
```

```{r}
otype(fit)
```

```{r}
ftype(nobs)
```


```{r}
mle_nobs <- method_from_call(nobs(fit))
isS4(mle_nobs)
```

```{r}
ftype(mle_nobs)
```

Use `is()` with one argument to list all classes that an object inherits from.  
Use `is()` with two arguments to test if an object inherits from a specific class.  

```{r}
is(fit)
```

```{r}
is(fit, "mle")
```

get a list of all S4 generics with `getGenerics()`.  

get a list of all S4 classes with `getClasses()`.  

list all S4 methods with `showMethods()`, optionally restricting selection either by `generic` or by `class` (or both).  



### Defining classes and creating objects  
An S4 class has three key properties:  
1. A **name**: an alpha-numeric class identifer. By convention, S4 class names use UpperCamelCase.  
2. A named list of **slots** (fields), which defines slot names and permitted classes. For example, a person class might be represented by a character name and a numeric age: lsit(name = "character", age = "numeric").  

3. A string giving the class it inherits from, or, in S4 terminology, that it **contains**. You can provide multiple classes for multiple inheritance, but this is an advanced technique which adds much complexity.  

In slots and contains you can use S4 classes, S3 classes registered with `setOldClass()`, or the implicit class of a base type. In slots you can also use the special class **ANY** which does not restrict the input.  


```{r}
setClass(Class = "Person",
         slots = list(name = "character", age = "numeric"))

setClass(Class = "Employee", 
         slots = list(boss = "Person"),
         contains = "Person")

alice <- new(Class = "Person", name = "Alice", age = 40)
john <- new(Class = "Employee", name = "John", age = 20, boss = alice)
```

To access slots of an S4 objects use @ or `slot()`  
```{r}
alice@name
slot(john, name = "boss")
```


**@ is equivalent to $, and slot() to [[**.  

```{r}
slotNames(john)
slot(object = john, name = "boss")
john
```


If an S4 object contains (inherits) from an S3 class or a base type, it will have a special `.Data` slot which contains the underlying base type or S3 object.  

```{r}
setClass(Class = "RangedNumeric",
         contains = "numeric",
         slots = list(min = "numeric", max = "numeric"))
rn <- new(Class = "RangedNumeric", 1:10, min = 1, max = 10)
slotNames(rn)
```


### Creating new methods and generics  
S4 provides special functions for creating new generics and methods.  

`setGeneric()` creates a new generic or converts an existing function into a generic.  

`setMethod()` takes the name of the generic, the classes the method shoud be associated with, and a function that implements the method.  

```{r}
setGeneric(name = "union")

setMethod(f = "union",
          c(x = "data.frame", y = "data.frame"),
          function(x, y){
              unique(rbind(x, y))
          })
```


If you create a new generic from scratch, you need to supply a function that calls `standardGeneric()`:  
```{r}
setGeneric(name = "myGeneric",
           def = function(x){
               standardGeneric(f = "myGeneric")
           })
```

`standardGeneric()` is the equivalent to `UseMethod()`.  



### Method dispatch  
```{r}
selectMethod(f = "nobs", 
             signature = list("mle"))
```






























































































