---
title: "2020-11-1"
author: "yincy"
date: "11/1/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, comment = "")
```

```{r, message=FALSE, warning=FALSE}
library(mosaic)
```


```{r load-data}
NCbirths <- read.csv(file = "births.csv")
```

### Question 1  
Answer.  
Variable `Habit` indicates that the mother is smoked or not.  
Variable `weight` indicates the weight of the baby.  

### Question 2  
Answer.  
Variable `Feduc`, an numerical variable, might affect the weight of the baby weight, because well educated women known more how scientifically take care of herself during pregnancy.   

Variable `Premie`, an categorical variable, will affect the weight of the baby, because the earlier birth of the baby the lower weight he/she will be.  

### Question 3  
Answer.  
```{r}
df <- subset(NCbirths[, c("Habit", "weight", "Feduc", "Premie")], subset = NCbirths$Habit == "NonSmoker")
```

### Question 4  
```{r}
head(df)
```


### Question 5  
Answer.  
(a)  
```{r}
mean(~ weight, data = NCbirths)
```

(b)  
```{r}
mean(~ weight, data = NCbirths[NCbirths$Habit == "Smoker", ])
```

(c)  
```{r}
mean(~ weight, data = NCbirths[NCbirths$Habit == "NonSmoker", ])
```

### Question 6  
Answer.  
(a)  
```{r}
sd(~ weight, data = NCbirths)
```

(b)  
```{r}
sd(~ weight, data = NCbirths[NCbirths$Habit == "Smoker", ])
```

(c)  
```{r}
sd(~ weight, data = NCbirths[NCbirths$Habit == "NonSmoker", ])
```


```{r}
Births <- NCbirths[NCbirths$Habit != '', ]
```


### Question 7  
Answer.  
```{r}
histogram(~ weight | Habit, data = Births, 
          layout = c(1, 2), 
          width = 10, 
          density = T, 
          fit = "normal", 
          v = mean(weight ~ Habit, data = Births), 
          stripes = "vertical", 
          gcol = "red")
```

It seems that the baby's weight of Non-smoker mother is a little more weight than smoker mother's.  


### Question 8  
Answer.  
```{r}
set.seed(49)

rflip(n = 20, prob = 0.4)
```

### Question 9  
Answer.  
```{r}
set.seed(49)

sim_toss <- do(1000) * rflip(20, 0.4)
head(sim_toss)
```


### Question 10  
Answer.  
```{r}
set.seed(19)

sample(1:1998, 25, replace = F)
```


### Question 11  
Answer.  
```{r}
set.seed(49)

mean(sample(c(1, 0), 20, prob = c(0.4, 0.6), replace = T))
```

### Question 12  
Answer.  
```{r}
hyp_null <- 75 / 100 # the null probability
hyp_alt <- 8 / 10 # the alternative probability 
```


### Question 13  
Answer.  
```{r}
set.seed(32)
n <- 10
N <- 1000

sim_shoot_prob <- numeric(length = N)

for(i in 1:N){
    shooting = sample(c(0, 1), size = 10, replace = T, prob = c(1 - hyp_null, hyp_null))
    sim_shoot_prob[i] <- mean(shooting)
}
```


### Question 14  
Answer.  
```{r}
dotPlot(sim_shoot_prob)
```

```{r}
plot(density(sim_shoot_prob, kernel = "gaussian", bw = 0.3), 
     main = "density plot", 
     xlab = NULL)
abline(v = mean(sim_shoot_prob), lty = 2, col = "blue", lwd = 2)
```


### Question 15  
Answer.  
```{r}
mean(sim_shoot_prob > hyp_alt)
```


### Question 16  
Answer.  
```{r}
z_score <- (sim_shoot_prob - mean(sim_shoot_prob)) / sd(sim_shoot_prob)
```


### Question 17  
Answer.  
```{r}
z <- (hyp_alt - hyp_null) / (sqrt((hyp_null * (1- hyp_null))/length(sim_shoot_prob)))
pnorm(z, mean(z_score), sd(z_score), lower.tail = F)
```

the player's free throws is indeed higher than the average of NBA player.    


### Question 18  
```{r}
hyp_alt <- 10 / 10 # alternative probability
z <- (hyp_alt - hyp_null) / (sqrt((hyp_null) * (1 - hyp_null)) / length(sim_shoot_prob))
pnorm(z, mean(z_score), sd(z_score), lower.tail = F)
```

- the player's free throws rate is higher than the average of the NBA player.  

- limitation: the test player should make multiple group of shooting and use the average shooting rate.  
