---
title: "function"
author: "YinCY"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Function components
All R functions have three parts:  

- the **body**, the code inside the function  
- the **formals**, the list of arguments which controls how you can call the function.  
- the **environment**, the 'map' of the location of the function's variables.  

When you print a function in R, it shows you these three important components. If the environment isn't displayed, it means that the function was created in the global environment.  

```{r}
f <- function(x) x^2
f
```

```{r}
formals(f)
```

```{r}
body(f)
```

```{r}
environment(f)
```


# Lexical scoping
Scoping is the set of rules that govern how R looks up the value of a symbol. In the example below, scoping is the set of rules that R applies to go from the symbol x to its value 10:  

```{r}
x <- 10
x
```

Understanding scoping allows you to:  

- build tools composing functions  
- overrule the usual evaluation rules and do non-standard evaluation  

R has two types of scoping:  

- **lexical scoping**, implemented automatically at the language level.  
- **dynamic scoping**, used in select functions to save typing during interactive analysis.  


```{r}
f <- function(){
  x <- 1
  y <- 2
  c(x, y)
}

f()
```

**If a name isn't defined inside a function, R will look one level up**.  

The same rules apply if a function is defined inside another function: look inside the current function, then where that function was defined, and so on, all the way up to the global environment, and then on to other loaded packages.  

The same rules apply to **closures**, functions created by other functions.  

```{r}
x <- 1
f <- function(){
  y <- 2
  c(x, y)
}

f()
```

```{r}
j <- function(x){
  y <- 2
  function(){
    c(x, y)
  }
}

k <- j(1)
k()
rm(j, k)
```


# Dynamic lookup
Lexical scoping determines where to look for values, not when to look for them. R looks for values when the function is run, not when it’s created. This means that the output of a function can be different depending on objects outside its environment.  

```{r}
f <- function() x + 1
x <- 15
f()

f <- function() x + 1
x <- 20
f()
```

```{r}
# find all the external dependence of a function
codetools::findGlobals(f)
```


# `...` (dot-dot-dot)  
Functions can have a special argument `...`. With it, a function can take any number of additional arguments. In other programming languages, this type of argument is often called *varargs* (short for variable arguments), and a function that uses it is said to be variadic.  

1. You can use `...` to pass those additional arguments on to another function.  

```{r}
i01 <- function(y, z){
    list(y = y, z = z)
}

i02 <- function(x, ...){
    i01(...)
}

i02(x = 1, y = 2, z = 3) %>% str
```

2. Using a special form, `..N`, it's possible (but rarely useful) to refer to elements of `...` by position.  

```{r}
i03 <- function(...){
    list(first = ..1, third = ..3)
}

i03(1, 2, 3) %>% str
```

3. More useful is `list(...)`, which evaluates the arguments and stores them in a list.  

```{r}
i04 <- function(...){
    list(...)
}

i04(a = 1, b = 2) %>% str
```




