---
title: "journal"
author: 
- Chunyou Yin
date: 2020-05-29
output: 
    html_document:
        toc: false
        highlight: tango
        
bibliography: library.bib  

csl: optometry-and-vision-science.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, messages = F, warning = F)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(monocle3)
library(nichenetr)
library(cowplot)
```


### Abstract  
The well known function of chroiocapillaris (CC) is the supplying oxygen and nutrients to the out retina and retina pigment epithelial (RPE) while remove waste. Besides oxygen and nutrients supplied by CC is important for RPE's survive, signals come from CC should also important for maintain RPE's homeostatsis. To date, most studies of the pathgenesis of age-related macular degeneration (AMD) are mainly focused on the RPE itself and largely ignored cells in RPE's microenvironment that might contribute to the progress of AMD. Recently Optical coherence tomography angiography (OCTA) and Histopathological studies found that changes have taken place in the CC at the early stage of AMD. Understand how the changes take place at CC and it's impact on the RPE will contribute to the understand of pathgenesis of age-related macular degeneration. Here we take the first attempt to explor how CC endothelial cells's change might contribute to the progress of AMD and find that TGFB1 expressed by choriocapillaris endothelail cells by regulate COL10A1 gene participate in the pregression of AMD.  

> Key words: age-related macular degeneration; choriocapillaris ; retina pigment epithelial 


### Introduction   
Age-related macular degeneration (AMD) is an leading cause of sever visual loss and blindness among the eldly population[@Wong2014a]. It's pathgenesis is poorly understood and it apears to involve a complex interaction of metabolic, functional, genetic and environment factors[@Evans2001;@Kaarniranta2019;@Kauppinen2016;@Wong2014a;@Fine1983;@Fritsche2013]. Genome-wide association studies (GWAS) conducted by The AMD Gene Consortium find numerous sites associated with AMD, which implies complement activity, lipid metabolism, extracellular matrix remodeling and angiogenesis is involved in the progress of AMD progression[@Fritsche2013].  

In light of the importance of RPE for the survive of photoreceptors, most of the research about the pathgenesis of AMD are focused on the retina-pigment epithelial (RPE), and ignored the importance of microenvirnment in which RPE might be affect by other cells. Histopathological and the recent advancement in imaging technology, Optical coherence tomography (OCT), have found that in the very early stage of AMD, notable changes have identified at the choriocapillaris, such as reduced blood flow, choriocapillaris dropout[@Sohn2019;@Chatziralli2018;@Seddon2016;@Parviz2017;@Wakatsuki2015]. This implies that the changes take place at CC is ahead of RPE and it might responsible for the onset of AMD. But the mechinism of how CC changes may affact retina degeneration is largely unexplored. In the light of above findings, here we trying to understand how CC might affact RPE in the progress of AMD, and find that TGFB1 expressed by CC endothelial cells affact the expression of AMD related gene COL10A1 in RPE and this might contribute to the progression of AMD.  


### Results  
#### Samples  
To find data suitable for our purpose, we searched the Gene Expression Omnibus database (GEO), and find the CD31 enriched single-cell RNA-seq of choriodal described by Voigt[@Voigt2019a], this data is used to define the potential ligands expressed by the choriocapillaris endothelial cells that might regulator macular degeneration related genes expressed by RPE. Single cell trajectory were conducted on choriocaplaris endothelial cells among CD31 enriched macular samples to find the differentially expressed ligands along the trajectory.  

For the RPE data, we use the bulk RNA-seq data mentioned by Kim[@Kim2018a], which they describe the siginificant impact on the anti-sense transcripte between the normal and AMD patients, this data is used to define the background expressed genes and expressed age-related macular degeneration genes at early AMD. Differential expression analysis of targets genes was carried out with edgeR-limma Bioconductor packages between normal and early AMD samples.  

#### ligand target paris  
Based on prior knowlege on signaling and gene regulatory networks, we use network propagation methods on the integrated networks to infer active ligands expressed by CC endothelial cells and their gene regulatory effects on RPE using methods metioned by Browaeys[@Browaeys2019], and we find numerous ligand-target paris 
```{r, echo=FALSE, message=FALSE}
ligand_target_heatmap <- readRDS(file = "../Retina/res/ligand_target_heatmap.rds")
ligand_activity_heatmap <- readRDS(file = "../Retina/res/ligand_activity_heatmap.rds")

plot_grid(
  ligand_activity_heatmap +
    ggtitle(label = "ligand activity") +
    theme(plot.title = element_text(size = 9, face = "bold", hjust = "0.5"), 
          legend.position = "right", 
          legend.box.spacing = unit(0, "cm")),
  ligand_target_heatmap +
  theme(legend.position = "top", 
        plot.title = element_text(size = 12, face = "bold", hjust = 0.5)), 
  nrow = 1, 
  rel_widths = c(1.8, 5))
```


#### Liagnds differentially expressed along trajectory  


#### Targets differentially expressed among normal and early AMD patients  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
expression_targets_RPE_boxplot <- readRDS(file = "../Retina/res/expression_targets_RPE_boxplot.rds")
expression_targets_RPE_boxplot
```






### Discussion  


### Data and code  
CD31 enriched choriocapillaris endothelial single-cell RNA-seq data: GSE139522.  
bulk RNA-seq RPE data: GSE99248 
code is available at GitHub: https://github.com/YinCY123/AMD

### Acknowledgements   




### References  

