---
title: "xml2"
author: "YinCY"
date: "6/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Node Modification
## Modifying Existing XML
Modifying existing XML can be done in xml2 by using the replacement functions of the accessors. They all have methods for both individual xml_node objects as well as xml_nodeset objects. If a vector of values is provided it is applied piecewise over the nodeset, otherwise the value is recycled.  

### Text Modification
Text modification only happens on text nodes. If a given node has more than one text node only the first will be affected. If you want to modify additional text nodes you need to select them explicitly with `/text()`.  

```{r}
x <- read_xml(x = "<p>This is some <b>text</b>. This is more.</p>")
xml_text(x)

xml_text(x) <- "This is some other text."
xml_text(x)

# you can avoid this by explicitly selecting the text node
x <- read_xml(x = "<p>This is some text. This is <b>bold!</b></p>")
text_only <- xml_find_all(x = x, xpath = "//text()")

xml_text(text_only) <- c("This is some other text.", "Still bold!")
xml_text(x)

xml_structure(x)
```


