---
title: "xml2"
author: "YinCY"
date: "6/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Node Modification
## Modifying Existing XML
Modifying existing XML can be done in `xml2` by using the replacement functions of the accessors. They all have methods for both individual `xml_node` objects as well as `xml_nodeset` objects. If a vector of values is provided it is applied piecewise over the nodeset, otherwise the value is recycled.  

### Text Modification
Text modification only happens on text nodes. If a given node has more than one text node only the first will be affected. If you want to modify additional text nodes you need to select them explicitly with `/text()`.  

```{r}
library(xml2)
x <- read_xml(x = "<p>This is some <b>text</b>. This is more.</p>")
xml_text(x)

xml_text(x) <- "This is some other text."
xml_text(x)

# you can avoid this by explicitly selecting the text node
x <- read_xml(x = "<p>This is some text. This is <b>bold!</b></p>")
text_only <- xml_find_all(x = x, xpath = "//text()")

xml_text(text_only) <- c("This is some other text.", "Still bold!")
xml_text(x)

xml_structure(x)
```

### Attribute and Namespace Definition Modification
Attributes and namespace definitions are modified one at a time with `xml_attr()` or all at once with `xml_attrs()`. In both cases using NULL as the value will remove the attribute completely.  

```{r}
x <- read_xml(x = "<a href='invalid!'>xml2</a>")
xml_attr(x, "href")

xml_attr(x, "href") <- "https://github.com/r-lib/xml2"
xml_attr(x, "href")

xml_attrs(x) <- c(id = "xml2", href = "https://github.com/r-lib/xml2")
xml_attrs(x)

x

xml_attrs(x) <- NULL
x

xml_attr(x, "xmlns") <- "http://foo"
x

xml_attr(x, "xmlna:bar") <- "http://bar"
x
```


### Name Modification
Node names are modified with `xml_name()`.  

```{r}
x <- read_xml("<a><b/></a>")
x

xml_name(x)
xml_name(x) <- "c"
x
```

## Node modification
All of these functions have a `.copy` argument. If this is set to `FALSE` they will remove the new node from its location before inserting it into the new location. Otherwise they make a copy of the node before insertion.

### Replacing existing nodes
```{r}
x <- read_xml("<parent><child>1</child><child>2<child>3</child></child></parent>")
xml_structure(x)
children <- xml_children(x)
t1 <- children[[1]]
t2 <- children[[2]]
t3 <- xml_children(children[[2]])[[1]]

xml_replace(t1, t3)
x
```

### Add a sibling
```{r}
x <- read_xml("<parent><child>1</child><child>2<child>3</child></child></parent>")
children <- xml_children(x)
t1 <- children[[1]]
t2 <- children[[2]]
t3 <- xml_children(children[[2]])[[1]]

xml_add_sibling(t1, t3)
x

xml_add_sibling(t1, t3, where = "after")
x
```


### Add a child
```{r}
x <- read_xml("<parent><child>1</child><child>2<child>3</child></child></parent>")
children <- xml_children(x)
t1 <- children[[1]]
t2 <- children[[2]]
t3 <- xml_children(children[[2]])[[1]]

xml_add_child(t1, t3)
x

xml_add_child(t1, .value = read_xml("<test/>"))
x
```

### Removing nodes
The `xml_remove()` can be used to remove a node (and its children) from a tree. The default behavior is to unlink the node from the tree, but does *not* free the memory for the node, so R objects pointing to the node are still valid.  

This allows code like the following to work without crashing R.  

```{r}
x <- read_xml("<foo><bar><baz/></bar></foo>")
x1 <- x %>% xml_children() %>% .[[1]]
x2 <- x1 %>% xml_children() %>% .[[1]]

xml_remove(x1)
rm(x1)
x2
```

If you are not planning on referencing these nodes again this memory is wasted. Calling `xml_remove(free = TRUE)` will remove the nodes and free the memory used to store them. **Note** In this case any node which previously pointed to the node or its children will instead be pointing to free memory and may cause R to crash. xml2 can’t figure this out for you, so it’s your responsibility to remove any objects which are no longer valid.  

```{r}
x <- read_xml("<a><b /><b><b /></b></a>")
xml_structure(x)
bees <- xml_find_all(x, "//b")
xml_remove(xml_child(x), free = TRUE)
rm(bees)
gc()
```

## Namespace
We want to construct a document with the following namespace layout.

```
<?xml version = "1.0" encoding="UTF-8"?>
<sld xmlns="http://www.o.net/sld"
     xmlns:ogc="http://www.o.net/ogc"
     xmlns:se="http://www.o.net/se"
     version="1.1.0" >
<layer>
<se:Name>My Layer</se:Name>
</layer>
</sld>
```

```{r}
d <- xml_new_root(.value = "sld", 
                  xmlns = "http://www.o.net/sld", 
                  "xmlns:ogc" = "http://www.o.net/ogc", 
                  "xmlns:se" = "http://o.net/se", 
                  version = "1.1.0") %>% 
    xml_add_child("layer") %>% 
    xml_add_child("se:Name", "My Layer") %>% 
    xml_root()
d
```



