---
title: "Matrix"
author: "yincy"
date: "1/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction  
The most automatic way to use the **Matrix** package is via the `Matrix()` function which is very similar to the standard R function `matrix()`.  

```{r}
library(Matrix)

M <- Matrix(data = 10 + 1:28, nrow = 4, ncol = 7)
M
```

```{r}
t(M)
```

```{r}
cbind(1, M)
```

```{r}
M[2, 1]
```

```{r}
M[4, ]
```

```{r}
M2 <- cbind(1, M)
M2[, c(2, 4, 6)] <- 0
M2[2, ] <- 0
M2 <- rbind(0, M2, 0)
M2[1:2, 2] <- M2[3, 4:5] <- NA
M2
```

```{r}
sM <- as(M2, Class = "sparseMatrix")
10 * sM
```

```{r}
identical(sM * 2,  sM + sM)
```

```{r}
is(sM / 10 + M2 %% 2, "sparseMatrix")
```

Where the last three calls show that multiplication by a scalar keeps sparcity, as does other arithmetic, but addition to a "dense" object does not, as you might have expected after some thought about "sensible" behavior.  

```{r}
sM + 10
```


```{r}
Mg2 <- (sM > 2)
Mg2
```

```{r}
str(Mg2)
```

```{r}
summary(Mg2)
```

As you see from both of these, Mg2 contains “extra zero” (here FALSE) entries; such sparse matrices may be created for different reasons, and you can use drop0() to remove (“drop”) these extra zeros.  

```{r}
Mg2 <- drop0(Mg2)
str(Mg2@x)
```

For large sparse matrices, visualization (of the sparsity pattern) is important, and we provide `image()` methods for that.  

```{r}
data(CAex)

print(image(CAex, main = "image(CAex)"))
```

# Matrix Classes  
The **Matrix** package provides classes for real (stored as double precision), logical and so-called "pattern" (binary) dense and sparse matrices. There are provisions to also provide integer and complex (stored as double precision complex) matrices.  

Note that in R, logical means entries TRUE, FALSE, or NA. To store just the non-zero pattern for typical sparse matrix algorithms, the pattern matrices are *binary*, i.e., conceptually just TRUE or FALSE. In **Matrix**, the pattern matrices all have class names starting with "n" (patter**n**).  

## Classes for dense matrices  
For the sake of brevity, we restrict ourselves to the real (double) classes, but they are paralleled by logical and pattern matrices for all but the positive definite ones.  

- **dgeMatrix** Real matrices in general storage mode.  
- **dsyMatrix** Symmetric real matrices in non-package storage.  
- **dspMatrix** Symmetric real matrices in packed storage (one triangle only).  
- **dtrMatrix** Triangular real matrices in non-packed storage.  
- **dtpMatrix** Triangular real matrices in packed storage (triangle only).  
- **dpoMatrix** Positive semi-definite symmetric real matrices in non-packed storage.  
- **dppMatrix** ditto in packed storage.  















