---
title: "tidygraph"
author: "yincy"
date: "11/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction  
As the name suggests, `tidygraph` is an entry into the tidyverse that provides a tidy framework for all things relational (networks/graphs, trees, etc.).   


## Tidy network data?  
There’s a discrepancy between relational data and the tidy data idea, in that relational data cannot in any meaningful way be encoded as a single tidy data frame. On the other hand, both node and edge data by itself fits very well within the tidy concept as each node and edge is, in a sense, a single observation. Thus, a close approximation of tidyness for relational data is two tidy data frames, one describing the node data and one describing the edge data.  


## The tbl_graph object  
Underneath the hood of `tidygraph` lies the well-oiled machinery of `igraph`, ensuring efficient graph manipulation.   

Rather than keeping the node and edge data in a list and creating `igraph` objects on the fly when needed, `tidygraph` subclasses `igraph` with the `tbl_graph` class and simply exposes it in a tidy manner. This ensures that all your beloved algorithms that expects `igraph` objects still works with `tbl_graph` objects.  

Further, `tidygraph` is very careful not to override any of `igraph`s exports so the two packages can coexist quite happily.  
To underline the tidyness of the `tbl_graph` class the print method shows the object as two tibbles along with additional network information.  

```{r}
library(tidygraph)

create_ring(n = 10)
```

`tbl_graph` objects can be created directly using the `tbl_graph` function that takes a node data.frame and an edge data.frame. On top of that, `tidygraph` also provides coercion from a huge amount of relational data structures.  


The following list gives the package/classes that can currently be converted to `tbl_graph`s, using the `as_tbl_graph` function:  
- `data.frame`, `list`, `matrix` from base  
- `igraph` from `igraph` package  
- `network` from `network` package  
- `dendrogram` and `hclust` from `stats` package  
- `Node` from `data.tree` package  
- `phylo` and `evonet` from `ape` package  
- `graphNEL`, `graphAM`, `graphBAM` from `graph` Bioconductor package  

```{r}
iris_clust <- hclust(dist(iris[, 1:4]))
iris_tree <- as_tbl_graph(x = iris_clust)
```

Lastly, tidygraph also wraps the multitude of graph constructors available in igraph and exports them under the **create_*()** family of functions for deterministic constructors (e.g. the call to create_ring(10) above) and the **play_*()** family for constructors that incorporate sampling (e.g. play_erdos_renyi() for creating graphs with a fixed edge probability). All of these functions provide a consistent argument naming scheme to make them easier to use and understand.  

## Meet a new verb  
There are many ways a multitable setup could fit into the tidyverse. There could be an added qualifier to the verbs such as mutate_nodes() and filter_edges() or each verb could take an additional argument specifying what is targeted e.g. arrange(..., target = 'nodes'). Both of these approachable are viable but would require a huge amount of typing as well as being taxing to support down the line.  

The approach used by `tidygraph` is to let the data object itself carry around a pointer to the active data frame that should be the target of manipulation. This pointer is changed using the `activate`() verb, which, on top of changing which part of the data is being worked on, also changes the print output to show the currently active data on top:  
```{r}
iris_tree %>% activate(edges)
```

As can be seen, `activate`() takes a single argument specifying the part of the data that should be targeted for subsequent operations as an unquoted symbol. `tidygraph` continues the naming conventions from `ggraph` using `nodes` and `edges` to denote the entities and their connections respectively, but `vertices` and `links` are allowed synonyms inside `activate`().  

The current active data can always be extracted as a tibble using `as_tibble()`  
```{r}
as_tibble(iris_tree)
```


## The dplyr verbs  
Using activate() it is possible to use the well known dplyr verbs as one would expect without much hassle:  
```{r}
iris_tree <- iris_tree %>% 
  activate(nodes) %>% 
  mutate(Species = ifelse(leaf, as.character(iris$Species)[label], NA)) %>% 
  activate(edges) %>% 
  mutate(to_setose = .N()$Species[to] == "setosa")
```
 
In the above the `.N()` function is used to gain access to the node data while manipulating the edge data. Similarly `.E()` will give you the edge data and `.G()` will give you the `tbl_graph` object itself.  

Some verbs have effects outside of the currently active data. `filter()/slice()` on node data will remove the edges terminating at the removed nodes and `arrange()` on nodes will change the indexes of the `to` and `from` column in the edge data.  

While one might expect all of `dplyr`s verbs to be supported in that manner, there is a clear limitation in the relational data structure that requires rows to maintain their identity. Thus, `summarise`() and `do`() are not allowed as there is no clear interpretation of how alterations on the node and edge data with these verbs should be interpreted. If these operations are required I suggest applying them to a tibble representation and then joining the result back in.  

Speaking of joining, all joins from `dplyr` are supported. Nodes and edges are added and removed as required by the join. New edge data to be joined in must have a `to` and `from` column referencing valid nodes in the existing graph.  
```{r}
library(dplyr)

iris_sum <- iris %>% 
  group_by(Species) %>% 
  summarise_all(mean) %>% 
  ungroup()
```

```{r}
iris_tree <- iris_tree %>% 
  activate(nodes) %>% 
  left_join(iris_sum)

iris_tree
```

## Expanding the vocabulary  
On top of what has been showed so far, `tidygraph` provides an assortment of graph specific verbs that can be used to power your analysis and manipulation. Analogous to `bind_rows`(), `tidygraph` provides three functions to expand your data: `bind_nodes`() and `bind_edges`() append nodes and edges to the graph respectively. As with the join functions `bind_edges`() must contain valid from and to columns.   

`bind_graphs`() allows you to combine multiple graphs in the same graph structure resulting in each original graph to become a component in the returned graph.  

While `bind_graphs`() cannot be used to create edges between the merged graphs `graph_join`() can do just that. It merges nodes using a `full_join`() semantic and keeps the individual edges from both graphs:  

```{r}
gr1 <- create_notable(name = "bull") %>% 
  mutate(name = letters[1:5])

gr2 <- create_ring(n = 5) %>% 
  mutate(name = letters[4:8])
```

```{r}
gr1 %>% bind_graphs(gr2) %>% 
  ggraph(layout = "kk") +
  geom_edge_link()+
  geom_node_point(size = 8, color = "steelblue") +
  geom_node_text(aes(label = name), color = "white", vjust = 0.4) +
  ggtitle(label = "Binding graphs") +
  theme_graph()
```


```{r}
gr1 %>% graph_join(gr2, by = 'name') %>% 
  ggraph(layout = "kk") +
  geom_edge_link()+
  geom_node_point(size = 8, color = "steelblue") +
  geom_node_text(aes(label = name), color = "white", vjust = 0.4) +
  ggtitle("Joining graphs") +
  theme_graph()
```


The standard `dplyr` verbs protects the `to` and `from` columns in the edge data in order to avoid accidental modification of the graph topology. If changing of the terminal nodes are necessary the `reroute`() verb will come in handy:  

```{r}
gr1 <- create_star(n = 6, directed = T)
layout <- create_layout(graph = gr1, layout = "fr")

gr1 <- gr1 %>% 
  mutate(x = layout$x, y = layout$y, graph = "original")

gr2 <- gr1 %>% 
  mutate(graph = "reverse") %>% 
  activate(edges) %>% 
  reroute(from = to, to = from)

gr3 <- gr1 %>% 
  mutate(graph = "using subset") %>% 
  activate(edges) %>% 
  reroute(from = to + 1, subset = to < 4)

ggraph(bind_graphs(gr1, gr2, gr3), layout = "nicely") +
  geom_edge_link(arrow = arrow(length = unit(x = 0.15, units = "inches"))) +
  geom_node_point(size = 2)+
  facet_nodes(~graph) +
  theme_graph(foreground = "steelblue")
```

As can be seen, reroute works pretty much as a `to` and `from` specific `mutate`(), with the added benefit of incorporating a subset operator if only a few edges should be changed.  


## Making the most of graphs  
Central to all of these functions is that they know about which graph is being computed on (in the same way that n() knows about which tibble is currently in scope). Furthermore they always return results matching the node or edge position so they can be used directly in mutate() calls.  

## Node and edge types  
On the top of our list of things we might be interested to know about is whether nodes or edges are of specific types such as leaf, sink, loop, etc. All of these functions return a logical vector indicating if the node or edge belong to the specified group. To easily find functions that queries types, all functions are prefixed with **node_is_*/edge_is_***.  
```{r}
create_tree(n = 20, children = 3) %>% 
  mutate(leaf = node_is_leaf(), root = node_is_root()) %>% 
  ggraph(layout = "tree")+
  geom_edge_diagonal() +
  geom_node_point(aes(filter = leaf), color = "forestgreen", size = 10) +
  geom_node_point(aes(filter = root), color = "firebrick", size = 10) +
  theme_graph()
```

Another example could be to remove loop edges using `filter(!edge_is_loop())`  

## Centrality  
One of the simplest concepts when computing graph based values is that of centrality, i.e. how central is a node or edge in the graph. As this definition is inherently vague, a lot of different centrality scores exists that all treat the concept of central a bit different. One of the famous ones is the pagerank algorithm that was powering Google Search in the beginning.  

`tidygraph` currently has 11 different centrality measures and all of these are prefixed with `centrality_*` for easy discoverability. All of them returns a numeric vector matching the nodes (or edges in the case of `centrality_edge_betweenness()`).  

```{r}
play_smallworld(n_dim = 1, dim_size = 100, order = 3, p_rewire = 0.05) %>% 
  mutate(centrality = centrality_authority()) %>% 
  ggraph(layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(size = centrality, color = centrality)) +
  scale_color_continuous(guide = "legend") +
  theme_graph()
```

## Clustering  
Another common operation is to group nodes based on the graph topology, sometimes referred to as community detection based on its commonality in social network analysis. All clustering algorithms from `igraph` is available in `tidygraph` using the **group_* ** prefix. All of these functions return an integer vector with nodes (or edges) sharing the same integer being grouped together.  

```{r}
play_islands(n_islands = 5, size_islands = 10, p_within = 0.8, m_between = 3) %>% 
  mutate(community  = as.factor(group_infomap())) %>% 
  ggraph(layout = "fr") +
  geom_edge_link(aes(alpha = ..index..), show.legend = F) +
  geom_node_point(aes(color = community), size = 7) +
  theme_graph()
```

## Node pairs  
Some statistics are a measure between two nodes, such as distance or similarity between nodes. In a tidy context one of the ends must always be the node defined by the row, while the other can be any other node.  

All of the node pair functions are prefixed with node_* and ends with _from/_to if the measure is not symmetric and _with if it is; e.g. there’s both a node_max_flow_to() and node_max_flow_from() function while only a single node_cocitation_with() function. The other part of the node pair can be specified as an integer vector that will get recycled if needed, or a logical vector which will get recycled and converted to indexes with which(). This means that output from node type functions can be used directly in the calls, e.g.  
```{r}
play_geometry(n = 50, radius = 0.25) %>% 
  mutate(dist_to_center = node_distance_to(nodes = node_is_center())) %>% 
  ggraph(layout = "kk") +
  geom_edge_link() +
  geom_node_point(aes(size = dist_to_center), color = "steelblue") +
  scale_size_continuous(range = c(6, 1)) +
  theme_graph()
```

## Searches  
**An integral type of operation on graphs is to perform a search, that is, start from one node and then traverse the edges until all nodes has been visited**. The most common approaches are either **breath first search** where all neighbors of a node is visited before moving on to the next node, or **depth first search** where you move along to the next node immediately and only backtracks and visit other neighbors when you’ve hit a dead end.  

Different statistics from these searches are available in `tidygraph` through the **bfs_*()** and **dfs_*()** family of functions e.g. the distance to the start node along the search can be obtained with **bfs_dist**()/**dfs_dist**().  

The root node can be specified in the same way as with node pairs. Sorting based on a search from the node with highest centrality can thus be done with:  
```{r}
play_geometry(n = 50, radius = 0.25) %>% 
  mutate(order = bfs_rank(which.max(centrality_alpha())))
```
 

## Local measures of a node   
Often we find ourselves interested in the local neighborhood of a node for various reasons. We might want to know the average degree around a node or the number of triangles each node participate in. The `local_*()` family of functions provide access to a range of node measures that are dependent on the local neighborhood of each node.  
```{r}
play_smallworld(n_dim = 1, dim_size = 100, order = 3, p_rewire = 0.05) %>% 
  mutate(weighted_degree = centrality_degree() / local_ave_degree())
```

## Graph measures  
Along with computations on the individual nodes and edges it is sometimes necessary to get summary statistics on the graph itself. These can be simple measures such as the number of nodes and edges as well as more involved measures such as assortativity (the propensity of similar nodes to be connected).  

All of these measures can be calculated through the `graph_*()` function family and they will all return a scale.  
```{r}
play_islands(n_islands = 5, size_islands = 10, p_within = 0.7, m_between = 3) %>% 
  mutate(norm_adhesion = node_adhesion_to(nodes = c(50, 1:49)) / graph_adhesion())
```

## Mapping over nodes  



























