---
title: "02 data structure"
author: "yincy"
date: "3/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# storage mode
storage mode describes how the object is stored.  

1. numeric
2. character
3. logical
4. complex
5. raw

# base data class
the class attribute tells generic function how to handle this object.  

1. vector
2. data.frame
3. matrix
4. array
5. list

```{r}
with(mtcars,
     {
         summary(mpg)
         plot(mpg, disp)
         abline(lm(disp ~ mpg))
         plot(mpg, wt)
         abline(lm(wt ~ mpg))
     })

with(mtcars,{
    nokeep <- summary(mpg)
    keep <<- summary(mpg) # special assignment to keep the assignment
})
keep
nokeep
```


```{r}
library(tibble)
as_tibble(mtcars)
```

- `tibble` never change the names of variable.  
- `tibble` never convert character variables to factors.  
- subsetting a `tibble` always returns a `tibble`.  
- `tibble` don't support row names.  


# functions to manipulate Connections
```{r}
?file
```





















