---
title: "09_Analysis-of-variance"
author: "yincy"
date: "3/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


When factors are included as explanatory variables, our focus usually shifts from prediction to understanding group differences, and the methodology is referred to as analysis of variance (ANOVA).  

ANOVA methodology is used to abalyze a wide variety of experimental and quasi-experimental designs.  


## Fit ANOVA models

**formulas for several common research designs**  

|**Design**|**Formula**|
|:--|:--|
|One-way ANOVA|y ~ A|
|One-way ANCOVA with a covariate|y ~ x + A|
|Two-way factorial ANOVA|y ~ A * B|
|Two-way factorial ANCOVA with 2 covariates|y ~ x1 + x2 + A * B|
|Randomized block|y ~ B + A (where B is a blocking factor)|
|One-way within-groups ANOVA|y ~ A + Error(Subject/A)|
|Repeated measures ANOVA with 1 within-groups factor(W) and 1 between-groups factor (B)|y ~ B * W + Error(Subject/W)|

- lowercase letters are quantitative variables  
- uppercase letters are grouping factors  
- `Subject` is a unique identifier variable for subjects

### One-way ANOVA
In a one-way ANOVA, the dependent variable is assumed to be normally distributed and have equal variance in each group.  


```{r}
library(dplyr)
data(cholesterol, package = "multcomp")

plotdata <- cholesterol %>% 
    group_by(trt) %>% 
    summarise(
        n = n(), 
        mean = mean(response), 
        sd = sd(response), 
        ci = qt(0.975, df = n - 1) * sd / sqrt(n)
    )

fit <- aov(response ~ trt, data = cholesterol)
summary(fit)
```

```{r}
library(ggplot2)

plotdata %>% 
    ggplot(aes(trt, mean, group = 1)) +
    geom_point(size = 3, color = "red") +
    geom_line(linetype = 2, color = "darkgrey") +
    geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.1) +
    theme_bw() +
    labs(x = "Treatment", 
         y = "Response", 
         title = "Mean Plot with 95% Confidence Interval")
```

```{r}
pairwise <- TukeyHSD(fit)
pairwise
```


```{r}
plotdata <- pairwise[[1]] %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column("conditions")

plotdata %>% 
    ggplot(aes(conditions, diff)) +
    geom_point(size = 3, color = "red") +
    geom_errorbar(aes(ymin = lwr, ymax = upr), width = 0.2) +
    geom_hline(yintercept = 0, color = "red", linetype = 2) +
    labs(y = "Difference in mean levels", 
         x = NULL, 
         title = "95% family-wise confidence level") +
    theme_bw() +
    coord_flip()
```


```{r}
library(car)
fit <- aov(response ~ trt, data = cholesterol)
qqPlot(fit, simulate = T, main = "Q-Q Plot")
```


```{r}
outlierTest(fit)
```


## One-way ANCOVA
A one-way analysis of covariance (ANCOVA) extends the one-way ANOVA to include one or more quantitative covariates.  

ANCOVA designs make the same normality and homogeneity of variance assumptions described for ANOVA designs.  

In addition, standard ANCOVA designs assume homogeneity of regression slopes.  

A test for the homogeneity of regression slopes can be obtained by including a gestation × dose interaction term in your ANCOVA model.  


```{r}
library(multcomp)
library(dplyr)
litter %>% 
    group_by(dose) %>% 
    summarise(
        n = n(), 
        mean = mean(gesttime), 
        sd = sd(gesttime)
    )

fit <- aov(weight ~ gesttime + dose, data = litter)
summary(fit)
```

obtain adjusted group means using `effect()` function from `effects` package.  
```{r}
library(effects)
effect(term = "dose", mod = fit)
```

```{r}
library(multcomp)
contrast <- rbind("no drug vs. drug" = c(3, -1, -1, -1))
glht(fit, linfct = mcp(dose = contrast)) %>% summary
```

A test for the homogeneity of regression slopes can be obtained by including a gestation × dose interaction term in your ANCOVA model.  

```{r}
library(multcomp)
fit2 <- aov(weight ~ gesttime * dose, data = litter)
fit2 %>% summary
```

```{r}
library(ggplot2)
pred <- predict(fit)

cbind(pred, litter) %>% 
    ggplot(aes(gesttime, weight)) +
    geom_point() +
    facet_wrap(~ dose, nrow = 1) +
    geom_line(aes(y = pred)) +
    labs(title = "ANCOVA for weight by gesttime and dose") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.position = "none")
```


## Two-way factorial ANOVA
```{r}
library(dplyr)
data("ToothGrowth")
ToothGrowth$dose <- factor(ToothGrowth$dose)
stats <- ToothGrowth %>% 
    group_by(supp, dose) %>% 
    summarise(
        n = n(), 
        mean = mean(len), 
        sd = sd(len), 
        ci = qt(0.975, df = n - 1) * sd/sqrt(n)
    )

stats

fit <- aov(len ~ supp * dose, data = ToothGrowth)
fit %>% summary()
```


```{r}
stats %>% 
    ggplot(aes(dose, mean, group = supp, color = supp, linetype = supp)) +
    geom_point(size = 2, position = position_dodge(width = 0.2)) +
    geom_line(position = position_dodge(width = 0.2)) +
    geom_errorbar(aes(ymin = mean - ci, ymax = mean + ci), width = 0.1, position = position_dodge(width = 0.2)) +
    theme_bw() +
    scale_color_manual(values = c("blue", "red")) +
    labs(title = "Mean Plot with 95% Confidence Interval", 
         x = "Dose", 
         y = "Mean Length")
```


## Repeated measures ANOVA
In repeated measures ANOVA, subjects are measured more than once.  

```{r}
data(CO2)
CO2$conc <- factor(CO2$conc)
w1b1 <- subset(CO2, Treatment == "chilled")
fit <- aov(uptake ~ conc*Type + Error(Plant/(conc)), data = w1b1)
fit %>% summary
```

```{r}
library(dplyr)
stats <- CO2 %>% 
    group_by(conc, Type) %>% 
    summarise(
        mean_conc = mean(uptake)
    )

stats %>% 
    ggplot(aes(conc, mean_conc, group = Type, color = Type, linetype = Type)) +
    geom_point(size = 2) +
    geom_line(size = 1) +
    theme_bw() +
    theme(legend.position = "top") +
    labs(title = "Interaction Plot for Plant Type and Concentration", 
         x = "Concentration", 
         y = "Mean Uptake")
```


```{r}
CO2 %>% 
    ggplot(aes(conc, uptake, fill = Type)) +
    geom_boxplot() +
    theme_bw() +
    theme(legend.position = "top") +
    scale_fill_manual(values = c("gold", "green")) +
    labs(title = "Chilled Quebec and Mississippi Plants")
```


## Multivariate analysis of variance (MANOVA)
The two assumptions underlying a one-way MANOVA are multivariate normality and homogenity of variance-covariance matrices.  

The first assumption states that the vector of dependent variables jointly follows a multivariate normal distribution. 



```{r}
data(UScereal, package = "MASS")

shelf <- factor(UScereal$shelf)
y <- cbind(UScereal$calories, UScereal$fat, UScereal$sugars) 
colnames(y) <- c("calories", "fat", "sugars")

aggregate(y, by = list(shelf = shelf), FUN = mean)
cov(y)

fit <- manova(y ~ shelf)
fit %>% summary
summary.aov(fit)
```

assess the normality  

```{r}
center <- colMeans(y)
n <- nrow(y)
p <- ncol(y)
cov <- cov(y)
d <- mahalanobis(y, center, cov)
(coord <- qqplot(qchisq(ppoints(n), df = p), d, 
                main = "Q-Q Plot Assessing Multivariate Normality", 
                ylab = "Mahalanobi D2"))
abline(a = 0, b = 1)
```


### Robust MANOVA
```{r}
library(rrcov)

Wilks.test(y, shelf, method = "mcd")
```


## ANOVA as regression
```{r}
fit.aov <- aov(response ~ trt, data = cholesterol)
fit.aov %>% summary
```

```{r}
fit.lm <- lm(response ~ trt, data = cholesterol)
fit.lm %>% summary
```

```{r}
contrasts(cholesterol$trt)
```








