---
title: "03 XML and JSON"
author: "YinCY"
date: "6/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(XML)
library(magrittr)
```


```{r}
parsed_stocks <- xmlParse("../../../Wiley-ADCR/ch-3-xml/stocks/technology.xml", 
                          validate = T)
xmlRoot(parsed_stocks) %>% xmlSize()
```

```{r}
bond <- xmlParse("../../../Wiley-ADCR/ch-3-xml/bond.xml")
bond %>% class()
```

```{r}
(root <- xmlRoot(bond))
root %>% class()
```

```{r}
root %>% xmlName() # return the name of the root element
root %>% xmlSize() # the number of child in the root element
```

```{r}
root[[1]]
```

```{r}
root[[1]][[1]]
```

Using double brackets yields the first element in the tree, single brackets return objects of class XMLInternalNodeList.  

```{r}
root[["movie"]]
```

```{r}
root["movie"]
```

Names and numbers can also be combined.  

```{r}
root[["movie"]][[1]][[1]]
```


### From XML to data frames or lists
```{r}
xmlSApply(root[[1]], xmlValue)
xmlSApply(root, xmlAttrs)
xmlSApply(root, xmlGetAttr, "id")
```

As long as XML documents are flat in the hierarchical sense, that is, the root node’s most
distant relatives are grandchildren or children, they can usually be transformed easily into a data frame with xmlToDataFrame()  

```{r}
xmlToDataFrame(root)
```

```{r}
xmlToList(bond)
```

### Event-driven parsing
Event-driven parsing differs from DOM- style parsing in that it skips the construction of the complete DOM at the C level. Instead, event-driven parsers sequentially traverse over an XML file, and once they find a specified element of interest they prompt an instant, user-defined reaction to this event. This procedure provides a huge advantage over DOM-style parsers because the machine’s memory never has to hold the complete document.  

```{r}
branchFun <- function(){
  container_close = numeric()
  container_date = numeric()
  
  "Apple" = function(node, ...){
    date = xmlValue(xmlChildren(node)[[c("date")]])
    container_date <<- c(container_date, date)
    close <- xmlValue(xmlChildren(node)[["close"]])
    container_close <<- c(container_close, close)
  }
  getContainer <- function()data.frame(date = container_date, close = container_close)
  list(Apple = Apple, getStore = getContainer)
}

(h5 <- branchFun())
```

```{r}
xmlEventParse(file = "../../../Wiley-ADCR/ch-3-xml/stocks/technology.xml", 
              branches = h5, handlers = list()) %>% invisible()

apple.stock <- h5$getStore()
apple.stock %>% head
```


```{r}
library(RJSONIO)
ls("package:RJSONIO")
```

```{r}
isValidJSON("../../../Wiley-ADCR/ch-3-xml/indy.json")
```

```{r}
indy <- fromJSON("../../../Wiley-ADCR/ch-3-xml/indy.json")
indy %>% class()
```

```{r}
library(stringr)

indy.vec <- unlist(indy, recursive = T, use.names = T)
indy.vec %>% head

indy.vec[str_detect(names(indy.vec), "name")]
```

```{r}
sapply(indy[[1]], "[[", "year")
```

```{r}
library(plyr)
indy.unlist <- sapply(indy[[1]], unlist)
indy.df <- do.call("rbind.fill", lapply(lapply(indy.unlist, t), data.frame, stringsAsFactors = F))
indy.df %>% names
```

```{r}
library(jsonlite)

jsonlite::fromJSON("../../../Wiley-ADCR/ch-3-xml/peanuts.json") 
```

```{r}
jsonlite::fromJSON("../../../Wiley-ADCR/ch-3-xml/indy.json")[[1]]
```



