---
title: "rayshader"
author: "yincy"
date: "1/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview  
**rayshader** is an open source package for producing 2D and 3D data visualizations in R. **rayshader** uses elevation data in a base R matrix and a combination of raytracing, spherical texture mapping, overlays, and ambient occlusion to generate beautiful topographic 2D and 3D maps. In addition to maps, **rayshader** also allows the user to translate **ggplot2** objects into beautiful 3D data visualizations.  

The models can be rotated and examined interactively or the camera movement can be scripted to create animations. The user can also create a cinematic depth of field post-processing effect to direct the user’s focus to important regions in the figure. The 3D models can also be exported to a 3D-printable format with a built-in STL export function.   


## Functions  
Rayshader has seven functions related to mapping:  

- `ray_shade` uses user specified light directions to calculate a global shadow map for an elevation matrix. By default, this also scales the light intensity at each point by the dot product of the mean ray direction and the surface normal (also implemented in function `lamb_shade`, this can be turned off by setting `lambert = FALSE`).  

- `sphere_shade` maps an RGB texture to hillshade by spherical mapping. A texture can be generated with the `create_texture` function, or loaded from an image. `sphere_shade` also includes 7 build-in palettes: 'imhof1', 'imhof2', 'imhof3', 'imhof4', 'desert', 'bw', 'unicorn'.  

- `create_texture` programmatically creates texture maps given five colors: a highlight, a shadow, a left fill light, a right fill light, and a center color flat areas. The user can also optionally specify the colors at the corners, but `create_texture` fill interpolate those if they aren't given.  

- `ambient_shade` creates an ambient occlusion shadow layer, darking areas that have less scattered light from the atmosphere. This results in valleys being darker than flat areas and ridges.  

- `lamb_shade` uses a single user specified light direction to calculate a local shodow map based on the dot product between the surface normal and light direction for an elevation matrix.  

- `add_shade` take two of the shadow maps above and combines them, scaling the second one (or, if the second is an RGB array, the matrix) as specified by the user.  

- `add_overlay` takes a 3 or 4-layer RGB/RGBA array and overlays it on the current map. If the map includes transparency, this is taken into account when overlaying the image. Otherwise, the user can specify a single color that will be marked as completely transparent, or set the full overlay as partly transparent.  



Rayshader also has three functions to detect and add water to maps:  

- `detect_water` uses a flood-fill algorithm to detect bodies of water of a user-specified minimum area.   

- `add_water` users the output of `detect_water` to add a water color to the map. The user can input their own color, or pass the name of one of the pre-defined palettes from `sphere_shade` to get a matching hue.  

- `render_water` adds a 3D transparent water layer to 3D maps, after the `rgl` device has already been created. This can either add to a 3D transparent water layer to a map that does not already have a water layer, or replace an existing water layer on the map.  



Also included are two functions to add additional effects and information to your 3D visulizations:  

- `render_depth` generates a depth of field effect for the 3D map. The user can specify the focal distance, focal length, and f-stop of the camera, as well as aperture shape and bokeh intensity. This either plots the image to the local device, or saves it to a file if given a filename.  

- `render_label` adds a text label to the `x` and `y` coordinate of the map at a specified altitude `z` (in units of the matrix). The altitude can either be specified relative ti the elevation at that point (the default), or absolutely.  


Four functions to display and save your visulizations:  

- `plot_map` plots the current map. Accepts either a matrix or an array.  

- `write_png` writes the current map to disk with a user-specified filename.  

- `plot_3d` create a 3D map, given a texture and an elevation matrix. You can customize the appearance of the map, as well as add a user-defined water level.  

- `render_snapshot` saves an image of the current 3D view to disk (if given a filename), or plots the 3D view to the current device (usefull for including images in R markdown files).  

- `render_movie` create and saves a mp4 file of the camera rotating around the 3D sense by either using a built-in orbit or by using one provided by the user.  



A single function to generate 3D plots using ggplot2 objects:  

- `plot_gg` takes a ggplt2 object (or a list of two ggplot2 object) and uses the fill or color aesthetic to transform the point into 3D surface. You can pass any of the arguments used to specify the camera and the background/shadow colors in `plot_3d()`, and manipulate the displayed 3D plot using `render_camera()` and `render_depth()`.  


All of these functions are designed to be used with the magrittr pipe ` %>% `.  



## Usage  
`Rayshader` can be used for two purposes: both creating hillshaded maps and 3D data visualizations plots.  

### Mapping with rayshader   
```{r}
library(rayshader)
```


load a map with the raster package.  
```{r}
loadzip <- tempfile()
download.file("https://tylermw.com/data/dem_01.tif.zip", loadzip)
localtif <- raster::raster("dem_01.tif")
unlink(loadzip)

# convert to matrix 
elmat <- raster_to_matrix(localtif)

elmat %>% 
    sphere_shade(texture = "desert") %>% 
    plot_map()
```


sphere shader can shift the sun direction  
```{r}
elmat %>% 
    sphere_shade(sunangle = 45, texture = "desert") %>% 
    plot_map()
```


detect_water and add_water adds a water layer to the map  
```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    plot_map()
```

we can add a layer from the sun direction as well
```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    add_shadow(ray_shade(elmat), 0.5) %>% 
    plot_map()
```

add an ambient occlusion shadow layer, which models lighting from atmospheric scattering.  
```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    add_shadow(ray_shade(elmat), 0.5) %>% 
    add_shadow(ambient_shade(elmat), 0) %>% 
    plot_map()
```


```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    add_shadow(ray_shade(elmat, zscale = 3), 0.5) %>% 
    add_shadow(ambient_shade(elmat), 0) %>% 
    plot_3d(elmat, zscale = 10, fov = 0, theta = 135, zoom = 0.75, phi = 45, windowsize = c(1000, 800))
Sys.sleep(0.2)
render_snapshot(clear = TRUE)
```

You can add a scale bar, as well as compass using `render_scalebar()` and `render_compass()`  
```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    add_shadow(ray_shade(elmat, zscale = 3), 0.5) %>% 
    add_shadow(ambient_shade(elmat), 0) %>% 
    plot_3d(elmat, zscale = 10, fov = 0, theta = 135, zoom = 0.75, phi = 45, windowsize = c(1000, 800))

render_camera(fov = 0, theta = 60, zoom = 0.75, phi = 45)
render_scalebar(limits = c(0, 5, 10), 
                label_unit = "km", 
                position = "W", 
                y = 50, 
                scale_length = c(0.33, 1))
render_compass(position = 'E')
render_snapshot(clear = T)
```

You can also render using the built-in pathtracer, powered by `rayrender`. Simply replace `render_snapshot()` with `render_highquality()`. When `render_highquality()` is called, there's no need to pre-compute the shadows with any of the `_shade` functions, so we remove those.  

```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(elmat), color = "desert") %>% 
    plot_3d(elmat, zscale = 10, fov = 0, 
            theta = 60, zoom = 0.75, phi = 45, 
            windowsize = c(1000, 800))
render_scalebar(limits = c(0, 5, 10), label_unit = "km", position = "W", y = 50, 
                scale_length = c(0.33, 1))

render_compass(position = "E")

Sys.sleep(0.2)

render_highquality(samples = 200, scale_text_size = 24, clear = T)
```

You can also easily add a water layer by setting `water=TRUE` in `plot_3d()` (and setting `waterdepth` if the water level is not 0), or by using the function `render_water()` after the 3D map has been rendered. You can customize the appearance and transparancy of the water layer via function arguments. Here's an example using bathymetric/topographic data of Monterey Bay, CA.  

```{r}
montshadow <- ray_shade(heightmap = montereybay, lambert = F)
montamb <- ambient_shade(heightmap = montereybay, zscale = 50)

montereybay %>% 
    sphere_shade(zscale = 10, texture = "imhof1") %>% 
    add_shadow(shadowmap = montamb, max_darken = 0) %>% 
    add_shadow(shadowmap = montshadow, max_darken = 0.5) %>% 
    plot_3d(montereybay, zscale = 50, fov = 0, theta = -45, phi = 45, 
            windowsize = c(1000, 800), zoom = 0.75, 
            water = T, waterdepth = 0, wateralpha = 0.5, 
            watercolor = "lightblue", 
            waterlinecolor = "white", 
            waterlinealpha = 0.5)

Sys.sleep(0.2)
render_snapshot(clear = T)
```

Water is also supported in `render_highquality()`. We load the `rayrender` package to change the ground material to include a checker pattern. By default, the camera looks at the origin, but we shift if down slightly to center the map.  

```{r}
library(rayrender)

montereybay %>% 
    sphere_shade(zscale = 10, texture = "imhof1") %>% 
    plot_3d(montereybay, zscale = 50, fov = 70, theta = 270, phi = 30, 
            windowsize = c(1000, 800), zoom = 0.6, 
            water = T, 
            waterdepth = 0, 
            watercolor = "#233aa1", 
            waterlinecolor = "white", 
            waterlinealpha = 0.5)

Sys.sleep(0.2)
render_highquality(lightdirection = c(-45, 45), 
                   lightaltitude = 30, 
                   clamp_value = 10, 
                   samples = 200, 
                   camera_lookat = c(0, -50, 0), 
                   ground_material = diffuse(color = "grey50", 
                                             checkercolor = "grey20", 
                                             checkerperiod = 100))
```

Rayshader also has map shapes other than rectangular included `c("hex", "circle")`, and you can customize the map into any shape you want by setting the areas you do not want to display to NA.  

```{r, message=F, warning=F}
par(mfrow = c(1, 2))

montereybay %>% 
    sphere_shade(zscale = 10, texture = "imhof1") %>% 
    add_shadow(montshadow, max_darken = 0.5) %>% 
    add_shadow(montamb, max_darken = 0) %>% 
    plot_3d(montereybay, zscale = 50, fov = 0, 
            theta = -45, phi = 45, 
            windowsize = c(1000, 800), 
            water = T, waterdepth = 0, 
            wateralpha = 0.5, watercolor = "lightblue", 
            waterlinecolor = "white", waterlinealpha = 0.5, 
            baseshape = "circle")
render_snapshot(clear = T)

montereybay %>% 
    sphere_shade(zscale = 10, texture = "imhof1") %>% 
    add_shadow(montshadow, max_darken = 0.5) %>% 
    add_shadow(montamb, max_darken = 0) %>% 
    plot_3d(montereybay, zscale = 50, fov = 0, 
            theta = -45, phi = 45, 
            windowsize = c(1000, 800), 
            water = T, waterdepth = 0, 
            wateralpha = 0.5, watercolor = "lightblue", 
            waterlinecolor = "white", waterlinealpha = 0.5, 
            baseshape = "hex")

render_snapshot(clear = T)
```

add text labels is done with the `render_label()` function, which also allows you to customize the line type, color, and size along with the font.  

```{r}
montereybay %>% 
    sphere_shade(zscale = 10, texture = "imhof1") %>% 
    add_shadow(shadowmap = montshadow, max_darken = 0.5) %>% 
    add_shadow(shadowmap = montamb, max_darken = 0) %>% 
    plot_3d(heightmap = montereybay, zscale = 50, 
            fov = 0, theta = -100, 
            phi = 30, windowsize = c(1000, 800), 
            water = T, waterdepth = 0, 
            waterlinecolor = "white", 
            waterlinealpha = 0.5,
            wateralpha = 0.5, 
            watercolor = "lightblue")

render_label(montereybay, x = 350, y = 160, z = 1000, 
             zscale = 50, text = "Moss Landing", 
             textsize = 1, linewidth = 5)

render_label(montereybay, x = 220, y = 100, z = 7000, zscale = 50, 
             text = "Santa Cruz", textcolor = "darkred", 
             linecolor = "darkred", textsize = 1, linewidth = 5)

render_label(montereybay, x = 300, y = 270, z = 4000, zscale = 50, 
             text = "Monterey", dashed = T, textsize = 1, linewidth = 5)

render_label(montereybay, x = 50, y = 330, z = 1000, zscale = 50, 
             textcolor = "white", text = "Monterey Canyon", 
             relativez = F, textsize = 1, linewidth = 5, linecolor = "white")

Sys.sleep(0.2)

render_highquality(samples = 200, line_radius = 1, 
                   text_size = 18, text_offset = c(0, 12, 0), 
                   clamp_value = 10, clear = T)
```

You can also apply a post-processing effect to the 3D maps to render maps with depth of field with the `render_depth()` function:  

```{r}
elmat %>% 
    sphere_shade(texture = "desert") %>% 
    add_water(detect_water(heightmap = elmat), color = "desert") %>% 
    add_shadow(ray_shade(elmat, zscale = 3), max_darken = 0.5) %>% 
    add_shadow(shadowmap = ambient_shade(elmat), max_darken = 0) %>% 
    plot_3d(elmat, zscale = 10, fov = 30, theta = -225, phi = 25, 
            windowsize = c(1000, 800))

Sys.sleep(0.2) 

render_depth(focus = 0.6, focallength = 200, clear = T)
```


## 3D plotting with rayshader and ggplot2  
Rayshader can also be used to make 3D plots out of ggplot2 objects using the `plot_gg()` function. Here, I turn a color density plot into 3D density plot. `plot_gg()` detects that the user mapped the `fill` aesthetic to color and uses that information to project the figure into 3D.  

```{r}
library(ggplot2)

ggdiamonds <- ggplot(diamonds) +
    stat_density_2d(aes(x = x, y = depth, fill = stat(nlevel)), 
                    geom = "polygon", n = 100, bins = 10, contour = T) +
    facet_wrap(clarity ~.) +
    scale_fill_viridis_c(option = "A")

par(mfrow = c(1, 2))

plot_gg(ggdiamonds, width = 5, height = 5, raytrace = F, preview = T)
plot_gg(ggdiamonds, width = 5, height = 5, multicore = T, scale = 250, 
        zoom = 0.7, theta = 10, phi = 30, windowsize = c(800, 800))

Sys.sleep(0.2)
render_snapshot(clear = T)
```

Rayshader will automatically ignore lines and other elements that should not be mapped to 3D. Here's a contour plot of the `volcano` dataset.  

```{r}
library(reshape2)

ggvolcano <- volcano %>% 
    melt() %>% 
    ggplot() + 
    geom_tile(aes(x = Var1, y = Var2, fill = value)) +
    geom_contour(aes(x = Var1, y = Var2, z = value), color = "black") +
    scale_x_continuous(name = "X", expand = c(0, 0)) +
    scale_y_continuous(name = "Y", expand = c(0, 0)) +
    scale_fill_gradientn(name = "Z", colours = terrain.colors(10)) +
    coord_fixed()

par(mfrow = c(1, 2))
plot_gg(ggvolcano, width = 7, height = 4, raytrace = F, preview = T)
```

```{r}
plot_gg(ggvolcano, multicore = T, raytrace = T, width = 7, height = 4, 
        scale = 300, windowsize = c(1400, 866), 
        zoom = 0.6, phi = 30, theta = 30)

Sys.sleep(0.2)
render_snapshot(clear = T)
```

Rayshader also detects when the user passes the `color` aesthetic, and maps theose values to 3D. If both `color` and `fill` are passed, however, rayshader will default to `fill`.  

```{r}
mtplot <- ggplot(mtcars) +
    geom_point(aes(x = mpg, y = disp, color = cyl)) +
    scale_color_continuous(limits = c(0, 8)) 

par(mfrow = c(1, 2)) 
plot_gg(mtplot, width = 3.5, raytrace = F, preview = T)

plot_gg(mtplot, width = 3.5, multicore = T, windowsize = c(800, 800), 
        zoom = 0.85, phi = 35, theta = 30, sunangle = 225, soliddepth = -100)

Sys.sleep(0.2)
render_snapshot(clear = T)
```

Utilize combinations of line color and fill to create different effects. Here is a terraced hexbin plot, created by mapping the line colors of the hexagons to black.  

```{r}
a <- data.frame(x = rnorm(20000, 10, 1.9), y = rnorm(20000, 10, 1.2))
b <- data.frame(x = rnorm(20000, 14.5, 1.9), y = rnorm(20000, 14.5, 1.9))
c <- data.frame(x = rnorm(20000, 9.5, 1.9), y = rnorm(20000, 15.5, 1.9))
data <- rbind(a, b, c)

pp <- ggplot(data, aes(x, y)) +
    geom_hex(bins = 20, size = 0.5, color = "black") +
    scale_fill_viridis_c(option = "C")

par(mfrow = c(1, 2))
plot_gg(pp, width = 5, height = 4, scale = 300, raytrace = F, preview = T)
plot_gg(pp, width = 5, height = 4, scale = 300, multicore = T, windowsize = c(1000, 800))
render_camera(fov = 70, zoom = 0.5, theta = 130, phi = 35)

Sys.sleep(0.2)
render_snapshot(clear = T)
```

You can also use the `render_depth()` function to direct the viewer's focus to important areas in the figure.  

```{r}
par(mfrow = c(1, 1))
plot_gg(pp, width = 5, height = 4, scale = 300, 
        multicore = T, windowsize = c(1200, 900), 
        fov = 70, zoom = 0.4, theta = 300, phi = 40)

Sys.sleep(0.2)
render_depth(focus = 0.68, focallength = 200, clear = T)
```

Finally, you can increase the allowable error in triangulating the model to vastly reduce the size. Here, we reduce the model to 1/100th of it's raw size, while maintaining model quality. This can improve the performance when rendering 3D ggplots with `render_highquality()`, as well as improve performance on slower computers. This triangulation is powered by the `terrainmeshr` package.  

Here, we make a 3D ggplot out of glass, using a triangulated model and `render_highquality()`  

```{r}
par(mfrow = c(1, 1))
plot_gg(pp, width = 5, height = 4, scale = 300, 
        raytrace = F, windowsize = c(1200, 900), 
        fov = 70, zoom = 0.4, theta = 300, phi = 20, 
        max_error = 0.01, verbose = T)

Sys.sleep(0.2)

render_highquality(samples = 400, 
                   aperture = 30, 
                   light = F, 
                   ambient = T, 
                   focal_distance = 1700, 
                   obj_material = rayrender::dielectric(attenuation = c(1, 1, 0.3)/200), 
                   ground_material = rayrender::diffuse(checkercolor = "grey80", sigma = 90,
                                                        checkerperiod = 100))
```




