---
title: "linear model with R"
author: "yincy"
date: "3/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Notes for <Linear Models with R>**  
The emphasis of this text is on the practice of regression and analysis of variance.  

The objective is to learn what methods are available and more importantly, when they should be applied.  

# Chapter 1 Introduction  
## Before You Start  
Statistics starts with a problem, proceeds with the collection of data, continues with the data analysis and finshes with conclusions.  

**To formulate the problem correctly**  

1. Understand the physical background.  
2. Understand the objective.  
3. Make sure you know what the client wants.  
4. Put the problem into statistical terms.  


## Inital Data Analysis
This is a critical step that should always be performed. It looks simple but it is vital. You should make numerical summaries such as means, standard deviations (SDs), maximum and minimum, correlations and whatever else is appropriate to the specific dataset. Equally important are graphical summaries.  

```{r}
library(faraway)
data(pima)

pima
```

```{r}
summary(pima)
```

```{r}
pima$diastolic[pima$diastolic == 0] <- NA
pima$glucose[pima$glucose == 0] <- NA
pima$triceps[pima$triceps == 0] <- NA
pima$insulin[pima$insulin == 0] <- NA
pima$bmi[pima$bmi == 0] <- NA
pima$test <- factor(pima$test, levels = c(0, 1), labels = c("negative", "positive"))

summary(pima)
```

```{r}
hist(pima$diastolic)
```

```{r}
plot(density(pima$diastolic, na.rm = TRUE))
```

```{r}
plot(sort(pima$diastolic), pch = ".")
```

```{r}
plot(diabetes ~ diastolic, pima, pch = 19, cex = 0.5)
```

```{r}
plot(diabetes ~ test, pima)
```

```{r}
pairs(pima)
```

## When to Use Regression Analysis  
Regression analysis is used for explaining or modeling the relationship between a single variable Y, called the response, output or dependent variable; and one or more predictor, input, independent or explanatory variables, X~1~, …, X~p~.  

When p=1, it is called simple regression but when p>1 it is called multiple regression or sometimes multivariate regression.  

When there is more than one Y, then it is called multivariate multiple regression.  

The response must be a continuous variable, but the explanatory variables can be continuous, discrete or categorical.  

A regression with diastolic and bmi as X and diabetes as Y would be a multiple regression involving only quantitative variables.  

A regression with diastolic and test as X and bmi as Y would have one predictor that is quantitative and one that is qualitative.  

A regression with test as X and diastolic as Y involves just qualitative predictors called analysis of variance (ANOVA).  

A regression of test as Y on diastolic and bmi as predictors would involve a qualitative response. A logistic regression could be used.  


**Regression analyses have several possible objectives including**:  

1. Prediction of future observations.  
2. Assessment of the effect of, or relationship between, explanatory variables and the response.  
3. A general description of data structure.  


## History  
Legendre developed the method of least squares in 1805. Gauss claimed to heve developed the method a few years earlier and in 1809 showed that **least squares is the optimal solution when the errors are normally distributed**.  

```{r}
data("stat500")

stat500 <- data.frame(scale(stat500))
plot(final ~ midterm, stat500)
abline(0, 1)

g <- lm(final ~ midterm, stat500)
abline(coef(g), lty = 5)
```

```{r}
cor(stat500)
```

### Exercises  
1. 
```{r}
data("teengamb")
teengamb$sex <- factor(teengamb$sex, levels = c(0, 1), labels = c('male', 'female'))

summary(teengamb)
```

```{r}
plot(income ~ gamble, teengamb, pch = 19, cex = 0.5)
l <- lm(income ~ gamble, teengamb)
abline(coef(l))
```

The expenditure on gambling seems positively correlated with income  


```{r}
plot(gamble ~ sex, teengamb)
```

male spend more on gambling than female.  


2.  
```{r}
data("uswages")
uswages
```

```{r}
uswages$race <- factor(uswages$race, levels = c(0, 1), labels = c("White", "Black"))
uswages$smsa <- factor(uswages$smsa, levels = c(0, 1), labels = c("Metropolitan", "Non Metropolitan"))
uswages$ne <- factor(uswages$ne, levels = c(0, 1), labels = c("North East", "Non North East"))
uswages$mw <- factor(uswages$mw, levels = c(0, 1), labels = c("Midwest", "Non Midwest"))
uswages$we <- factor(uswages$we, levels = c(0, 1), labels = c("West", "Non West"))
uswages$so <- factor(uswages$so, levels = c(0, 1), labels = c("South", "Non South"))
uswages$pt <- factor(uswages$pt, levels = c(0, 1), labels = c("part time", "non pt"))
summary(uswages)
```

```{r}
par(mfrow = c(3, 3))
plot(wage ~ educ + exper + race + smsa + ne + mw + so + we, uswages, pch = 19)
```

3. 
```{r}
data("prostate")
prostate$svi <- factor(prostate$svi, levels = c(0, 1))
prostate$gleason <- factor(prostate$gleason, levels = unique(prostate$gleason))
prostate$pgg45 <- factor(prostate$pgg45, levels = unique(prostate$pgg45))

summary(prostate)
```

```{r}
plot(lcavol ~ svi, prostate)
```

4.  
```{r}
data(sat)

sat

summary(sat)
```

```{r}
par(mfrow = c(2, 2))
plot(expend ~ salary, sat)
l <- lm(expend ~ salary, sat)
abline(l, lty = 5)

plot(ratio ~ salary, sat)

plot(expend ~ takers, sat)

plot(expend ~ total, sat)
```


5.  
```{r}
data(divusa)
divusa
```

```{r}
summary(divusa[, -1])
```

```{r}
par(mfrow = c(2, 2))
plot(divorce ~ unemployed, divusa)
plot(divorce ~ year, divusa)
plot(divorce ~ femlab, divusa)
plot(divorce ~ marriage, divusa)
```


# Chapter 2 Estimation  
Suppose we want to model the response Y in terms of three predictors, X~1~, X~2~ and X~3~. One very general form for the model would be:  

$$Y = f(X_{1}, X_{2}, X_{3}) + e$$

Where `f` is some unknown function and `e` is the error in this representation. `e` is additive in this instance, but could enter in some more general form. Still, if we assume that `f` is a smooth, continuous function, that still leaves a very wide range of possibilities. Even with just three predictors, we typically will not have enough data try to estimate `f` directly. So we usually have to assume that it has some more restricted form, perhaps linear as in:  

$$Y = b_{0} + b_{1}X_{1} + b_{2}X_{2} + b_{3}X_{3} + e$$

Where b~i~, i=0, 1, 2, 3 are unknown parameters. b~0~ is called the intercept term.  

Thus the problem is reduced to the estimation of four parameters rather than the infinite dimensional `f`.  


> Linear models seem rather restrictive, but because the predictors can be transformed and combined in any way, they are actually very flexible.  


$$Data = Systematic Structure + Random Variation$$

$$N dimensions = P dimensions + (N - P) dimensions$$

The estimation of `b` can also be considered from a nongeometrical point of view. We might define the best estimate of b as the one which minimizes the sum of the squared errors.  

```{r}
library(faraway)
data(gala)
gala
```
 
```{r}
mdl <- lm(Species ~ Area + Elevation + Nearest + Scruz + Adjacent, data = gala)

summary(mdl)
```

```{r}
x <- model.matrix( ~ Area + Elevation + Nearest + Scruz + Adjacent, data = gala)
y <- gala$Species

xtxi <- solve(t(x) %*% x) # get the reverse of (t(x) %*% x)

# we can get `b` directly using (X^T^X)~-1~X~T~y

xtxi %*% t(x) %*% y
```

a better way to compute `b`

```{r}
solve(crossprod(x, x), crossprod(x, y))
```

`crossprod(x, y)` compute x^T^y  


```{r}
names(mdl)
```

```{r}
mdls <- summary(mdl)
names(mdls)
```

```{r}
mdls$sigma

mdls$r.squared
```

```{r}
fitted(mdl)
```





