---
title: "Advanced-Object-oriented-Programming-in-R"
author: "yincy"
date: "3/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Chapter 1 Classes and Generic Functions  
R’s approach to object-oriented programming is through `generic functions` and `classes`.  

## Generic Functions  
The term generic functions refers to functions that can be used on more than one data type.  

Generic functions are functions that work differently on different types of objects. They are therefore also known as `polymorphic functions`.  

```{r}
top <- function(stack)UseMethod(generic = "top")
pop <- function(stack)UseMethod(generic = "pop")
push <- function(stack, element)UseMethod(generic = "push")
is_empty <- function(stack)UseMethod(generic = "is_empty")
```

default method  
```{r}
top.default <- function(stack) .NotYetImplemented()
pop.default <- function(stack) .NotYetImplemented()
push.default <- function(stack, element) .NotYetImplemented()
is_empty.default <- function(stack).NotYetImplemented()
```


## Classes  
In the S3 system, you create a class, and assign a class to an object, just by setting an attribute on the object. The name of the class is all that defines it, so there is no real type checking involved. **Any object can have an attribute called “class” and any string can be the name of a class**.  

```{r}
empty_vector_stack <- function(){
    stack <- numeric()
    class(stack) <- "vector_stack"
    return(stack)
}
```


```{r}
stack <- empty_vector_stack()
stack
```

```{r}
attributes(stack)
```

```{r}
class(stack)
```

```{r}
top.vector_stack <- function(stack)stack[1]

pop.vector_stack <- function(stack){
    new_satck <- stack[-1]
    class(new_stack) <- "vector_stack"
    return(new_satck)
}

push.vector_stack <- function(element, stack){
    new_stack <- c(element, stack)
    class(new_stack) <- "vector_stack"
    return(new_stack)
}

is_empty.vector_stack <- function(stack){
    return(length(stack) == 0)
}
```

```{r}
stack <- push(stack, 1)
stack <- push(stack, 2)
stack <- push(stack, 3)
stack
```

```{r}
while(is_empty(stack)){
    stack <- pop(stack)
}
stack
```


```{r}
make_vector_stack <- function(elements){
    structure(elements, class = "vector_stack")
}

empty_vector_stack <- function(){
    make_vector_stack(vector("numeric"))
}

top.vector_stack <- function(stack){
    stack[1]
}

pop.vector_stack <- function(stack){
    make_vector_stack(stack[-1])
}

push.vector_stack <- function(stack, element){
    make_vector_stack(c(element, stack))
}

is_empty.vector_stack <- function(stack) length(stack) == 0
```


## Polymorphism in Action  
```{r}
make_list_node <- function(head, tail){
    list(head = head, tail = tail)
}

make_list_stack <- function(elements){
    structure(list(elements = elements), class = "list_stack")
}
empty_list_stack <- function()make_list_stack(NULL)

top.list_stack <- function(stack)stack$element$head
pop.list_stack <- function(stack) make_list_stack(stack$element$tail)
push.list_stack <- function(stack, element){
    make_list_stack(make_list_node(element, stack$elements))
}
is_empty.list_stack <- function(stack) is.null(stack$elements)
```


```{r}
stack <- empty_list_stack()
stack <- push(stack, 1)
stack <- push(stack, 2)
stack <- push(stack, 3)
stack
```






