---
title: "reticulate"
author: "yincy"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_engines$set(python = reticulate::eng_python)
```

# R Interface to Python  
The `reticulate` package provides a comprehensive set of tools for interoperability between Python and R. The package includes facilities for:  

- Calling Python from R in a variety of ways including R Markdown, sourcing Python scripts, importing Python modules, and using Python interactively within an R session.  

- Translation between R and Python objects (for example, between R and Pandas data frames, or between R matrices and NumPy arrays).  

- Flexible binding to different versions of Python including virtual environments and Conda environments.   

Reticulate embeds a Python session within your R session, enabling seamless, high-performance interoperability. If you are an R developer that uses Python for some of your work or a member of data science team that uses both languages, reticulate can dramatically streamline your workflow!  

## Getting Started  
```{r, include=FALSE}
# BiocManager::install("reticulate")
library(reticulate)
#use_python(python = "/home/yincy/miniconda3/envs/py37/bin/python")
```

### Python version  
By default, reticulate uses the version of python found on your `PATH`.  
The `use_python` function enables you to specify an alternative version.  
```
use_python(python = "/path/to/a/python")
```

The `use_virtualenv` and `use_condaenv` functions enables you to specify version of python in virtual or conda environments.  


### Python packages 
You can install any required Python packages using standard shell tools like pip and conda. Alternately, reticulate includes a set of functions for managing and installing packages within virtualenvs and Conda environments.   


### Calling Python  
there are a varity of ways to integrate Python code into your R projects  

- Python in R markdown: A new Python language engine for R Markdown that supports bi-directional communication between R and Python (R chunks can access Python objects and vice-versa).  

- Importing Python modules: The `import` function enables you to import any python module and call its functions directly from R.  

- Sourcing Python scripts: The `source_python()` function enables you to source a Python script the same way you would source() an R script (Python functions and objects defined within the script become directly available to the R session).  

- Python REPL: The `repl_python()` function creates an interactive Python console within R. Objects you create within Python are available to your R session (and vice-versa).  


#### Python in RMarkdown  
The reticulate package includes a Python engine for R Markdown with the following features  

1. Run Python chunks in a single Python session embedded within your R session (shared variables/state between Python chunks).  

2. Printing of Python output, including graphical output from matplotlib.  

3. Access to objects created within Python chunks from R using the py object (e.g. py$x would access an x variable created within Python from R).  

4. Access to objects created within R chunks from Python using the r object (e.g. r.x would access to x variable created within R from Python).  

Built in conversion for many Python object types is provided, including NumPy arrays and Pandas data frames. From example, you can use Pandas to read and manipulate data then easily plot the Pandas data frame using ggplot2:  
```{r}

```






















