---
title: "Fuzzy_Clustering"
author: "yincy"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The **fuzzy clustering** is considered as soft clustering, in which each element has a probability of belonging to each cluster. In other words, each element has a set of membership coe?cients corresponding to the degree of being in a given cluster.  

This is di?erent from k-means and k-medoid clustering, where each object is a?ected exactly to one cluster. K-means and k-medoids clustering are known as hard or non-fuzzy clustering.  

In fuzzy clustering, points close to the center of a cluster, may be in the cluster to a higher degree than points in the edge of a cluster. The degree, to which an element belongs to a given cluster, is a numerical value varying from 0 to 1.  

The **fuzzy c-means** (FCM) algorithm is one of the most widely used fuzzy clustering algorithms. The centroid of a cluster is calculated as the mean of all points, weighted by their degree of belonging to the cluster.  

# example
```{r}
library(cluster)
library(factoextra)
library(magrittr)

df <- scale(USArrests)
res.fanny <- fanny(x = df, k = 2)
res.fanny$membership %>% head

fviz_cluster(res.fanny, 
             ellipse.type = "norm", 
             repel = T, 
             palette = "jco", 
             ggtheme = theme_classic(), 
             legend = "right")
```

```{r}
fviz_silhouette(res.fanny, 
                palette = "jco", 
                ggtheme = theme_classic())
```














