---
title: "Model-Based-Clustering"
author: "yincy"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**model-based clustering**, consider the data as coming from a distribution that is mixture of two or more clusters. Unlike k-means, the model-based clustering uses a soft assignment, where each data point has a probability of belonging to each cluster.  


# Concept of model-based clustering
In model-based clustering, the data is considered as coming from a mixture of density. Each component k is modeled by the normal or Gaussian distribution which is characterized by the parameters:  

- $u_{k}$: mean vector.  

- $\sum_{k}$: covariance matrix.  

- An associated probability in the mixture. Each point has a probability of belonging to each cluster.  

```{r}
library(MASS)
data("geyser")
library(ggpubr)
library(ggplot2)

ggplot(geyser, aes(duration, waiting)) +
    geom_point(size = 2) +
    geom_density_2d() +
    theme_classic2() +
    theme(aspect.ratio = 1)
```

# example
```{r}
library(mclust)
data("diabetes")
diabetes %>% head

df <- scale(diabetes[, -1])
mc <- Mclust(df)
summary(mc)
mc$modelName # optimal selected model
mc$G # optimal number of cluster
mc$z %>% head # probality to belong to a given cluster
mc$classification %>% head # cluster assignment of each observation
```

```{r}
library(factoextra)
fviz_mclust(mc, 
            what = "BIC", 
            palette = "jco")
```

The first identifier refers to volume, the second to shape and the third to orientation.  

- `E` stands for 'equal'  
- `V` stands for 'variable'  
- `I` stands for 'coordinate axes'  


```{r}
fviz_mclust(mc, 
            what = "classification", 
            geom = "point", 
            pointsize = 1.5, 
            palette = "jco")
```

```{r, message=FALSE, warning=FALSE}
fviz_mclust(mc, 
            what = "uncertainty", 
            palette = "jco")
```



