---
title: "ggraph circular packing"
author: "yincy"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(ggraph)
library(igraph)
library(tidyverse)
library(viridis)
```


# Hiding the first level for better styling
To hide a specific layer, just spcify a color equal to the background color in the `scale_color_manual()` and `scale_fill_manual()` function.  

```{r}
edges <- flare$edges
vertices <- flare$vertices

mygraph <- graph_from_data_frame(d = edges, vertices = vertices)

# hide the first level

ggraph(mygraph, layout = "circlepack", weight = size) +
    geom_node_circle(aes(fill = as.factor(depth), color = as.factor(depth))) +
    scale_fill_manual(values = c("0" = "white", 
                                 "1" = viridis(4)[1], 
                                 "2" = viridis(4)[2], 
                                 "3" = viridis(4)[3], 
                                 "4" = viridis(4)[4])) +
    scale_color_manual(values = c("0" = "white", 
                                  "1" = "black", 
                                  "2" = "black", 
                                  "3" = "black", 
                                  "4" = "black")) +
    theme_void() +
    theme(legend.position = "none")
```

```{r}
# hide 2 first levels

ggraph(mygraph, layout = "circlepack", weight = size) +
    geom_node_circle(aes(fill = as.factor(depth), color = as.factor(depth))) +
    scale_fill_manual(values = c("0" = "white", 
                                 "1" = "white", 
                                 "2" = magma(4)[2], 
                                 "3" = magma(4)[3], 
                                 "4" = magma(4)[4])) +
    scale_color_manual(values = c("0" = "white", 
                                  "1" = "white", 
                                  "2" = "black", 
                                  "3" = "black", 
                                  "4" = "black")) +
    theme_void() +
    theme(legend.position = "none")
```

# add labels to a specific level of the hierarchy
To solve this issue, the trickiest part is to determine the level of each node in the edge list data frame. Fortunately, the `data.tree` library is here to help with its `FromDataFrameNetwork()` function. It allows to isolate the level of each node, making it a breeze to select the labels to display.  

```{r, message=FALSE, warning=FALSE}
library(data.tree)

# rebuild the data
edges <- flare$edges
vertices <- flare$vertices

# transform it in a 'tree' format
tree <- FromDataFrameNetwork(edges)

# get the level of each node, and add it to the inital data frame
mylevels <- data.frame(name = tree$Get("name"), level = tree$Get("level"))
vertices <- vertices %>% left_join(., mylevels, by = c("name" = "name"))

# add label for level 1 and 2
vertices <- vertices %>% 
    mutate(new_label = ifelse(level == 2, shortName, NA))

mygraph <- graph_from_data_frame(edges, vertices = vertices)

ggraph(mygraph, layout = "circlepack", weight = size) +
    geom_node_circle(aes(fill = as.factor(depth), color = as.factor(depth))) +
    scale_fill_manual(values = c("0" = "white", 
                                 "1" = viridis(4)[1], 
                                 "2" = viridis(4)[2], 
                                 "3" = viridis(4)[3], 
                                 "4"=viridis(4)[4])) +
    scale_color_manual(values = c("0" = "white", 
                                  "1" = "black", 
                                  "2" = "black", 
                                  "3" = "black", 
                                  "4"="black")) +
    geom_node_label(aes(label = new_label), size = 2.5, nudge_x = 5, nudge_y = 5) +
    theme_void() +
    theme(legend.position = "none", plot.margin = unit(rep(0, 4), "cm"))
```










