---
title: "pipeR"
author: "yincy"
date: "10/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting started  
`pipeR` is an Rpackage that helps you better organize your code in pipeline built with `%》>%`, `Pipe()` or `pipeline()` which is much easier to read, wirte and maintain.  

```{r}
sample_mtcars <- sample(mtcars$mpg, 10000, replace = T)
density_mtcars <- density(sample_mtcars, kernel = "gaussian")
plot(density_mtcars, col = "red", main = "density of mpg (bootstrap)")
```

```{r}
mtcars$mpg %>>% 
    sample(size = 10000, replace = T) %>>% 
    density(kernel = "guassian") %>>% 
    plot(col = "red", main = "density of mpg")
```

```{r}
mtcars %>>% 
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.05)) %>% 
    (mpg) %>>% 
    sample(size = 10000, replace = T) %>>% 
    density(kernel = "guassian") %>>% 
    plot(col = "red", main = "density of mpg")
```


**Design priciples**  

- define a set of syntax rather than symbols  
- syntax should be simple, clear and intuitive  
- give user full control of the piping behavior  

**Pipe operator**  
- to first argument of a function  
- to `.` in an expression  
- by formula to avoid ambiguity in symbol names.  
- to an expression only for its side effect  
- to an expression and save the intermediate results  
- to a symbol in order to extract that element  

Basically, the behavior of the operator is fully determined by the syntax of the expression that follows the operator.  


# Pipe operator  
## pipe to first argument  
```{r, message=FALSE, warning=F}
rnorm(100, 0) %>>% 
    log %>>% 
    diff %>>% 
    sample(size = 10000, replace = T) %>>% 
    summary()
```

```{r}
mtcars$mpg %>% 
    plot(col = "red", main = sprintf("number of points: %d", length(.)))
```


There are situations where one calls a function in a namespace with ::. In this case, the call must end up with parentheses with or without parameters.  

```{r}
mtcars$mpg %>>% 
    stats::median()

mtcars$mpg %>>% 
    graphics::plot(col = "red")
```

```{r}
functions <- list(average = function(x) mean(x))

mtcars$mpg %>>% functions$average()
mtcars$mpg %>>% functions[["average"]]() # need parenthese
```

```{r}
mtcars %>>% 
    subset(mpg <= quantile(mpg, 0.95), c(mpg, wt)) %>>% 
    summary()
```

One important thing to notice here is that pipeR does not support lazy evaluation on left value, that is, the left value will be evaluated immediately which cannot be substituted by the function on the right.  

```{r}
10000 %>>% 
    replicate(rnorm(10000)) %>>% 
    system.time()
```

```{r}
system.time(replicate(10000, rnorm(1000)))
```

```{r}
mtcars %>>% 
    lm(mpg ~ cyl + wt)
```

```{r}
mtcars %>>% 
    lm(formula = mpg ~ cyl + wt)
```

## pipe to dot  
```{r}
mtcars %>>% 
    (lm(mpg ~ cyl + wt, data = .))
```

```{r}
mtcars %>>% 
    {lm(mpg ~ cyl + wt, data = .)}
```

**The difference between `{}` and `()` used above is**:  

- `{}` accepts more than one expression within the braces and its value is determined by the last one; but `()` accepts only one expression.  

- `{}` has only one feature: pipe to `.` in the enclosed expression while `()` has more features.  

```{r}
mtcars %>>% {
    model <- lm(mpg ~ wt + cyl, data = .)
    summ <- summary(model)
    summ[c('r.squared', 'adj.r.squared')]
}
```

```{r}
mtcars %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    summary %>>% {
    if(.$r.squared >= 0.8){
        return("Model A")
    }
    cat("Model B should be considered.\n")
    "Model B"
}
```

```{r}
mtcars %>>% {
    par(mfrow = c(1, 2))
.$mpg %>>% plot(col = "red", main = "mtcars (mpg)")
.$mpg %>>% hist(main = "distribution")
}
```

```{r}
mtcars %>>% {
    par(mfrow = c(1, 2))
.$mpg %>>% plot(col = "red", main = sprintf("mtcars (mpg: %d)", length(.)))
.$mpg %>>% hist(main = "distribution")
}
```


### Creating closure  
```{r}
density_plot <- mtcars$mpg %>>% 
    sample(size = 10000, replace = T) %>>% 
    (function(kernel){
        . %>>% 
            density(kernel = kernel) %>% 
            plot(main = sprintf("%s kernel", kernel))
    })
```

```{r}
par(mfrow = c(1,3))
density_plot("guassian")
density_plot("rectanglar")
density_plot("trianglar")
```


```{r}
1:10 %>>% (function(x, pow) x^pow)(2)
```

```{r}
mtcars %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    (function(model, warn_level){
        warning("r.squared is too low", call. = FALSE)
        model
    })(0.9) %>>%
    coef
```

## Pipe by formula  
```{r}
mtcars %>>% 
    (lm(mpg ~., data = .))
```

the above code an be rewritten using lambda expression like 

```{r}
mtcars %>>% 
    (df ~ lm(mpg ~., data = df))
```

the formula tells `%>>%` to use `df` to represent `mtcars` so that the expression of linear model fit won't look ambiguous any more.  

```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (x ~ lm(mpg ~., data = x))
```

```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (x ~ {
        summ <- lm(mpg ~., data = x) %>>% 
            summary
        list(n = nrow(x), r.squared = summ$r.squared)
    })
```


## Pipe for side effect  
In a pipeline, one may be interested not only in the final outcome but sometimes also in intermediate results. In other words, as the operator pipes the value forward for streamlined processing, one may also want to save the intermediate result for further use, visualize the intermediate values for better understanding, or print some message to indicate the progress.  

```{r}
z <- plot(mtcars$mpg)
z
```

It can be clearly seen that `plot()` does two things:  

1. Draw a scatter image of `mtcars$mpg`  
2. return `NULL` value  

The first thing is often considered as side-effect because **it is an influence the function has on the outside environment**: it changes what to show in the graphics area. However, when we are building a pipeline and want to plot with some intermediate results, we only want the side effect and have to avoid the returned value, in this case, NULL, which is not the desired data we would like to further process.  


## Syntax for side effect  
```{r}
mtcars %>>% 
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95), c(mpg, cyl, wt)) %>>% 
    (~ cat("rows: ", nrow(.), "\n")) %>>% 
    summary()
```

```{r, eval=FALSE}
mtcars %>>% 
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95)) %>>% 
    cat("rows: ", nrow(.), "\n") %>>% 
    summary()
```


```{r}
mtcars %>>% 
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95)) %>>% 
    (~ plot(mpg ~ wt, data = .)) %>>% 
    lm(formula = mpg ~ wt) %>>% 
    summary()
```

## Lambda expression  
```{r}
mtcars %>>% 
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95), c(mpg, cyl ,wt)) %>>% 
    (~ filtered ~ cat("rows:", nrow(filtered), "\n\n")) %>>% 
    summary()
```

## String
```{r}
mtcars %>>%
    "Running linear model" %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    "summarizing the model" %>>% 
    summary() %>>% 
    coef()
```

```{r}
data(diamonds, package = "ggplot2")

diamonds %>>% 
    "filtering data ..." %>>% 
    (~ Sys.sleep(2)) %>>% 
    filter(carat >= quantile(carat, 0.05) & carat <= quantile(carat, 0.95)) %>>% 
    select(carat, price, depth) %>>% 
    "estimating linear model ..." %>>% 
    (~ Sys.sleep(2)) %>>% 
    lm(formula = carat ~ price + depth) %>>% 
    "summarizing model ..." %>>% 
    (~ Sys.sleep(2)) %>>% 
    summary()
```

## Question mark  
```{r}
mtcars %>>% 
    subset(vs == 1, c(mpg, wt, cyl)) %>>% 
    (? nrow(.)) %>>% 
    (? data ~ ncol(data)) %>>% 
    summary()
```

```{r}
mtcars %>>% 
    ("Raw data rows:" ? nrow(.)) %>>% 
    subset(vs == 1, c(mpg, cyl, wt)) %>>% 
    ("Filtered data rows:" ? nrow(.)) %>>% 
    lm(formula = mpg ~ wt + cyl)
```


## Stopping  
```{r}
mtcars %>>% 
    subset(vs == 1, c(mpg, wt, cyl)) %>>% 
    lm(formula = mpg ~ cyl + wt) %>>% 
    (~ stopifnot(summary(.)$r.squared >= 0.5)) %>>% 
    predict(newdata = list(cyl = 4.5, wt = 3.0))
```

```{r,eval=FALSE}
mtcars %>>% 
    subset(vs == 1, c(mpg, wt, cyl)) %>>% 
    lm(formula = mpg ~ cyl + wt) %>>% 
    (~ stopifnot(summary(.)$r.squared >= 0.8)) %>>% 
    predict(newdata = list(cyl = 4.5, wt = 3.0))
```


## Debugging  
```{r, eval=F}
mtcars %>>% 
    subset(vs == 1, c(mpg, wt, cyl)) %>>% 
    (~ message("debug1")) %>>% 
    (~ browser()) %>>% 
    lm(formula = mpg ~ cyl + wt) %>>% 
    summary()
```

Since `browser()` supports conditional breakpoint with `expr =` being set, the breakpoint can be smart to examine the value of the expression and then decide whether or not to pause.  

```{r}
mtcars %>>% 
    subset(vs == 1, c(mpg, cyl, wt)) %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    summary() %>>% 
    (~ browser(expr = .$r.squared < 0.5))

# if the R^2 of the inear model is less than 0.8, the evaluation will pause and wait for browser.
```


# Pipe with assignment  
```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (~ sub_mtcars) %>>% 
    (~ Sys.sleep(2)) %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    (~ lm_mtcars) %>>% 
    (~ Sys.sleep(2)) %>>% 
    summary()
```

```{r}
sub_mtcars %>% head()
```

```{r}
lm_mtcars
```

```{r}
ls.str() # inspect the environment variable and it's structure
```

```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (~ summ = summary(.)) %>>% # side-effect assignment
    lm(formula = mpg ~ wt + cyl)
```

```{r}
summ
```

```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (~ summ = df ~ summary(df)) %>>% # side-effect assignment
    lm(formula = mpg ~ wt + cyl)
```

```{r}
summ
```

```{r}
mtcars %>>% 
    subset(select = c(mpg, wt, cyl)) %>>% 
    (~ summ = df ~ summary(df)) %>>% 
    (model = lm(mpg ~ wt + cyl, data = .)) # pipe and assign
```

```{r}
summ
model
```


```{r}
mtcars %>>% 
    (~ summary(.) -> summ) %>>% 
    subset(cyl == 6) %>>% 
    ("number of rows after selection: " ? nrow(.))
```

```{r}
summ
```


The `(~ expression -> symbol)` and `(~ symbol <- expression)` syntax work for side-effect assignment, and `(expression -> symbol)` and `(symbol <- expression)` work for piping with assignment.  


```{r}
mtcars %>>% 
    (~ summary(.) -> summ) %>>%  # side-effect assignment
    (lm(formula = mpg ~ wt + cyl, data = .) -> lm_mtcars) %>>% # pipe assignment 
    summary()
```

```{r}
summ
```

```{r}
lm_mtcars
```

```{r}
results <- list()
mtcars %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    (~ results$mtcars = . %>>% summary %>>% (r.squared)) # side-effects assignment

results
```

```{r}
iris %>>% 
    lm(formula = Sepal.Length ~ Sepal.Width) %>>% 
    (~ results$iris = . %>>% summary %>>% (r.squared))

results
```

```{r}
results <- numeric()
rnorm(100) %>>% 
    (~ mean(.) -> results["mean"]) %>>% 
    (~ median(.) -> results['median']) %>>%
    summary

results
```

```{r}
numbers <- 1:5

letters %>>% 
    sample(length(numbers)) %>>% 
    (~ . -> names(numbers))

numbers
```


# extract element  
The syntax `x %>>% (y)` indicates to extract the element named `y` from object `x` where `y` must be a valid symbol name and `x` can be a vector, list, environment or anything else that supports `[[]]`. It also works with S4 object and able to extract a slot from it in exactly the same way.  

```{r}
mtcars %>>% 
    (lm(mpg ~ wt + cyl, data = .)) %>>% 
    summary %>>% 
    (r.squared)
```

```{r}
list(a = 1, b = 2) %>>% (a)
```

```{r}
c(a = 1, b = 2) %>>% (a)
```

```{r}
env <- new.env()
env$a <-  1
env$b <-  2

env %>>% (a) 
```

```{r}
set.seed(0)
x <- cumsum(rnorm(150))
y <- 2 * x + rnorm(150) * 0.5

matplot(cbind(x, y), type = "l")
```

```{r}
lm(y ~ x + 0) %>>% 
    residuals %>>%
    fUnitRoots::adfTest(type = "nc") %>>% # S4 object
    (~ testobj) %>>% # save test object to environment
    ("Which class of this object?" ? class(.)) %>>% 
    (test) %>>%
    (statistic)
```


Since the syntax only accepts symbol name, to evaluate an expression with the piped object being the frame if it is a list or environment, using `with()` or `within()` can be helpful.  

```{r}
list(a = 1, b = 2) %>>%
    with(a + 2 * b)
```


# Pipe function  
`Pipe()` function is the light-weight version of Pipe operator yet has more features.  

1. Support all features of `%>>%` but use `$` for command chaining.  
2. Support direct subsetting, extraction, element assignment.  

Basically, `Pipe()` creates a `Pipe` object for which `$` is specically defined to perform first-argument piping, and `.(...)` is defined to act like `x %>>% (...)`. Pipe object also supports subsetting (`[]`), extraction (`[[]]`), and their corresponding assignment operations.  

```{r}
mtcars %>>% 
    lm(formula = mpg ~ wt + cyl) %>>% 
    summary()
```

can be rewritten like  
```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)$ 
    summary()
```


## Pipe to first-argument  
```{r}
set.seed(123)

Pipe(rnorm(100, mean = 10))$
    log()$
    .(~ Sys.sleep(2))$
    diff()$
    sample(size = 10000, replace = T)$
    summary() # eventially return an Pipe object
```


```{r}
Pipe(rnorm(100, mean = 10))$
    log()$
    diff()$
    sample(size = 10000, replace = T)$
    summary()$
    value # the output object is no longer an Pipe object
```

```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)
```

```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)[]
```


```{r}
Pipe(mtcars$mpg)$
    plot(col = "red", main = sprintf("mpg (%d obs.)", length(.)))
```

```{r}
pmtcars <- Pipe(mtcars)$
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95))

pmtcars$
    lm(formula = mpg ~ wt + cyl)
```

```{r}
pmtcars$
    lm(formula = mpg ~ wt + cyl + qsec)$
    summary()$
    coef()
```


### Creating partial function  
```{r}
density_plot <- Pipe(mtcars$mpg)$
    sample(size = 10000, replace = T)$
    density(kernel = "gaussian")$
    plot
```


```{r}
par(mfrow = c(1, 2))
density_plot(col = "blue", main = "blue points")
density_plot(col = "grey", main = "gray circles", type = "o")
```


## Dot function  
### Pipe to dot  
```{r}
Pipe(mtcars)$
    .(lm(mpg ~ cyl + wt, data = .))
```


### Pipe by formula  
```{r}
Pipe(mtcars)$
    .(df ~ lm(mpg ~ cyl + wt, data = df))
```


### Pipe for side effect  
```{r}
Pipe(mtcars)$
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95), c(mpg, cyl, wt))$
    .(~ cat('rows:', nrow(.), "\n\n"))$
    lm(formula  = mpg ~ .)$
    summary()$
    coef()
```

#### Question mark  
```{r}
Pipe(mtcars)$
    subset(vs == 1, c(mpg, cyl, wt))$
    .("number of rows after filtering:"? nrow(.))$
    .("number of columns after filtering:"? data ~ ncol(data))$
    summary()
```

```{r}
mtcars %>>% 
    subset(vs == 1, c(mpg, cyl, wt)) %>>% 
    (~ browser()) %>>% 
    lm(formula = mpg ~ cyl + wt) %>>% 
    summary()
```


### Pipe with assignment  
```{r}
pmodel <- Pipe(mtcars)$
    subset(select = c(mpg, wt, cyl))$
    .(~ sub_mtcars)$
    lm(formula = mpg ~ wt + cyl)

summary(pmodel)
```

```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)$
    .(~ summary(.) -> summ)$ 
    coef() # works on the lm results
```

```{r}
summ
```


```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)$
    .(summary(.) -> summ)$
    coef() # works on the summary results
```

### Extract element  
```{r}
Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)$
    summary()$
    .(r.squared)
```

### Dot functions as closure  
Note that `.()` is only defined within a `Pipe` object and is not an exportedd function in PipeR, you cannot directly call `.()` outside a `Pipe()`.  

In fact, if you are familiar with functional programming, it is called a *closure*, which is basically defined as a function returned by a function. However, you can save `.()` in a `Pipe` object to a symbol as a shortcut to call it elsewhere.  

```{r}
model <- Pipe(mtcars)$
    lm(formula = mpg ~ wt + cyl)$
    .

model()
```

```{r}
model(coefficients) # pipe extract
```

```{r}
model()$summary()$coef() # pipe to first argument
```


## Subsettting  
### Using `[]`  
```{r}
Pipe(list(a = 1, b = 2, c = 3))[c("a", "b")]
```

```{r}
Pipe(mtcars)[c("mpg", "cyl", "wt")]$
    head()
```

```{r}
library(data.table)
dt <- data.table(id = 1:6, x = rnorm(6), y = rep(letters[1:3]), key = "id")
dt
```

```{r}
Pipe(dt)[1:3]
```

```{r}
Pipe(dt)[J(3)]
```

```{r}
Pipe(dt)[, sum(x), by = list(y)]
```

```{r}
Pipe(dt)[, z := x^2 + 1] # reference mutate
```

```{r}
pmtcars <- Pipe(mtcars)$
    .(~ rownames(.) -> row_names)$
    as.data.table()[, name := row_names]$
    setkey(name)
```

```{r}
row_names
```


```{r}
pmtcars[mpg >= quantile(mpg, 0.05)]$
    lm(formula = mpg ~ wt + cyl)$
    summary()$
    coef()
```


```{r}
Pipe(mtcars$mpg)[-length(.)]
```

```{r}
Pipe(mtcars)[['mpg']]$
    summary()
```

```{r}
Pipe(mtcars)$ 
    .(mpg)$
    summary()
```

```{r}
Pipe(mtcars)$
    with(summary(mpg))
```


### Assigning values to elements  
```{r}
lst <- Pipe(list(a = 1, b  =2))
lst
```

```{r}
lst$a <- 2
lst
```

```{r}
lst[['b']] <- NULL
lst
```

```{r}
lst$c <- 1:3
lst
```

```{r}
lst[c("a", "c")] <- list(1:3, 2:5)
lst
```


# Pipeline
This function evaluates both argument-based and expression-based pipelines. What make it special is that it does not require using pipe operator `%>>%`, nor does it create a `Pipe` object as a wrapper. Using `pipeline()`, you will be able to write cleaner pipeline expressions with less noise.  

In other words, pipeline() implements two similar operator-free mechanisms for building a pipeline. The good thing is that you don't have to sacrifice any feature to use it, that is, all features of %>>% are supported in pipeline() and some are even made easier to use.  

## Argument-based pipeline  
```{r}
mtcars$mpg %>>% 
    sample(size = 10000, replace = T) %>>% 
    density(kernel = "gaussian", bw = 5) %>>% 
    plot(col = "red", main = "density of mpg (bootstrap)")
```

rewrite with `pipeline()`  
```{r}
pipeline(mtcars$mpg, 
         sample(size = 10000, replace = T), 
         density(kernel = "gaussian", bw = 5), 
         plot(col = "red", main = "density of mpg (bootstrap)"))
```

Internally, `pipeline()` translate such a function call into expressions connected by `%>>%`.  

```{r}
pipeline(mtcars$mpg, 
         ("The number of samples: "? length(.)), 
         sample(size = 10000, replace = T), 
         (~ mtcars_sample), # side effect assignment 
         density(kernel = "gaussian", bw = 5), 
         plot(col = "red", main = "density of mpg (bootstrap)"))
```

## expression-based pipeline  
```{r}
mtcars$mpg %>>% 
    sample(size = 10000, replace = T) %>>% 
    density(kernel = "gaussian", bw = 5) %>>% 
    plot(col = "red", main = "density of mpg (bootstrap)")
```


```{r}
pipeline({
    mtcars$mpg
    sample(size = 10000, replace = T)
    density(kernel = "gaussian", bw = 5)
    plot(col = "red", main = "density of mpg (bootstrap)")
})
```

```{r}
pipeline({
    mtcars$mpg
    sample(size = 10000, replace = T)
    "number of samples: "? length(.)
    ~ mtcars_sample
    density(kernel = "gaussian", bw = 5)
    plot(col = "red", main = "density of mpg (bootstrap)")
})
```

```{r}
pipeline({
    mtcars
    "Total number of records" ? nrow(.)
    subset(mpg >= quantile(mpg, 0.05) & mpg <= quantile(mpg, 0.95))
    "Qualified number of records" ? nrow(.)
    lm(formula = mpg ~ wt + cyl)
    summary()
    "R Squared" ? .$r.squared
    coef
})
```





