---
title: "janitor"
author: "yincy"
date: "4/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Major functions
## clean
### clean data.frame names with `clean_names()`
It works in a `%>%` pipeline, and handles problematic variable names, especially those that are so well-preserved by `readxl::read_excel()` and `readr::read_csv()`.

- Parses letter cases and separators to a consistent format.  
    + Default is to snake_case, but other cases like camelCase are available  
    
- Handles special characters and spaces  
- Appends numbers to duplicated names  
- Converts "%" to "percent" and "#" to "number" to retain meaning  
- Spacing (or lack thereof) around numbers is preserved  

```{r}
test_df <- as.data.frame(matrix(ncol = 6))
names(test_df) <- c("firstName", 
                    "abc@!*", 
                    "% successful (2009)", 
                    "REPEAT VALUE", 
                    "REPEAT VALUE", 
                    "")
```

Clean the variable names, returning a data.frame:  

```{r}
library(janitor)

test_df %>% 
    clean_names
```

Compare to what base R produces

```{r}
make.names(names(test_df))
```

This function is powered by the underlying exported function `make_clean_names()`, which accepts and returns a character vector of names (see below). This allows for cleaning the names of any object, not just a data.frame. clean_names() is retained for its convenience in piped workflows, and can be called on an `sf` simple features object or a `tbl_graph` `tidygraph` object in addition to a data.frame.  


# Do those data.frames actually contain the same columns?
## Check with `compare_df_cols()`
For cases when you are given a set of data files that should be identical, and you wish to read and combine them for analysis. But then `dplyr::bind_rows()` or `rbind()` fails, because of different columns or because the column classes don’t match across `data.frames`.  

`compare_df_cols()` takes unquoted names of `data.frames / tibbles`, or a list of `data.frames`, and returns a summary of how they compare. See what the column types are, which are missing or present in the different inputs, and how column types differ.

```{r}
df1 <- data.frame(a = 1:2, 
                  b = c("big", "small"))
df2 <- data.frame(a = 10:12, 
                  b = c("medium", "small", "big"), 
                  c = 0, 
                  stringsAsFactors = T)
df3 <- df1 %>% 
    dplyr::mutate(b = as.character(b))
compare_df_cols(df1, df2, df3)
```

```{r}
compare_df_cols(df1, df2, df3, return = "mismatch")
```

```{r}
compare_df_cols(df1, df2, df3, return = "mismatch", bind_method = "rbind")
```

`compare_df_cols_same()` returns `TRUE` or `FALSE` indicating if the `data.frames` can be successfully row-bound with the given binding method:

```{r}
compare_df_cols_same(df1, df3)
compare_df_cols_same(df2, df3)
```


# Exploring
## `tabyl()` - a better version of `table()`
`tabyl()` is a tidyverse-oriented replacement for `table()`. It counts combinations of one, two, or three variables, and then can be formatted with a suite of `adorn_*` functions to look just how you want. For instance:

```{r}
mtcars %>% 
    tabyl(gear, cyl) %>% 
    adorn_totals("col") %>% 
    adorn_percentages(denominator = "row") %>% 
    adorn_pct_formatting(digits = 2) %>% 
    adorn_ns() %>% 
    adorn_title()
```










