---
title: "RSQLite"
author: "yincy"
date: "12/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

RSQLite is the easiest way to use a database from R because the package itself contains SQLite; no external software is needed. This vignette will walk you through the basics of using a SQLite database.  

RSQLite is a DBI-compatible interface which means you primarily use functions defined in the DBI package, so you should always start by loading `DBI`, not `RSQLite`.  

```{r}
library(DBI)
library(RSQLite)
```


# Creating a new database  
To create a new SQLite database, you simply supply the filename to `dbConnect()`:  
```{r}
mydb <- dbConnect(drv = RSQLite::SQLite(), "my-db.sqlite")
```

If you just need a temporary database, use either `""` (for an on-disk database) or `":memory:"` or `"file::memory:"` (for an in-memory database). This database will be automatically deleted when you disconnect from it.  

```{r}
mydb <- dbConnect(drv = RSQLite::SQLite(), "")
dbDisconnect(mydb)
```


# Loading data  
you can easily copy an R data frame into a SQLite database with `dbWriteTable()`:  
```{r}
mydb <- dbConnect(drv = RSQLite::SQLite(), "")
dbWriteTable(conn = mydb, name = "mtcars", value = mtcars)
dbWriteTable(conn = mydb, name = "iris", value = iris)
dbListTables(conn = mydb)
```


# Queries  
Issue a query with `dbGetQuery()`:  
```{r}
dbGetQuery(conn = mydb, statement = "SELECT * FROM mtcars LIMIT 5")
```

Not all R variable names are valid SQL variable names, so you may need to escape them with `"`:  

```{r}
dbGetQuery(mydb, 
           statement = 'select * from iris where "Sepal.Length" < 4.6')
```

If you need to insert the value from a user into a query, don't use `paste()`! That makes it easy for malicious attacker to insert SQL that damage your database or reveal sensitive information. Instead, use a parameterrised query.  

```{r}
dbGetQuery(conn = mydb, 
           statement = 'select * from iris where "Sepal.Length" < :x', 
           params = list(x = 4.6))
```


# Batched queries  
If you run a query and the results don’t fit in memory, you can use `dbSendQuery()`, `dbFetch()` and `dbClearResults()` to retrieve the results in batches. By default `dbFetch()` will retrieve all available rows: use n to set the maximum number of rows to return.  

```{r}
rs <- dbSendQuery(conn = mydb, 
                  statement = 'select * from mtcars')

while(!dbHasCompleted(rs)){
    df <- dbFetch(rs, n = 10)
    print(nrow(df))
}
dbClearResult(rs)
```


# Multiple parameterised queries  
You can use the same approach to run the same parameterised query with different parameters. Call `dbBind()` to set the parameters:  

```{r}
rs <- dbSendQuery(conn = mydb, 
                  statement = 'select * from iris where "Sepal.Length" < :x')

dbBind(rs, params = list(x = 5))
nrow(dbFetch(rs))

dbBind(rs, params = list(x = 4.5))
nrow(dbFetch(rs))

dbClearResult(rs)
```

You can also pass multiple parameters in one call to `dbBind()`:  

```{r}
rs <- dbSendQuery(conn = mydb, 
                  statement = 'select * from iris where "Sepal.Length" < :x')

dbBind(rs, params = list(x = seq(4.5, 5, 0.1)))
nrow(dbFetch(rs))

dbClearResult(rs)
```


# Statements  
DBI has new functions `dbSendStatement()` and `dbExecute()`, which are the counterparts of `dbSendQuery()` and `dbGetQuery()` for SQL statements that do not return a tabular result, such as inserting records into a table, updating a table, or setting engine parameters. It is good practice, although currently not enforced, to use the new functions when you don’t expect a result.  

```{r}
dbExecute(conn = mydb, 
          statement = 'delete from iris where "Sepal.Length" < 4')

rs <- dbSendStatement(conn = mydb, 
                      statement = 'delete from iris where "Sepal.Length" < :x')

dbBind(rs, params = list(x = 4.5))
dbGetRowsAffected(res = rs)
dbClearResult(res = rs)
```



