---
title: "Visualizations"
author: "yincy"
date: "6/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `featurePlot` function is a wrapper for different `lattice` plots to visualize the data. For example, the following figures show the default plot for continuous outcomes generated using the `featurePlot` function.  

For classification data sets, the `iris` data are used for illustration.  

```{r}
str(iris)
```

**Scatterplot Matrix**  
```{r}
library(AppliedPredictiveModeling)
transparentTheme(trans = 0.4)
library(caret)

featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "pairs", 
            auto.key = list(columns = 3))
```

**Scatterplot Matrix with Ellipses**  
```{r}
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "ellipse",
            auto.key = list(columns = 3))
```

**Overlayed Density Plots**  
```{r}
transparentTheme(trans = 0.9)
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "density", 
            scales = list(x = list(relation = "free"), 
                          y = list(relation = "free")), 
            adjust = 1.5, 
            pch = "|", 
            layout = c(4, 1), 
            auto.key = list(columns = 3))
```

**Box Plots**  
```{r}
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "box", 
            scales = list(x = list(rot = 90), 
                          y = list(relation = "free")), 
            layout = c(4, 1), 
            auto.key = list(columns = 2))
```

**Scatter Plots**
```{r}
library(mlbench)
data("BostonHousing")

regVar <- c("age", "lstat", "tax")
str(BostonHousing[, regVar])
```

When the predictors are continuous, featurePlot can be used to create scatter plots of each of the predictors with the outcome.  

```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col = rgb(0.2, 0.2, 0.2, 0.4)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(1, 0, 0, 0.7)
theme1$plot.line$lwd = 2
trellis.par.set(theme1)

featurePlot(x = BostonHousing[, regVar], 
            y = BostonHousing$medv, 
            plot = "scatter", 
            layout = c(3, 1))
```

Note that the x-axis scales are different. The function automatically uses `scales = list(y = list(relation = "free"))` so you don’t have to add it. We can also pass in options to the `lattice` function `xyplot`. For example, we can add a scatter plot smoother by passing in new options:  

```{r}
featurePlot(x = BostonHousing[, regVar], 
            y = BostonHousing$medv, 
            plot = "scatter", 
            type = c("p", "smooth"), 
            span = 0.5, 
            layout = c(3, 1))
```

The options `degree` and `span` control the smoothness of the smoother.  

